### 2025-01-31 products/manuscript/Manuscript.qmd
initializing project repo

---
title: "Prasanga Paudel- Data Analysis Project"
author: ""
format:
  html:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/american-journal-of-epidemiology.csl
---

The structure below is one possible setup for a manuscript, or a general data analysis project (including the course project). Adjust as needed. 
You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses HTML as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like docx or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

**Authors**

* First Autor$^{1,2,*}$ (ORCID: 0000-0000-1234-5678) 
* Second Author$^{3, *}$
* Third Autor$^{3, \dagger}$
* Last Author$^{1, 2, \land}$

**Author affiliations**

1. College of Public Health, University of Georgia, Athens, GA, USA.
2. Another Affiliation, Athens, GA, USA.
3. Yet another Affiliation.

$*$ These authors contributed equally to this work.

$\land$ Corresponding author: some@email.com

$\dagger$ Disclaimer: The opinions expressed in this article are the
author's own and don't reflect their employer.

{{< pagebreak >}}

# Summary/Abstract
_The data is based on World Development Indicator from World Bank Group. This is a comprehensive database that includes information including employment, public and private expenditure in health sector, fertility, immunization, literacy, mortality, and many more.The dataset has a total of 1496 variables and across 11 years from 2010 to 2020. The actual source of data has information from 1960 to 2023 but to deal with the problem of missing values in recent years and overcomplication in analysis due to too many year data , we will focus on 10 recent years between 2010 and 2020._

{{< pagebreak >}}

# Introduction

## General Background Information
_We will try to analyze the most influential factors affecting the expenditure in health sector by different countries. The predictor variables will include factors ranging from education, mortality, population distribution among rural and urban spaces, and so on. We would use multiple regression analysis with health sector expenditure as dependent variable and the variable of interests as predictors._

_We are using data from OECD members for analysis, and there have already been some research which have focused on GDP as predictor for health sector expenditure. Some research have also focused on the relation between life expectancy and health expenditure. However, these research were based on fixed effect and random effect models. In our analysis we will try to follow OLS based multiple regression model._

## Description of data and data source
_This is a time series data based on 38 member countries from the Organisation for Economic Co-operation and Development (OECD). This organization is a unique forum with governments from 37 democracies collaborate to develop policy standards to promote sustainable economic growth. However, the data obtained from this dataset is crude and needs to be treated before attempting analysis based on it. For example, in its present form, the variable name is a long sentense describing the data it contains._

## Objective of the research
_The main objective of this research is to:_

_1. To analyze the factors that affect the expenditure in health sector among OECD countries._

_2. To analyze the magnitude of the influence of these factors on health expenditure_

To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above and have the right bibtex key. Then you can include like this:

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].

{{< pagebreak >}}

## Data aquisition
_The data for this research is available at the official World Bank- Data Bank website.It is a free database managed by world bank group and can be accessed anytime. We can apply various filters as per our need and export the data as a .CSV or an .xlsx file._

Link to the source: <https://databank.worldbank.org/source/world-development-indicators/Type/TABLE/preview/on#>

# Methods

_To be added_

## Schematic of workflow

Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it). @fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender.
We store those figures in the `assets` folder.

```{r}
#| label: fig-schematic
#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
#| echo: FALSE
knitr::include_graphics(here("assets","antigen-recognition.png"))
```

## Data import and cleaning
_Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along._

## Statistical analysis
_Explain anything related to your statistical analyses._

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below. I generally recommend the `here` package.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/tables/summarytable.rds")
knitr::kable(resulttable)
```

## Basic statistical analysis

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results","figures","height-weight-stratified.png"))
```

## Full analysis

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like.

{{< pagebreak >}}

# References

### 2025-01-31 products/manuscript/Manuscript.qmd
changes in data summary

@@
 library(knitr)
+library(here)
+library(readxl)
 ```
@@
 However, the data obtained from this dataset is crude and needs to be treated before attempting analysis based on it.
-For example, in its present form, the variable name is a long sentense describing the data it contains._
+For example, in its present form, the variable name is a long sentense describing the data it contains.The years are listed as column names and variables are listed as Series._
 
@@
 ## Data import and cleaning
-_Write code that reads in the file and cleans it so it's ready for analysis.
-Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts.
-If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g.
-as supplement in a paper).
-All materials, including files that contain code, should be commented well so everyone can follow along._
+```{r}
+file_path <- here("data","raw-data", "wdi-oecd.xlsx")
+data <- read_excel(file_path, sheet="Data")
+```
 
 ## Statistical analysis
-_Explain anything related to your statistical analyses._
+
+_This is the snapshot of the data as of now._
+```{r}
+head(data)
+```
 

### 2025-01-31 products/manuscript/Manuscript.qmd
updates to page title

@@
 ---
-title: "Prasanga Paudel- Data Analysis Project"
-author: ""
+title: "Factors affecting health sector expenditure in OECD countries"
+author: "Prasanga Paudel"
 format:
@@
 
-* First Autor$^{1,2,*}$ (ORCID: 0000-0000-1234-5678) 
+* Prasanga Paudel$^{1,2,*}$ (ORCID: 0000-0000-1234-5678) 
 * Second Author$^{3, *}$
@@
 
-1. College of Public Health, University of Georgia, Athens, GA, USA.
+1. College of Agricultural and Environmental Science, University of Georgia, Athens, GA, USA.
 2. Another Affiliation, Athens, GA, USA.
@@
 The predictor variables will include factors ranging from education, mortality, population distribution among rural and urban spaces, and so on.
-We would use multiple regression analysis with health sector expenditure as dependent variable and the variable of interests as predictors._
+We would use multiple regression analysis with health sector expenditure as dependent variable and the variable of interests as predictors for this panel data._
 
@@
 However, these research were based on fixed effect and random effect models.
-In our analysis we will try to follow OLS based multiple regression model._
+In our analysis we will try to follow OLS multiple regression model._
 
@@
 ## Objective of the research
+
+The expenditure in health sector is one of the important factor that helps in determining the efficiency of health sector.
+We can relate the expenditure with different national development indicators and hope to find important connections between them.
+This will help government to make specific interventions through their policy to enhance their health sector.
+We can also find out how countries are currently operating and which factor has helped or negatively affected their health sector.
+
 _The main objective of this research is to:_
@@
 
+## Data import and cleaning
+```{r}
+file_path <- here("data","raw-data", "wdi-oecd.xlsx")
+data <- read_excel(file_path, sheet="Data")
+```
+
+## Statistical analysis
+
+_This is the snapshot of the data as of now._
+```{r}
+head(data)
+```
+
 ## Schematic of workflow
@@
 knitr::include_graphics(here("assets","antigen-recognition.png"))
-```
-
-## Data import and cleaning
-```{r}
-file_path <- here("data","raw-data", "wdi-oecd.xlsx")
-data <- read_excel(file_path, sheet="Data")
-```
-
-## Statistical analysis
-
-_This is the snapshot of the data as of now._
-```{r}
-head(data)
 ```

### 2025-02-21 code/eda-code/eda.qmd
work #2

---
title: "Codes for exploratory analysis"
date: "2024-02-17"
author: "Prasanga Paudel"
output: html_document
---

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
library(patchwork)
```

# Loading the data.

We will import the processed data from the processed-data folder which we is the cleaned and transformed dataset.

```{r}
#Path to data. 
data_location <- here::here("data","processed-data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

# Data exploration through tables

Here we will present the summary stat for some of the interesting varible based on year 2010 for the OECD countries.

```{r}
library(dplyr)
library(knitr)
library(kableExtra)

# Filtering data for the year 2010
data_2010 <- mydata %>%
  filter(Year == 2010)

# Create a summary table, excluding specific rows
country_summary_table <- data_2010 %>%
  filter(!CountryName %in% c("Data from database: World Development Indicators", "OECD members"
                              "Last Updated: 01/28/2025")) %>%
  select(
    CountryName,
    GDP_current_US,
    Population_total,
    Life_expectancy_at_birth_total_years,
    Physicians_per_1000_people,
    Hospital_beds_per_1000_people,
    Government_Effectiveness_Estimate,
    Control_of_Corruption_Estimate,
    Rule_of_Law_Estimate,
    Access_to_electricity__of_population,
    Unemployment_total__of_total_labor_force_modeled_ILO_estimate,
    Poverty_headcount_ratio_at_national_poverty_lines__of_population
  )

# Create and display an HTML table in R
country_summary_table %>%
  kable("html", escape = FALSE, digits = 2) %>%
  kable_styling(full_width = F, position = "left") %>%
  column_spec(1, bold = TRUE) %>%  # Making CountryName column bold
  add_header_above(c(" " = 1, "National Development Indicators" = 11)) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

```

# Data exploration through figures

In this section, we will explore some of the variables and try to look at some of the similarity among the varibles.

First, we will see how "Health Expenditure as a Percentage of GDP Over Time (2010–2020)" is distributed over the years for top 5 OECD countries.The top 5 countries are calculated based on 2010 data.  _Note: The aggregate values visible as "OECD members" will not be used for analysis purpose. It is just for demonstration._

```{r}
# Converting column to numeric before plotting
mydata <- mydata %>%
  mutate(Current_health_expenditure__of_GDP = as.numeric(Current_health_expenditure__of_GDP))

# Filtering top 5 countries based on 2010 health expenditure
top_5_countries <- mydata %>%
  filter(Year == 2010) %>%
  arrange(desc(Current_health_expenditure__of_GDP)) %>%
  slice_head(n = 5) %>%
  pull(CountryName)  # Extracting country names

# Filtering data for only 5 countries over time
data_filtered <- mydata %>%
  filter(CountryName %in% top_5_countries)

y_breaks <- seq(min(data_filtered$Current_health_expenditure__of_GDP, na.rm = TRUE), 
                max(data_filtered$Current_health_expenditure__of_GDP, na.rm = TRUE), 
                length.out = 4)

# Creating the line plot
ggplot(data_filtered, aes(x = Year, y = Current_health_expenditure__of_GDP, color = CountryName, group = CountryName)) +
  geom_line(size = 1) +
  geom_point(size = 2) +  
  labs(
    title = "Current Health Expenditure as a % of GDP (2010–2020): Top 5 nations",
    x = "Year",
    y = "Health Expenditure (% of GDP)",
    color = "Country"
  ) +
  scale_y_continuous(breaks = y_breaks) +  
  theme_minimal()

ggsave(
  filename = here("results","figures", "health_expenditure_plot.png"),  
  plot = last_plot(),  # Saves the last plotted figure
  width = 8, 
  height = 6, 
  dpi = 300
)

```

Now, we will try to see if there is any relation between GDP per capita (current US) and current Health Expenditure as a % of GDP.

```{r}

# Converting columns to numeric 
mydata <- mydata %>%
  mutate(
    Current_health_expenditure__of_GDP = as.numeric(Current_health_expenditure__of_GDP),
    GDP_per_capita_current_US = as.numeric(GDP_per_capita_current_US)
  )

# Filtering data for the year 2010
data_2010 <- mydata %>%
  filter(Year == 2010)

# Creating scatter plot
ggplot(data_2010) +

  # Plot GDP per capita 
  geom_point(aes(x = GDP_per_capita_current_US, y = Current_health_expenditure__of_GDP), 
             color = "blue", alpha = 0.7, size = 3) +

  # Plot Health Expenditure 
  geom_point(aes(x = GDP_per_capita_current_US, y = Current_health_expenditure__of_GDP), 
             color = "red", alpha = 0.7, size = 3, shape = 17) +  # Shape 17 = Triangle

  # Adding a green regression line
  geom_smooth(aes(x = GDP_per_capita_current_US, y = Current_health_expenditure__of_GDP), 
              method = "lm", color = "green", se = FALSE, size = 1) +

  # Labels and title
  labs(
    title = "Health Expenditure vs GDP Per Capita (2010)",
    x = "GDP Per Capita (Current US$)",
    y = "Current Health Expenditure (% of GDP)"
  ) +

  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

ggsave(
  filename = here("results","figures", "health_expenditure_gdp.png"),  
  plot = last_plot(),  # Saves the last plotted figure
  width = 8, 
  height = 6, 
  dpi = 300
)
```

Now, we will try to see if there is any relation between hospital beds per 1000 people and current Health Expenditure as a percentage of GDP.

```{r}

# Converting columns to numeric
mydata <- mydata %>%
  mutate(
    Current_health_expenditure__of_GDP = as.numeric(Current_health_expenditure__of_GDP),
    Hospital_beds_per_1000_people = as.numeric(Hospital_beds_per_1000_people)
  )

# Filtering data for the year 2010
data_2010 <- mydata %>%
  filter(Year == 2010)

# Creating scatter plot
ggplot(data_2010) +

  # Plot GDP per capita 
  geom_point(aes(x = Hospital_beds_per_1000_people, y = Current_health_expenditure__of_GDP), 
             color = "blue", alpha = 0.7, size = 3) +

  # Plot Health Expenditure 
  geom_point(aes(x = Hospital_beds_per_1000_people, y = Current_health_expenditure__of_GDP), 
             color = "red", alpha = 0.7, size = 3, shape = 17) +  # Shape 17 = Triangle

  # Adding green regression line
  geom_smooth(aes(x = Hospital_beds_per_1000_people, y = Current_health_expenditure__of_GDP), 
              method = "lm", color = "green", se = FALSE, size = 1) +

  # Labels and title
  labs(
    title = "Health Expenditure vs GDP Per Capita (2010)",
    x = "Hospital beds per 1000 people",
    y = "Health Expenditure (% of GDP)"
  ) +

  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

ggsave(
  filename = here("results","figures", "hospital_per_bed.png"),  
  plot = last_plot(),  # Saves the last plotted figure
  width = 8, 
  height = 6, 
  dpi = 300
)
```

Let's check if annual population growth shows any relation with current Health Expenditure as a percentage of GDP for Australia

```{r}

# Filtering data for Australia
australia_data <- mydata %>%
  filter(CountryName == "Australia")

# Histogram for Population Growth
plot_pop_growth <- ggplot(australia_data, aes(x = as.factor(Year), y = Population_growth_annual_)) +
  geom_col(fill = "blue", color = "black", alpha = 0.7) +
  labs(
    title = "Annual Population Growth \n Country: Australia",
    x = "Year",
    y = "Annual Population Growth (%)"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(
  filename = here("results","figures", "pop_growth_aus.png"),  
  plot = last_plot(),  # Saves the last plotted figure
  width = 8, 
  height = 6, 
  dpi = 300
)

# Histogram for Health Expenditure
plot_health_expenditure <- ggplot(australia_data, aes(x = as.factor(Year), y = Current_health_expenditure__of_GDP)) +
  geom_col(fill = "red", color = "black", alpha = 0.7) +
  labs(
    title = "Health Expenditure (% of GDP) \n Country: Australia",
    x = "Year",
    y = "Health Expenditure (% of GDP)"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(
  filename = here("results","figures", "health_exp_aus.png"),  
  plot = last_plot(),  
  width = 8, 
  height = 6, 
  dpi = 300
)

# Combine both plots side by side
plot_pop_growth + plot_health_expenditure

```
We can not see a visible relation between these two factors.  Let's try it for France now

```{r}

# Filter data for France
france_data <- mydata %>%
  filter(CountryName == "France")

# Histogram for Population Growth
plot_pop_growth <- ggplot(france_data, aes(x = as.factor(Year), y = Population_growth_annual_)) +
  geom_col(fill = "blue", color = "black", alpha = 0.7) +
  labs(
    title = "Annual Population Growth \n Country: France",
    x = "Year",
    y = "Annual Population Growth (%)"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(
  filename = here("results","figures", "pop_growth_france.png"),  
  plot = last_plot(),  
  width = 8, 
  height = 6, 
  dpi = 300
)

# Histogram for Health Expenditure
plot_health_expenditure <- ggplot(australia_data, aes(x = as.factor(Year), y = Current_health_expenditure__of_GDP)) +
  geom_col(fill = "red", color = "black", alpha = 0.7) +
  labs(
    title = "Health Expenditure (% of GDP) \n Country: France",
    x = "Year",
    y = "Health Expenditure (% of GDP)"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(
  filename = here("results","figures", "health_exp_france.png"),  
  plot = last_plot(),  
  width = 8, 
  height = 6, 
  dpi = 300
)

# Combining both plots side by side
plot_pop_growth + plot_health_expenditure
```
We can not observe a distinct relation between the two variable, we can observe that the Health Expenditure (% of GDP) variable also show less fluctuation across the years for multiple countries. Lets see if Health Expenditure Per Capita (USD) over the Years shows any reation with annual population growth.

```{r}

# Histogram for Population Growth
plot_pop_growth <- ggplot(australia_data, aes(x = as.factor(Year), y = Population_growth_annual_)) +
  geom_col(fill = "blue", color = "black", alpha = 0.7) +
  labs(
    title = "Annual Population Growth \n Country: Australia",
    x = "Year",
    y = "Annual Population Growth (%)"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))

# Histogram for Health Expenditure Per Capita (Current USD)
plot_health_expenditure <- ggplot(australia_data, aes(x = as.factor(Year), y = Current_health_expenditure_per_capita_current_US)) +
  geom_col(fill = "red", color = "black", alpha = 0.7) +
  labs(
    title = "Health Expenditure Per Capita (USD) \n Country: Australia",
    x = "Year",
    y = "Health Expenditure Per Capita (USD)"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

ggsave(
  filename = here("results","figures", "health_exp_per_capita_aus.png"),  
  plot = last_plot(),  
  width = 8, 
  height = 6, 
  dpi = 300
)

# Combine both plots side by side
plot_pop_growth + plot_health_expenditure

```
The Health Expenditure Per Capita (USD) varriable in fact shows more variation and also show some connections with annual population growth.
```{r}

# Histogram for Population Growth
plot_pop_growth <- ggplot(france_data, aes(x = as.factor(Year), y = Population_growth_annual_)) +
  geom_col(fill = "blue", color = "black", alpha = 0.7) +
  labs(
    title = "Annual Population Growth \n Country: France",
    x = "Year",
    y = "Annual Population Growth (%)"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

# Histogram for Health Expenditure Per Capita (Current USD)
plot_health_expenditure <- ggplot(france_data, aes(x = as.factor(Year), y = Current_health_expenditure_per_capita_current_US)) +
  geom_col(fill = "red", color = "black", alpha = 0.7) +
  labs(
    title = "Health Expenditure Per Capita (USD) \n Country: France",
    x = "Year",
    y = "Health Expenditure Per Capita (USD)"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(
  filename = here("results","figures", "health_exp_per_capita_france.png"),  
  plot = last_plot(),  
  width = 8, 
  height = 6,
  dpi = 300
)

# Combine both plots side by side
plot_pop_growth + plot_health_expenditure
```
The results for Australia can also confirmed for France, we may be able to use Health Expenditure Per Capita (USD) as a dependent variabe in our future analysis.

# Notes

For your own explorations, tables and figures can be "quick and dirty". As long as you can see what's going on, there is no need to polish them. That's in contrast to figures you'll produce for your final products (paper, report, presentation, website, etc.). Those should look as nice, polished and easy to understand as possible.

### 2025-02-21 code/processing-code/processingfile-v1.qmd
work #2

---
title: "Data cleaning for Project"
author: "Prasanga Paudel"
date: "2024-02-17"
output: html_document
---

```{r}
#loading the necessary packages
library(here)
library(dplyr)
library(skimr)
library(ggplot2)
library(readxl)
library(tidyr)
```

## Data import and cleaning
Here, we will import the original file downloaded from the website.

```{r}
file_path <- here("data","raw-data", "wdi-oecd.xlsx")
data <- read_excel(file_path, sheet="Data")
```

## Preliminary check of the original data

_This is the snapshot of the data as of now._
```{r}
head(data)
```

## Data Cleaning

We will now change the name of the variables so that R can better understand the names and to avoid errors during the data analysis process.

```{r}
# Cleaning year columns: Removing brackets, text, and spaces
colnames(data) <- gsub("\\[.*\\]|\\s", "", colnames(data))

# We will now check the corrected names
print(colnames(data))
```
Here, we can observe that the names have been assigned as required and we no longer have spaces and unwanted words and characters in the variable names.

## Data Transformation

As the necesary variables require to go further with analysis are present in rows under the column "SeriesName", we need to make sure that we have these variables as column names and not as rows. We will now perform the necessary transformations. We will also need to transform the year values from column names to rows as it would make more convinient to run a time series analysis. We will therefore create a new variable Year to save the values for different years.

```{r}
# Reshaping the data from wide to long format
data_long <- data %>%
  pivot_longer(
    cols = starts_with("20"),  # Selecting all year columns (2010 to 2020)
    names_to = "Year",         # Naming the new column for years
    values_to = "Value"        # Naming the new column for values
  )

# Viewing the reshaped data
print(head(data_long))
```
As we can see the year data has been arranged as required. Now, we will transform the variables into column names. But, before doing that, we will remove all the special characters from the to-be variable names, and also replace spaces with unerscore.

```{r}
# Cleaning the Series_Name column
data_long <- data_long %>%
  mutate(
    SeriesName = gsub("[^a-zA-Z0-9 ]", "", SeriesName),  # Removing special characters
    SeriesName = gsub(" ", "_", SeriesName)               # Replacing spaces with _
  )

# View the updated Series_Name column
print(unique(data_long$SeriesName))
```
Now, the variables are ready to be transformed.

We will further insure that no replication has been produuced during out data analysis process.

```{r}
# Remove duplicates
data_long <- data_long %>%
  distinct(CountryName, CountryCode, SeriesCode, Year, SeriesName, .keep_all = TRUE)
```

Here, we will transform the variables as column names with appropriates values assigned to them correctly.
```{r}
# Spread the SeriesName into separate columns
data_final <- data_long %>%
  pivot_wider(
    names_from = SeriesName,  # Use SeriesName to create new columns
    values_from = Value       # Fill the new columns with the corresponding values
  )

```

## Checking if the transformation is accurate

In this section, we will verify if our transformation of the dataset was accurate.

```{r}
# Extract value from the original dataset
original_value <- data %>%
  filter(CountryName == "Hungary", SeriesName == "Primary education, teachers") %>%
  pull(`2010`)  # Replace `2010` with the correct column name for the year 2010

# Print the original value
print(original_value)
```
This provides the value for Primary education, teachers for Hungary based on original dataset.

```{r}
# Convert Year to numeric
data_final <- data_final %>%
  mutate(Year = as.numeric(Year))

# Extract the value
transformed_value <- data_final %>%
  filter(CountryName == "Hungary", Year == 2010) %>%
  pull(Primary_education_teachers)

# Remove missing values and print non-missing values
non_missing_values <- na.omit(transformed_value)
cat(non_missing_values, sep = "\n")
```
This provides the value for Primary education, teachers for Hungary based on transformed dataset. 
Since they are equal, we have verified that our transformation was accurate.

## Creating a more compact dataset

As the dataset has 10 years information for each country and currently this structure has resulted in creation of 641,817 observations, we will try to make it more compact and less complex.

```{r}
data_collapsed <- data_final %>%
  group_by(CountryName, CountryCode, Year) %>%
  summarize(across(everything(), ~ max(.x, na.rm = TRUE)), .groups = "drop")

```

## Saving the dataset for EDA

Now, we will save the dataset so that we will be able run EDA based on it.

```{r}
rds_path <- here("data", "processed-data", "processeddata.rds")
saveRDS(data_collapsed, rds_path)
```

### 2025-02-21 products/manuscript/Manuscript.qmd
changes to rendered files
work #2

@@
 
-The structure below is one possible setup for a manuscript, or a general data analysis project (including the course project). Adjust as needed. 
-You don't need to have exactly these sections, but the content covering those sections should be addressed.
-
 This uses HTML as output format.
@@
 library(readxl)
+library(writexl)
 ```
@@
 
-## Data import and cleaning
+# Data import and cleaning
 ```{r}
@@
 
-## Statistical analysis
+## A look into the original for of dataset
 
@@
 
-## Schematic of workflow
+# EDA based on clean dataset
 
-Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it).
-@fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender.
-We store those figures in the `assets` folder.
-
-```{r}
-#| label: fig-schematic
-#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
-#| echo: FALSE
-knitr::include_graphics(here("assets","antigen-recognition.png"))
-```
+## Descriptive Analysis
 
@@
 
-@tbl-summarytable shows a summary of the data.
+### Looking into the trend for our possible dependent variable
 
-Note the loading of the data providing a **relative** path using the `../../` notation.
-(Two dots means a folder up).
-You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path.
-You can also use the `here` R package to create paths.
-See examples of that below.
-I generally recommend the `here` package.
+@fig-health-expenditure-plot how "Health Expenditure as a Percentage of GDP Over Time (2010–2020)" is distributed over the years for top 5 OECD countries.
 
 ```{r}
-#| label: tbl-summarytable
-#| tbl-cap: "Data summary table."
+#| label: fig-health-expenditure-plot
+#| fig-cap: "Current Health Expenditure as a % of GDP (2010–2020): Top 5 nations"
 #| echo: FALSE
-resulttable=readRDS("../../results/tables/summarytable.rds")
-knitr::kable(resulttable)
+knitr::include_graphics(here("results","figures","health_expenditure_plot.png"))
+```
+
+@fig-health-exp-aus how "Health Expenditure as a Percentage of GDP changes over time for Australia.
+
+```{r}
+#| label: fig-health-exp-aus
+#| fig-cap: "Health Expenditure (% of GDP), Country: Australia"
+#| echo: FALSE
+knitr::include_graphics(here("results", "figures", "health_exp_aus.png"))
+```
+
+@fig-pop-growth-aus checks if Health Expenditure as a Percentage of GDP over the Years shows any reation with annual population growth for Australia.
+
+```{r}
+#| label: fig-pop-growth-aus
+#| fig-cap: "Health Expenditure (% of GDP), Country: Australia"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","pop_growth_aus.png"))
+```
+
+@fig-he_per_capita checks if Health Expenditure Per Capita (USD) over the Years shows any reation with annual population growth for Australia.
+
+```{r}
+#| label: fig-he_per_capita
+#| fig-cap: "Health Expenditure Per Capita (USD, Country: Australia"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","health_exp_per_capita_aus.png"))
 ```
@@
 
-_To get some further insight into your data, if reasonable you could compute simple statistics (e.g.
-simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable.
-Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._
-
-@fig-result shows a scatterplot figure produced by one of the R scripts.
-
-```{r}
-#| label: fig-result
-#| fig-cap: "Height and weight stratified by gender."
-#| echo: FALSE
-knitr::include_graphics(here("results","figures","height-weight-stratified.png"))
-```
+_To be added_
 
@@
 
-_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
-This might again be code that is best placed in one or several separate R scripts that need to be well documented.
-You want the code to produce figures and data ready for display as tables, and save those.
-Then you load them here._
-
-Example @tbl-resulttable2 shows a summary of a linear model fit.
-
-```{r}
-#| label: tbl-resulttable2
-#| tbl-cap: "Linear model fit table."
-#| echo: FALSE
-resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
-knitr::kable(resulttable2)
-```
+_To be added._
 
@@
 ## Summary and Interpretation
-_Summarize what you did, what you found and what it means._
+_To be added._
 
 ## Strengths and Limitations
-_Discuss what you perceive as strengths and limitations of your analysis._
+_To be added._
 
 ## Conclusions
-_What are the main take-home messages?_
+_To be added._
 

### 2025-02-21 code/eda-code/edacode.R

@@ -1,44 +1,0 @@
-## ---- packages --------
-#load needed packages. make sure they are installed.
-library(here) #for data loading/saving
-library(dplyr)
-library(skimr)
-library(ggplot2)
-
-## ---- loaddata --------
-#Path to data. Note the use of the here() package and not absolute paths
-data_location <- here::here("data","processed-data","processeddata.rds")
-#load data
-mydata <- readRDS(data_location)
-
-## ---- table1 --------
-summary_df = skimr::skim(mydata)
-print(summary_df)
-# save to file
-summarytable_file = here("results","tables", "summarytable.rds")
-saveRDS(summary_df, file = summarytable_file)
-
-## ---- height --------
-p1 <- mydata %>% ggplot(aes(x=Height)) + geom_histogram() 
-plot(p1)
-figure_file = here("results", "figures", "height-distribution.png")
-ggsave(filename = figure_file, plot=p1) 
-
-## ---- weight --------
-p2 <- mydata %>% ggplot(aes(x=Weight)) + geom_histogram() 
-plot(p2)
-figure_file = here("results", "figures", "weight-distribution.png")
-ggsave(filename = figure_file, plot=p2) 
-
-## ---- fitfig1 --------
-p3 <- mydata %>% ggplot(aes(x=Height, y=Weight)) + geom_point() + geom_smooth(method='lm')
-plot(p3)
-figure_file = here("results","figures", "height-weight.png")
-
-## ---- fitfig2 --------
-p4 <- mydata %>% ggplot(aes(x=Height, y=Weight, color = Gender)) + geom_point() + geom_smooth(method='lm')
-plot(p4)
-figure_file = here("results","figures", "height-weight-stratified.png")
-ggsave(filename = figure_file, plot=p4) 
-
-

### 2025-02-21 code/processing-code/processingcode.R

@@ -1,138 +1,0 @@
-###############################
-# processing script
-#
-#this script loads the raw data, processes and cleans it 
-#and saves it as Rds file in the processed-data folder
-#
-# Note the ## ---- name ---- notation
-# This is done so one can pull in the chunks of code into the Quarto document
-# see here: https://bookdown.org/yihui/rmarkdown-cookbook/read-chunk.html
-###############################
-
-
-## ---- packages --------
-#load needed packages. make sure they are installed.
-library(readxl) #for loading Excel files
-library(dplyr) #for data processing/cleaning
-library(tidyr) #for data processing/cleaning
-library(skimr) #for nice visualization of data 
-library(here) #to set paths
-
-
-## ---- loaddata --------
-#path to data
-#note the use of the here() package and not absolute paths
-data_location <- here::here("data","raw-data","exampledata.xlsx")
-#load data. 
-#note that for functions that come from specific packages (instead of base R)
-# I often specify both package and function like so
-#package::function() that's not required one could just call the function
-#specifying the package makes it clearer where the function "lives",
-#but it adds typing. You can do it either way.
-rawdata <- readxl::read_excel(data_location)
-# We might also want to load the codebook to look at it
-codebook <- readxl::read_excel(data_location, sheet ="Codebook")
-
-
-## ---- exploredata --------
-#take a look at the data
-dplyr::glimpse(rawdata)
-#another way to look at the data
-summary(rawdata)
-#yet another way to get an idea of the data
-head(rawdata)
-#this is a nice way to look at data
-skimr::skim(rawdata)
-#look in the Codebook for a variable explanation
-print(codebook)
-
-
-## ---- cleandata1 --------
-# Inspecting the data, we find some problems that need addressing:
-# First, there is an entry for height which says "sixty" instead of a number. 
-# Does that mean it should be a numeric 60? It somehow doesn't make
-# sense since the weight is 60kg, which can't happen for a 60cm person (a baby)
-# Since we don't know how to fix this, we might decide to remove the person.
-# This "sixty" entry also turned all Height entries into characters instead of numeric.
-# That conversion to character also means that our summary function isn't very meaningful.
-# So let's fix that first.
-d1 <- rawdata %>% dplyr::filter( Height != "sixty" ) %>% 
-                  dplyr::mutate(Height = as.numeric(Height))
-# look at partially fixed data again
-skimr::skim(d1)
-hist(d1$Height)
-
-
[…]

### 2025-02-22 code/eda-code/eda.qmd
#3

@@
 library(patchwork)
+library(kableExtra)
+library(htmltools)
+library(magrittr)
+library(webshot2)
 ```
@@
 ```{r}
-library(dplyr)
-library(knitr)
-library(kableExtra)
 
@@
 country_summary_table <- data_2010 %>%
-  filter(!CountryName %in% c("Data from database: World Development Indicators", "OECD members"
-                              "Last Updated: 01/28/2025")) %>%
+  filter(!CountryName %in% c("Data from database: World Development Indicators", "OECD members",        "Last Updated: 01/28/2025")) %>%
   select(
@@
 
-# Create and display an HTML table in R
-country_summary_table %>%
+# Generate the HTML table
+table_html <- country_summary_table %>%
   kable("html", escape = FALSE, digits = 2) %>%
-  kable_styling(full_width = F, position = "left") %>%
-  column_spec(1, bold = TRUE) %>%  # Making CountryName column bold
+  kable_styling(full_width = FALSE, position = "left") %>%
+  column_spec(1, bold = TRUE) %>%
   add_header_above(c(" " = 1, "National Development Indicators" = 11)) %>%
   kable_styling(bootstrap_options = c("striped", "hover"))
+
+# Save the table object as an RDS file
+rds_file <- here::here("results", "tables", "health_exp_2010.rds")
+saveRDS(table_html, file = rds_file)
 

### 2025-02-22 products/manuscript/Manuscript.qmd
#3

@@
     toc: false
+    embed-resources: true
     number-sections: true
@@
 library(writexl)
+library(magrittr)
+library(patchwork)
+library(kableExtra)
 ```
@@
 * Prasanga Paudel$^{1,2,*}$ (ORCID: 0000-0000-1234-5678)
-* Second Author$^{3, *}$
-* Third Autor$^{3, \dagger}$
-* Last Author$^{1, 2, \land}$
 
@@
 1. College of Agricultural and Environmental Science, University of Georgia, Athens, GA, USA.
-2. Another Affiliation, Athens, GA, USA.
-3. Yet another Affiliation.
 
@@
 
-_To be added_
+We wll be following a panel data time series regression model. The details will be discussed as we progress.
 
-# Data import and cleaning
+# Importing the original dataset
+
 ```{r}
@@
 
-## A look into the original for of dataset
+## A look into the original format of the dataset
 
-_This is the snapshot of the data as of now._
+_This was the state of the data when first downloaded_
 ```{r}
@@
 
-# EDA based on clean dataset
+# Importing the cleaned dataset
 
-## Descriptive Analysis
+```{r}
+#Path to data. 
+data_location <- here::here("data","processed-data","processeddata.rds")
+#load data
+mydata <- readRDS(data_location)
+```
+
+## A look into the cleaned dataset
+
+_This is the present state of the dataset_
+
+```{r}
+head(mydata)
+```
 
@@
 
+In exploratory data analysis, we will just look at data from Australia as a reference for relation between variables in most of the cases.
+We are currently thinking of using "Current Health Expenditure as a % of GDP" as our dependent variable.
+But, we will see the it's distribution and connections with other variables before making decision.
+
 ## Exploratory/Descriptive analysis
 
-_Use a combination of text/tables/figures to explore and describe your data.
-Show the most important descriptive results here.
-Additional ones should go in the supplement.
-Even more can be in the R and Quarto files that are part of your project._
+_Here, we will now use a combination of tables and figues to have a look into the data._
+
+As this is a time series data, it is quite difficult to visualize all the variables across the years for all the countries.
+So, we will try to keep it simple and see simple relations between variables within a year or trend of a particular variable for some countries across the years.
+
+@tbl-summary shows the values under interesting variables for all OECD countries for the year 2010.
+
+```{r}
+#| label: tbl-summary
+#| tbl-cap: "Summary of variables"
+#| echo: FALSE
+table_html <- readRDS(here::here("results", "tables", "health_exp_2010.rds"))
+table_html
+
+```
 
@@
 
-@fig-health-expenditure-plot how "Health Expenditure as a Percentage of GDP Over Time (2010–2020)" is distributed over the years for top 5 OECD countries.
+@fig-health-expenditure-plot shows how "Health Expenditure as a Percentage of GDP Over Time (2010–2020)" is distributed over the years for top 5 OECD countries.
 
@@
 
-@fig-health-exp-aus how "Health Expenditure as a Percentage of GDP changes over time for Australia.
+@fig-health-exp-aus shows the distribution of "Health Expenditure as a Percentage of GDP changes over time for Australia.
 
@@
 
-@fig-pop-growth-aus checks if Health Expenditure as a Percentage of GDP over the Years shows any reation with annual population growth for Australia.
+@fig-pop-growth-aus shows the distribution  of Health Expenditure as a Percentage of GDP over the Years Comparing with shows any reation with annual population growth for Australia.This figure shows that it has very less in common with @fig-health-exp-aus as long as distribution and correlation is concerned.
 
@@
 
-@fig-he_per_capita checks if Health Expenditure Per Capita (USD) over the Years shows any reation with annual population growth for Australia.
+@fig-he_per_capita shows the distribution of Health Expenditure Per Capita (USD) over the Years.
+It shows that it is some what related to @fig-pop-growth-aus in it's distribution.
+So we can say that there is indeed a presence of some pattern in the variables of interest.
 
@@
 ```
+ I think it will be better for us to use the variable "Health Expenditure Per Capita $USD" as a dependent variable in our analysis as it show more variation and connections with other variables.
 

### 2025-03-12 code/analysis-code/statistical-analysis.qmd
project part 3

---
title: "statistical-analysis"
format: html
editor: visual
---

## Quarto

```{r}
#Path to data. 
data_location <- here::here("data","processed-data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

```{r}
# Load necessary libraries
library(dplyr)

# Define the dependent variable
dependent_var <- "Current_health_expenditure_per_capita_current_US"

# Select 50 variables that are likely to predict the dependent variable
selected_vars <- c(
  # Economic Factors
  "GDP_per_capita_current_US", "GNI_per_capita_Atlas_method_current_US", "GDP_growth_annual_",
  "Inflation_GDP_deflator_annual_", "Domestic_credit_provided_by_financial_sector__of_GDP",
  "Gross_capital_formation__of_GDP", "Exports_of_goods_and_services__of_GDP",
  "Imports_of_goods_and_services__of_GDP", "Foreign_direct_investment_net_inflows_BoP_current_US",
  "Industry_including_construction_value_added__of_GDP", "Agriculture_forestry_and_fishing_value_added__of_GDP",
  "Services_value_added__of_GDP", "Unemployment_total__of_total_labor_force_modeled_ILO_estimate",
  "Income_share_held_by_lowest_20", "Income_share_held_by_highest_20", "Poverty_headcount_ratio_at_215_a_day_2017_PPP__of_population",

  # Health Factors
  "Life_expectancy_at_birth_total_years", "Mortality_rate_under5_per_1000_live_births",
  "Prevalence_of_HIV_total__of_population_ages_1549", "Immunization_measles__of_children_ages_1223_months",
  "Births_attended_by_skilled_health_staff__of_total", "Contraceptive_prevalence_any_method__of_married_women_ages_1549",
  "Hospital_beds_per_1000_people", "Physicians_per_1000_people", "Prevalence_of_underweight_weight_for_age__of_children_under_5",
  "Adolescent_fertility_rate_births_per_1000_women_ages_1519", "Current_health_expenditure__of_GDP",
  "Domestic_general_government_health_expenditure__of_GDP",

  # Social Factors
  "Population_total", "Population_growth_annual_", "Population_density_people_per_sq_km_of_land_area",
  "Urban_population__of_total_population", "Rural_population__of_total_population",
  "Literacy_rate_adult_total__of_people_ages_15_and_above", "School_enrollment_primary__gross",
  "School_enrollment_secondary__gross", "School_enrollment_tertiary__gross", "Unemployment_youth_total__of_total_labor_force_ages_1524_modeled_ILO_estimate",
  "Net_migration", "Refugee_population_by_country_or_territory_of_asylum",

  # Government Factors
  "Military_expenditure__of_GDP", "Tax_revenue__of_GDP", "Government_expenditure_on_education_total__of_GDP",
  "Government_Effectiveness_Estimate", "Rule_of_Law_Estimate", "Control_of_Corruption_Estimate",
  "Political_Stability_and_Absence_of_ViolenceTerrorism_Estimate", "Regulatory_Quality_Estimate",

  # Trade Factors
  "Trade__of_GDP", "Merchandise_trade__of_GDP", "Exports_of_goods_and_services_current_US",
  "Imports_of_goods_and_services_current_US", "Foreign_direct_investment_net_inflows_BoP_current_US",
  "External_debt_stocks_total_DOD_current_US", "Net_barter_terms_of_trade_index_2015__100"
)

# Add the dependent variable to the selected variables
selected_vars <- c(dependent_var, selected_vars)

# Subset the dataset with the selected variables
mydata_selected <- mydata %>%
  select(all_of(selected_vars))

# Categorize the selected variables
variable_categories <- list(
  Economic = c("GDP_per_capita_current_US", "GNI_per_capita_Atlas_method_current_US", "GDP_growth_annual_",
               "Inflation_GDP_deflator_annual_", "Domestic_credit_provided_by_financial_sector__of_GDP",
               "Gross_capital_formation__of_GDP", "Exports_of_goods_and_services__of_GDP",
               "Imports_of_goods_and_services__of_GDP", "Foreign_direct_investment_net_inflows_BoP_current_US",
               "Industry_including_construction_value_added__of_GDP", "Agriculture_forestry_and_fishing_value_added__of_GDP",
               "Services_value_added__of_GDP", "Unemployment_total__of_total_labor_force_modeled_ILO_estimate",
               "Income_share_held_by_lowest_20", "Income_share_held_by_highest_20", "Poverty_headcount_ratio_at_215_a_day_2017_PPP__of_population"),

  Health = c("Life_expectancy_at_birth_total_years", "Mortality_rate_under5_per_1000_live_births",
             "Prevalence_of_HIV_total__of_population_ages_1549", "Immunization_measles__of_children_ages_1223_months",
             "Births_attended_by_skilled_health_staff__of_total", "Contraceptive_prevalence_any_method__of_married_women_ages_1549",
             "Hospital_beds_per_1000_people", "Physicians_per_1000_people", "Prevalence_of_underweight_weight_for_age__of_children_under_5",
             "Adolescent_fertility_rate_births_per_1000_women_ages_1519", "Current_health_expenditure__of_GDP",
             "Domestic_general_government_health_expenditure__of_GDP"),

  Social = c("Population_total", "Population_growth_annual_", "Population_density_people_per_sq_km_of_land_area",
             "Urban_population__of_total_population", "Rural_population__of_total_population",
             "Literacy_rate_adult_total__of_people_ages_15_and_above", "School_enrollment_primary__gross",
             "School_enrollment_secondary__gross", "School_enrollment_tertiary__gross", "Unemployment_youth_total__of_total_labor_force_ages_1524_modeled_ILO_estimate",
             "Net_migration", "Refugee_population_by_country_or_territory_of_asylum"),

  Government = c("Military_expenditure__of_GDP", "Tax_revenue__of_GDP", "Government_expenditure_on_education_total__of_GDP",
                 "Government_Effectiveness_Estimate", "Rule_of_Law_Estimate", "Control_of_Corruption_Estimate",
                 "Political_Stability_and_Absence_of_ViolenceTerrorism_Estimate", "Regulatory_Quality_Estimate"),

  Trade = c("Trade__of_GDP", "Merchandise_trade__of_GDP", "Exports_of_goods_and_services_current_US",
            "Imports_of_goods_and_services_current_US", "Foreign_direct_investment_net_inflows_BoP_current_US",
            "External_debt_stocks_total_DOD_current_US", "Net_barter_terms_of_trade_index_2015__100")
)

# Print the categorized variables
print(variable_categories)
```

```{r}
# Load necessary libraries
library(dplyr)
library(tidyr)
library(broom)      # For tidy regression results
library(purrr)      # For functional programming
library(kableExtra) # For advanced table formatting

# Ensure all selected variables are numeric
mydata <- mydata %>%
  mutate(across(everything(), as.numeric))

# Function to run multiple regression for a specific category and country
run_regression <- function(data, category_vars, category_name) {
  # Drop variables with all NA values within the category
  category_vars <- category_vars[sapply(data[category_vars], function(x) !all(is.na(x)))]

  # If no variables are left after dropping, return NULL
  if (length(category_vars) == 0) {
    return(NULL)
  }

  # Create a subset of data with the dependent variable and independent variables
  regression_data <- data %>%
    select(Current_health_expenditure_per_capita_current_US, all_of(category_vars)) %>%
    na.omit()  # Remove rows with NA values

  # Check if there are enough rows to run the regression
  if (nrow(regression_data) == 0) {
    return(NULL)
  }

  # Create the regression formula
  formula <- as.formula(paste("Current_health_expenditure_per_capita_current_US ~", 
                              paste(category_vars, collapse = " + ")))

  # Run the regression
  regression_model <- lm(formula, data = regression_data)

  # Tidy up the regression results and add a "Significant" column
  tidy_results <- tidy(regression_model) %>%
    mutate(
      Category = category_name,
      Significant = ifelse(p.value <= 0.10, "YES", "NO")  # Add significance column
    )

  return(tidy_results)
}

# Function to run regression for all countries and a specific category
run_regression_all_countries <- function(category_vars, category_name) {
  results <- mydata %>%
    group_by(CountryName) %>%
    group_modify(~ {
      regression_results <- run_regression(.x, category_vars, category_name)
      if (!is.null(regression_results)) {
        return(regression_results)
      } else {
        return(tibble())  # Return empty tibble if no results
      }
    })

  return(results)
}

# Run regression for each category and create tables
economic_results <- run_regression_all_countries(variable_categories$Economic, "Economic")
health_results <- run_regression_all_countries(variable_categories$Health, "Health")
social_results <- run_regression_all_countries(variable_categories$Social, "Social")
government_results <- run_regression_all_countries(variable_categories$Government, "Government")
trade_results <- run_regression_all_countries(variable_categories$Trade, "Trade")

# Combine all results into one table
all_results <- bind_rows(
  economic_results,
  health_results,
  social_results,
  government_results,
  trade_results
)

# Ensure the "Significant" column has no NA values
all_results <- all_results %>%
  mutate(Significant = ifelse(is.na(Significant), "NO", Significant))  # Replace NA with "NO"

# Remove the CountryName column
all_results <- all_results %>%
  select(-CountryName)

# Rename the "term" column to "Predictor variables"
all_results <- all_results %>%
  rename(`Predictor variables` = term)

# Add category separators
all_results <- all_results %>%
  group_by(Category) %>%
  group_modify(~ {
    # Add a separator row before each category
    separator_row <- tibble(
      `Predictor variables` = paste("Independent variables from", .y$Category, "category"),
      estimate = NA, std.error = NA, statistic = NA, p.value = NA, Significant = NA
    )
    bind_rows(separator_row, .x)
  }) %>%
  ungroup()

# Create an advanced table using kableExtra
all_results %>%
  kable("html", caption = "Regression Results by Category") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(width = "100%", height = "500px") %>%
  row_spec(which(all_results$Significant == "YES"), background = "#D9F2D9") %>%  # Highlight significant rows in green
  row_spec(which(all_results$`Predictor variables` %in% paste("Independent variables from", unique(all_results$Category), "category")), 
           bold = TRUE, background = "#FFA500") %>%  # Highlight separator rows in orange
  column_spec(1, bold = TRUE)  # Bold the "Predictor variables" column
```

### 2025-03-12 code/processing-code/processingfile-v1.qmd
project part 3

@@
 library(tidyr)
+library(writexl)
 ```
@@
 ```
+
+Saving an excel file as well.
+
+```{r}
+xl_path <- here("data", "processed-data", "processeddata.xlsx")
+writexl::write_xlsx(processeddata, xl_path)
+```

### 2025-03-12 code/analysis-code/statistical-analysis.R

@@ -1,56 +1,0 @@
-###############################
-# analysis script
-#
-#this script loads the processed, cleaned data, does a simple analysis
-#and saves the results to the results folder
-
-#load needed packages. make sure they are installed.
-library(ggplot2) #for plotting
-library(broom) #for cleaning up output from lm()
-library(here) #for data loading/saving
-
-#path to data
-#note the use of the here() package and not absolute paths
-data_location <- here::here("data","processed-data","processeddata.rds")
-
-#load data. 
-mydata <- readRDS(data_location)
-
-
-######################################
-#Data fitting/statistical analysis
-######################################
-
-############################
-#### First model fit
-# fit linear model using height as outcome, weight as predictor
-
-lmfit1 <- lm(Height ~ Weight, mydata)  
-
-# place results from fit into a data frame with the tidy function
-lmtable1 <- broom::tidy(lmfit1)
-
-#look at fit results
-print(lmtable1)
-
-# save fit results table  
-table_file1 = here("results", "tables", "resulttable1.rds")
-saveRDS(lmtable1, file = table_file1)
-
-############################
-#### Second model fit
-# fit linear model using height as outcome, weight and gender as predictor
-
-lmfit2 <- lm(Height ~ Weight + Gender, mydata)  
-
-# place results from fit into a data frame with the tidy function
-lmtable2 <- broom::tidy(lmfit2)
-
-#look at fit results
-print(lmtable2)
-
-# save fit results table  
-table_file2 = here("results", "tables", "resulttable2.rds")
-saveRDS(lmtable2, file = table_file2)
-
-  

### 2025-03-14 code/analysis-code/statistical-analysis.qmd
project part3

@@
 
-## Quarto
+```{r}
+# Load necessary libraries
+library(dplyr)
+library(tidyr)
+library(broom)      
+library(kableExtra) 
+library(webshot)
+library(magick)
+library(here)
+library(modelsummary)
+library(plm)
+library(lmtest)
+```
+
+At first we will import the dataset
 
@@
 
+We will now pick the most useful variables from the large pool of 1400+ variables in our dataset, and we will then categorize these variables based on the aspects they cover.
+For example, if there is avariable named "Annual GDP growth", we will categorize it as an "Economic" variable.
+Similarly, if there is a variable named "Unemployment rate" or "Total population", we will categorize them as "Social" variable.
+
 ```{r}
-# Load necessary libraries
-library(dplyr)
 
@@
 
+Now that we have selected our variables, we will try to look at the relation between our dependent variable "Health Expenditure Per Capita \$USD" with other variables from a specific category.
+
 ```{r}
-# Load necessary libraries
-library(dplyr)
-library(tidyr)
-library(broom)      # For tidy regression results
-library(purrr)      # For functional programming
-library(kableExtra) # For advanced table formatting
 
-# Ensure all selected variables are numeric
+# Convert all columns to numeric except CountryName, CountryCode, and SeriesCode
 mydata <- mydata %>%
-  mutate(across(everything(), as.numeric))
+  mutate(across(-c(CountryName, CountryCode, SeriesCode), as.numeric))
 
-# Function to run multiple regression for a specific category and country
+# Function to run regression for a specific category of independent variables
 run_regression <- function(data, category_vars, category_name) {
@@
 
-# Function to run regression for all countries and a specific category
-run_regression_all_countries <- function(category_vars, category_name) {
-  results <- mydata %>%
-    group_by(CountryName) %>%
-    group_modify(~ {
-      regression_results <- run_regression(.x, category_vars, category_name)
-      if (!is.null(regression_results)) {
-        return(regression_results)
-      } else {
-        return(tibble())  # Return empty tibble if no results
-      }
-    })
-
-  return(results)
-}
-
 # Run regression for each category and create tables
-economic_results <- run_regression_all_countries(variable_categories$Economic, "Economic")
-health_results <- run_regression_all_countries(variable_categories$Health, "Health")
-social_results <- run_regression_all_countries(variable_categories$Social, "Social")
-government_results <- run_regression_all_countries(variable_categories$Government, "Government")
-trade_results <- run_regression_all_countries(variable_categories$Trade, "Trade")
+economic_results <- run_regression(mydata, variable_categories$Economic, "Economic")
+health_results <- run_regression(mydata, variable_categories$Health, "Health")
+social_results <- run_regression(mydata, variable_categories$Social, "Social")
+government_results <- run_regression(mydata, variable_categories$Government, "Government")
+trade_results <- run_regression(mydata, variable_categories$Trade, "Trade")
 
@@
   mutate(Significant = ifelse(is.na(Significant), "NO", Significant))  # Replace NA with "NO"
-
-# Remove the CountryName column
-all_results <- all_results %>%
-  select(-CountryName)
 
@@
 # Create an advanced table using kableExtra
-all_results %>%
+table_html <- all_results %>%
   kable("html", caption = "Regression Results by Category") %>%
@@
   column_spec(1, bold = TRUE)  # Bold the "Predictor variables" column
+
+# Display the table
+table_html
 ```
+
+In the above table, we can observe that a lot of factors affect the dependent variable and a lot of them don't.
+We can also observe that a lot of our variables have missing values make it unavailable for the regression estimation.
+
+Now, we will save our regression result is a HTML table so that we can use to present it in our manuscript.
+
+```{r}
+# Save the table as a PDF file
+html_file <- here("results", "tables", "regression_results.html")
+save_kable(table_html, file = html_file)
+
+```
+
+Now that we have tested some variables in our model across different categories, we will pick the most useful variable and add them into our preliminary model.
+We will also add other variable that were not present in the above table.
+Variable selection is a continuous process and the variables selected in the preliminary model may also be dropped later to make space for new variables that may appear as important.
+
+```{r}
+
+model <- lm(Current_health_expenditure_per_capita_current_US ~ 
+              GDP_growth_annual_ + 
+              Income_share_held_by_lowest_20 + 
+              Control_of_Corruption_Estimate + 
+              Life_expectancy_at_birth_total_years + 
+Unemployment_youth_total__of_total_labor_force_ages_1524_modeled_ILO_estimate + 
+              Literacy_rate_adult_total__of_people_ages_15_and_above + 
+              Trade__of_GDP+ Foreign_direct_investment_net_inflows_BoP_current_US+ 
+  Access_to_electricity_rural__of_rural_population +
+  Age_dependency_ratio__of_workingage_population
+  , 
+            data = mydata)
+
+# Save the regression results to an HTML file
+modelsummary(
+  model,
+  output = here("results", "tables", "preliminary_regression_results.html"),
+  title = "First Attempt at Regression Model",
+  stars = c('*' = 0.1, '**' = 0.05, '***' = 0.01),
+  notes = list(
+    "Standard errors are shown in parentheses.",
+    "Significance levels: * p < 0.1, ** p < 0.05, *** p < 0.01."
+  ),
+  encoding = "UTF-8"  # Ensure UTF-8 encoding
+)
+```
+
+The variables have a significant impact on the dependent variable as expected.
+But, as this is a time series dataset, we must also test for autocorrelation.
+There are many other tests as well which can be performed on time-series models such ADF test for unit root, and complex models such as ARCH and GARCH can also be used.
+However, we will only perform simple procedures and try to build a simple model.
+
+# Autocorrelation test
+
+Below, we will perform the autocorrelation test.
+
+```{r}
+# Perform Durbin-Watson test
+dw_test <- dwtest(model)
+print(dw_test)
+```
+
+Unfortunately, the Durbin-Watson test is significant, and our model shows presence of autocorrelation.
+
+We will try to use a lagged model to solve the issue of autocorrelation.
+
+```{r}
+# Create a lagged version of the dependent variable
+mydata$Current_health_expenditure_per_capita_current_US_lag1 <- lag(mydata$Current_health_expenditure_per_capita_current_US, 1)
+
+mydata$GDP_growth_annual_lag1 <- lag(mydata$GDP_growth_annual_, 1)
+
+# Refit the model with lagged variables
+model_lagged <- lm(Current_health_expenditure_per_capita_current_US ~ 
+                     GDP_growth_annual_ + GDP_growth_annual_lag1 +
+                     Income_share_held_by_lowest_20 + 
+                     Control_of_Corruption_Estimate + 
+                     Life_expectancy_at_birth_total_years + 
+                     Unemployment_youth_total__of_total_labor_force_ages_1524_modeled_ILO_estimate + 
+                     Literacy_rate_adult_total__of_people_ages_15_and_above + 
+                     Trade__of_GDP + 
+                     Foreign_direct_investment_net_inflows_BoP_current_US + 
+                     Access_to_electricity_rural__of_rural_population +
+                     Age_dependency_ratio__of_workingage_population +
+                     Current_health_expenditure_per_capita_current_US_lag1,  # Include lagged dependent variable
+                   data = mydata)
+
+# autocorrelation test
+
+dw_test_lagged <- dwtest(model_lagged)
+print(dw_test_lagged)
+```
+
+Unfortunately, the DW test is still significant, and lagged model also shows auto correlation.
+
+We will try to use a country fixed effect to account for the all time-invariant differences between countries (e.g., cultural, institutional, or geographical factors).
+Although these country fixed effect is not a solution to the autocorrelation problem in itself, we believe that the new model specification will show a different property than our previous two models with ordinary pooled OLS and lagged model.
+
+```{r}
+# Fit a fixed effects model using lm() with dummy variables
+fixed_model_lm <- lm(Current_health_expenditure_per_capita_current_US ~ 
+                       GDP_growth_annual_ + 
+                       Income_share_held_by_lowest_20 + 
+                       Control_of_Corruption_Estimate + 
+                       Life_expectancy_at_birth_total_years + 
+   Unemployment_youth_total__of_total_labor_force_ages_1524_modeled_ILO_estimate +
+                       Literacy_rate_adult_total__of_people_ages_15_and_above + 
+                       Trade__of_GDP + 
+                       Foreign_direct_investment_net_inflows_BoP_current_US + 
+                       Access_to_electricity_rural__of_rural_population +
+                       Age_dependency_ratio__of_workingage_population + 
+                       factor(CountryName),  # Include dummy variables for each entity
+                     data = mydata)
+
+#autocorrelation test
+
+dw_test_lagged <- dwtest(fixed_model_lm)
+print(dw_test_lagged)
+```
+
+Thankfully, our country fixed effect model no longer shows autocorrelation. We will now be able to build upon this model for our further analysis.

### 2025-03-14 products/manuscript/Manuscript.qmd
project part3

@@
 This uses HTML as output format.
-[See here](https://quarto.org/docs/output-formats/ms-word.html) for more information.
-You can switch to other formats, like docx or pdf.
-See [the Quarto documentation](https://quarto.org/) for other formats.
 
@@
 
-* Prasanga Paudel$^{1,2,*}$ (ORCID: 0000-0000-1234-5678)
+-   Prasanga Paudel$^{1,2,*}$ (ORCID: 0000-0000-1234-5678)
 
@@
 
-$\land$ Corresponding author: some@email.com
+$\land$ Corresponding author: some\@email.com
 
-$\dagger$ Disclaimer: The opinions expressed in this article are the
-author's own and don't reflect their employer.
+$\dagger$ Disclaimer: The opinions expressed in this article are the author's own and don't reflect their employer.
 
@@
 # Summary/Abstract
-_The data is based on World Development Indicator from World Bank Group.
+
+The data is based on World Development Indicator from World Bank Group.
 This is a comprehensive database that includes information including employment, public and private expenditure in health sector, fertility, immunization, literacy, mortality, and many more.The dataset has a total of 1496 variables and across 11 years from 2010 to 2020.
-The actual source of data has information from 1960 to 2023 but to deal with the problem of missing values in recent years and overcomplication in analysis due to too many year data , we will focus on 10 recent years between 2010 and 2020._
+The actual source of data has information from 1960 to 2023 but to deal with the problem of missing values in recent years and overcomplication in analysis due to too many year data , we will focus on 10 recent years between 2010 and 2020.
 
@@
 ## General Background Information
-_We will try to analyze the most influential factors affecting the expenditure in health sector by different countries.
+
+We will try to analyze the most influential factors affecting the expenditure in health sector by different countries.
 The predictor variables will include factors ranging from education, mortality, population distribution among rural and urban spaces, and so on.
-We would use multiple regression analysis with health sector expenditure as dependent variable and the variable of interests as predictors for this panel data._
+We would use multiple regression analysis with health sector expenditure as dependent variable and the variable of interests as predictors for this panel data.
 
-_We are using data from OECD members for analysis, and there have already been some research which have focused on GDP as predictor for health sector expenditure.
+We are using data from OECD members for analysis, and there have already been some research which have focused on GDP as predictor for health sector expenditure.
 Some research have also focused on the relation between life expectancy and health expenditure.
 However, these research were based on fixed effect and random effect models.
-In our analysis we will try to follow OLS multiple regression model._
+In our analysis we will try to follow OLS multiple regression model.
 
 ## Description of data and data source
-_This is a time series data based on 38 member countries from the Organisation for Economic Co-operation and Development (OECD).
+
+This is a time series data based on 38 member countries from the Organisation for Economic Co-operation and Development (OECD).
 This organization is a unique forum with governments from 37 democracies collaborate to develop policy standards to promote sustainable economic growth.
 However, the data obtained from this dataset is crude and needs to be treated before attempting analysis based on it.
-For example, in its present form, the variable name is a long sentense describing the data it contains.The years are listed as column names and variables are listed as Series._
+For example, in its present form, the variable name is a long sentense describing the data it contains.The years are listed as column names and variables are listed as Series.
 
@@
 
-_The main objective of this research is to:_
+*The main objective of this research is to:*
 
-_1. To analyze the factors that affect the expenditure in health sector among OECD countries._
+*1. To analyze the factors that affect the expenditure in health sector among OECD countries.*
 
-_2. To analyze the magnitude of the influence of these factors on health expenditure_
-
-To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above and have the right bibtex key.
-Then you can include like this:
-
-Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].
+*2. To analyze the magnitude of the influence of these factors on health expenditure*
 
@@
 ## Data aquisition
-_The data for this research is available at the official World Bank- Data Bank website.It is a free database managed by world bank group and can be accessed anytime.
-We can apply various filters as per our need and export the data as a .CSV or an .xlsx file._
+
+The data for this research is available at the official World Bank- Data Bank website.It is a free database managed by world bank group and can be accessed anytime.
+We can apply various filters as per our need and export the data as a .CSV or an .xlsx file.
 
@@
 
-_This was the state of the data when first downloaded_
+This was the state of the data when first downloaded
+
 ```{r}
@@
 
-_This is the present state of the dataset_
+This is the present state of the dataset
 
@@
 
-_Here, we will now use a combination of tables and figues to have a look into the data._
+*Here, we will now use a combination of tables and figues to have a look into the data.*
 
@@
 table_html
-
 ```
@@
 
+But, before moving on to be fixed at a single variable as our dependent variable, we will also look if there are any alternative dependent varibles that could prove to be more useful.
+
 @fig-health-exp-aus shows the distribution of "Health Expenditure as a Percentage of GDP changes over time for Australia.
+Here we can see that there is not much variation in the data.
+Therefore we will also look at other similar variables.
 
@@
 #| echo: FALSE 
+#| out.width: "60%"
 knitr::include_graphics(here("results", "figures", "health_exp_aus.png"))
@@
 
-@fig-pop-growth-aus shows the distribution  of Health Expenditure as a Percentage of GDP over the Years Comparing with shows any reation with annual population growth for Australia.This figure shows that it has very less in common with @fig-health-exp-aus as long as distribution and correlation is concerned.
-
-```{r}
-#| label: fig-pop-growth-aus
-#| fig-cap: "Health Expenditure (% of GDP), Country: Australia"
-#| echo: FALSE
-knitr::include_graphics(here("results","figures","pop_growth_aus.png"))
-```
-
 @fig-he_per_capita shows the distribution of Health Expenditure Per Capita (USD) over the Years.
-It shows that it is some what related to @fig-pop-growth-aus in it's distribution.
-So we can say that there is indeed a presence of some pattern in the variables of interest.
+It can be observed that this variable indeed has variation and could be useful as our dependent variable.
 
@@
 #| echo: FALSE
+#| out.width: "60%"
 knitr::include_graphics(here("results","figures","health_exp_per_capita_aus.png"))
 ```
- I think it will be better for us to use the variable "Health Expenditure Per Capita $USD" as a dependent variable in our analysis as it show more variation and connections with other variables.
 
-## Basic statistical analysis
+@fig-pop-growth-aus shows the distribution of population growth of australia.This figure shows that it has very less in common with @fig-health-exp-aus and more in common with @fig-he_per_capita
 
-_To be added_
+```{r}
+#| label: fig-pop-growth-aus
+#| fig-cap: "Population Growth, Country: Australia"
+#| echo: FALSE
+#| out.width: "60%"
+knitr::include_graphics(here("results","figures","pop_growth_aus.png"))
+```
+
+I think it will be better for us to use the variable "Health Expenditure Per Capita \$USD" as a dependent variable in our analysis as it show more variation and connections with other variables.
+
+# Basic statistical analysis
+
+## Preparing a model
+
+To help finalize our model, we will first try to look at the relation between our dependent variable "Health Expenditure Per Capita \$USD" with other variables from a specific category.
+Specifically, we will pick the most useful variables from the large pool of 1400+ variables in our dataset, and we will then categorize these variables based on the aspects they cover.
+For example, if there is avariable named "Annual GDP growth", we will categorize it as an "Economic" variable.
+Similarly, if there is a variable named "Unemployment rate" or "Total population", we will categorize them as "Social" variable
+
+We will then run our initial regression with "Health Expenditure Per Capita \$USD" as dependent variable, and variables from each of the category as independent variables in separate regressions
+
+@fig-regression-results shows the different regresion results based on different categories.
+The categories are: Economic, Health, Social, Government, Trade.
+We will pick the most useful variables among the variables that came up significant and create our preliminary model for our research.
+However, this would not be the end of our variable selection process.
+
+```{r}
+#| label: fig-regression-results
+#| fig-cap: "Regression Results by Category"
+#| echo: FALSE
+htmltools::tags$iframe(
+  src = here("results", "tables", "regression_results.html"),
+  width = "100%",
+  height = "600px"
+)
+```
+
+## Observing the relationship with dependent variable
+
+@fig-correlation-p1 show the correlation of different variables that had a statistical significant effect in @fig-regression-results (and other variables of interest).
+We are just trying to see the pattern present in the dataset for the year 2010, with all the countries treated as cross-sectional units.
+
+```{r}
+#| label: fig-correlation-p1
+#| fig-cap: "Correlation based on 2010 data"
+#| echo: FALSE
+#| out.width: "100%"
+knitr::include_graphics(here("results","figures","correlation_plot1.png"))
+```
+
+We do the same with few other variables.
+
+```{r}
+#| label: fig-correlation-p2
+#| fig-cap: "Correlation based on 2010 data"
+#| echo: FALSE
+#| out.width: "100%"
+knitr::include_graphics(here("results","figures","correlation_plot2.png"))
+```
+
+It can be seen that most of the variables seem to have a high correlation with our variable of interest.
+Although some variables lack observations, we must recall that this is just for the year 2010, and the whole dataset might have a different distribution pattern.
+
+## Proposed Model for the research
+
+Based on the informations obtained in @fig-regression-results, we have selected variable that would best explain our dependent variable. The model is as follow:
+
+Current Health Expenditure Per Capita (\$USD) = $b_0$ + $b_1$ Annual GDP growth + $b_2$ Income share held by lowest 20% + $b_3$ Control of corruption + $b_4$ Life expectancy at birth + $b_5$ Unemployed youth population as a percentage of total labor force + $b_6$ Adult literacy rate + $b_7$ Trade as a percentage of GDP + $b_8$ FDI net inflow + $b_9$ Access to electricity (rural) + $b_{10}$ Age dependency ratio
+
+```{r}
+#| label: fig-preliminary-regression-results
+#| fig-cap: "Preliminary Regression Results"
+#| echo: FALSE
+#| out.width: "70%"
+knitr::include_graphics(here("results", "tables", "preliminary_regression_results.png"))
+
+```
+
+But, it turns out that the model shows significant amount of autocorrelation
+
+```{r}
+#| label: fig-DW-test-results
+#| fig-cap: "Durbin Watson Test Results"
+#| echo: FALSE
+#| out.width: "25%"
+knitr::include_graphics(here("results", "figures", "durbin_watson_test_results.png"))
+
+```
+
+**Therefore, we will use a regression model with country fixed effect which is given below:**
+
+Current Health Expenditure Per Capita (\$USD) = $b_0$ + $b_1$ Annual GDP growth + $b_2$ Income share held by lowest 20% + $b_3$ Control of corruption + $b_4$ Life expectancy at birth + $b_5$ Unemployed youth population as a percentage of total labor force + $b_6$ Adult literacy rate + $b_7$ Trade as a percentage of GDP + $b_8$ FDI net inflow + $b_9$ Access to electricity (rural) + $b_{10}$ Age dependency ratio + country fixed effect.
+
+```{r}
+#| label: fig-DW-test-results-fe
+#| fig-cap: "Durbin Watson Test Results FE-Model"
+#| echo: FALSE
+#| out.width: "30%"
+knitr::include_graphics(here("results", "figures", "durbin_watson_test_results_fe.png"))
+
+```
 
@@
 
-_To be added._
+*To be added.*
 
@@
 ## Summary and Interpretation
-_To be added._
+
+*To be added.*
 
 ## Strengths and Limitations
-_To be added._
+
+*To be added.*
 
 ## Conclusions
-_To be added._
 
-_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_
-
-This paper [@leek2015] discusses types of analyses.
-
-These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.
-
-Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header.
-Many more style files for almost any journal [are available](https://www.zotero.org/styles).
-You also specify the location of your bibtex reference file in the YAML.
-You can call your reference file anything you like.
+*To be added.*
 

### 2025-03-14 code/analysis-code/statistical-analysis.qmd
project #3

@@
 library(lmtest)
+library(gt)
+library(webshot2)
+
 ```
@@
 
-# Save the regression results to an HTML file
 modelsummary(
   model,
-  output = here("results", "tables", "preliminary_regression_results.html"),
-  title = "First Attempt at Regression Model",
+  output = here("results", "tables", "preliminary_regression_results.png"),
+  title = "This table represents the preliminary regresssion model",
   stars = c('*' = 0.1, '**' = 0.05, '***' = 0.01),
@@
     "Significance levels: * p < 0.1, ** p < 0.05, *** p < 0.01."
-  ),
-  encoding = "UTF-8"  # Ensure UTF-8 encoding
   )
+)
+
 ```
@@
 ```{r}
-# Perform Durbin-Watson test
+
+# Perform the Durbin-Watson test
 dw_test <- dwtest(model)
-print(dw_test)
+
+# Extract the test statistic and p-value
+dw_statistic <- dw_test$statistic
+dw_p_value <- dw_test$p.value
+
+# Create a data frame with the results
+dw_results <- data.frame(
+  Statistic = dw_statistic,
+  P_Value = dw_p_value
+)
+
+# Generate a table using gt
+dw_table <- gt(dw_results) %>%
+  tab_header(
+    title = "Durbin-Watson Test Results"
+  ) %>%
+  fmt_number(
+    columns = vars(Statistic, P_Value),
+    decimals = 4
+  )
+
+# Define the file path for the PNG
+png_file <- here::here("results", "figures", "durbin_watson_test_results.png")
+
+# Save the table as a PNG file
+gtsave(dw_table, png_file)
+
 ```
@@
 
-dw_test_lagged <- dwtest(fixed_model_lm)
-print(dw_test_lagged)
+dw_test_fe <- dwtest(fixed_model_lm)
+print(dw_test_fe)
+
+# Extract the test statistic and p-value
+dw_lstatistic <- dw_test_fe$statistic
+dw_lp_value <- dw_test_fe$p.value
+
+# Create a data frame with the results
+dw_lresults <- data.frame(
+  Statistic = dw_lstatistic,
+  P_Value = dw_lp_value
+)
+
+# Generate a table using gt
+dw_table <- gt(dw_lresults) %>%
+  tab_header(
+    title = "Durbin-Watson Test Results for FE "
+  ) %>%
+  fmt_number(
+    columns = vars(Statistic, P_Value),
+    decimals = 4
+  )
+
+# Define the file path for the PNG
+png_file <- here::here("results", "figures", "durbin_watson_test_results_fe.png")
+
+# Save the table as a PNG file
+gtsave(dw_table, png_file)
+
 ```

### 2025-03-14 code/eda-code/eda.qmd
project #3

@@
 library(webshot2)
+library(patchwork)
 ```
@@
 
+# EDA (part2)
+
+This is the EDA analysis performed for part 3 of the project.
+
+```{r}
+# Define dataset
+df <- data_2010
+
+# Function to create a scatter plot with correlation
+create_corr_plot <- function(x_var, df, point_color) {
+  cor_value <- round(cor(df[[x_var]], df$Current_health_expenditure_per_capita_current_US, use = "complete.obs"), 2)
+
+  ggplot(df, aes_string(x = x_var, y = "Current_health_expenditure_per_capita_current_US")) +
+    geom_point(color = point_color, alpha = 0.6) +  # Set color based on input
+    geom_smooth(method = "lm", color = "red", se = FALSE) +
+    labs(title = paste("Correlation:", cor_value), x = x_var, y = "Health Expenditure per Capita") +
+    theme_minimal()
+}
+
+# Generate individual plots with different colors
+plot1 <- create_corr_plot("GDP_growth_annual_", df, "blue")
+plot2 <- create_corr_plot("Income_share_held_by_lowest_20", df, "blue")
+plot3 <- create_corr_plot("Control_of_Corruption_Estimate", df, "blue")
+plot4 <- create_corr_plot("Life_expectancy_at_birth_total_years", df, "blue")
+
+# Arrange plots in a 2x2 grid
+final_plot1 <- (plot1 + plot2) / (plot3 + plot4)
+png_file1 <- here("results", "figures", "correlation_plot1.png")
+ggsave(filename = png_file1, plot = final_plot1, width = 10, height = 8, dpi = 300)
+
+# Generate individual plots with different colors
+plot5 <- create_corr_plot("Literacy_rate_adult_total__of_people_ages_15_and_above", df, "green")
+plot6 <- create_corr_plot("Trade__of_GDP", df, "green")
+plot7 <- create_corr_plot("Foreign_direct_investment_net_inflows_BoP_current_US", df, "green")
+plot8 <- create_corr_plot("Age_dependency_ratio__of_workingage_population", df, "green")
+
+# Arrange plots in a 2x2 grid
+final_plot2 <- (plot5 + plot6) / (plot7 + plot8)
+png_file2 <- here("results", "figures", "correlation_plot2.png")
+ggsave(filename = png_file2, plot = final_plot2, width = 10, height = 8, dpi = 300)
+```
+
 # Notes

### 2025-03-14 products/manuscript/Manuscript.qmd
project #3

@@
 We are using data from OECD members for analysis, and there have already been some research which have focused on GDP as predictor for health sector expenditure.
-Some research have also focused on the relation between life expectancy and health expenditure.
-However, these research were based on fixed effect and random effect models.
+Some research have such as Nkemgha in 2020 [@Nkemgha2020] studied the link among public health expenditure, private health expenditure and life expectancy in Cameroon.
+Christopoulos and Eleftheriou [@CHRISTOPOULOS2020195] using panel data analyzed the fiscal effects of health expenditures on health outcomes and found a significant impact of healthcare expenditures on increasing revenue.
 In our analysis we will try to follow OLS multiple regression model.

### 2025-03-21 code/eda-code/eda.qmd
changed to numeric

@@
 # Define dataset
+
 df <- data_2010
+
+df$Current_health_expenditure_per_capita_current_US <- as.numeric(df$Current_health_expenditure_per_capita_current_US)
+df$GDP_growth_annual_ <- as.numeric(df$GDP_growth_annual_)
+df$Income_share_held_by_lowest_20 <- as.numeric(df$Income_share_held_by_lowest_20)
+df$Control_of_Corruption_Estimate <- as.numeric(df$Control_of_Corruption_Estimate)
+df$Life_expectancy_at_birth_total_years <- as.numeric(df$Life_expectancy_at_birth_total_years)
+df$Literacy_rate_adult_total__of_people_ages_15_and_above <- as.numeric(df$Literacy_rate_adult_total__of_people_ages_15_and_above)
+df$Trade__of_GDP <- as.numeric(df$Trade__of_GDP)
+df$Foreign_direct_investment_net_inflows_BoP_current_US <- as.numeric(df$Foreign_direct_investment_net_inflows_BoP_current_US)
+df$Age_dependency_ratio__of_workingage_population <- as.numeric(df$Age_dependency_ratio__of_workingage_population)
 

### 2025-04-03 code/analysis-code/statistical-analysis.qmd
reg summary
part #4

--- 2025-03-12 code/analysis-code/statistical-analysis.qmd
@@
 
-## Quarto
+We will load the necessary libraries.
+
+```{r}
+# Load necessary libraries
+library(dplyr)
+library(tidyr)
+library(broom)      
+library(kableExtra) 
+library(webshot)
+library(magick)
+library(here)
+library(modelsummary)
+library(plm)
+library(lmtest)
+library(gt)
+library(webshot2)
+library(fixest)
+library(lubridate)
+library(Metrics)
+
+```
+
+Then we will import the dataset
 
@@
 
+We will now pick the most useful variables from the large pool of 1400+ variables in our dataset, and we will then categorize these variables based on the aspects they cover.
+For example, if there is avariable named "Annual GDP growth", we will categorize it as an "Economic" variable.
+Similarly, if there is a variable named "Unemployment rate" or "Total population", we will categorize them as "Social" variable.
+
+But, before that we will import the training data which was obtained after splitting the dataset into 80% train data and 20% test data. We will be performing all our analysis on train data.
+
 ```{r}
-# Load necessary libraries
-library(dplyr)
+train <- readRDS(here("data", "processed-data", "train-data.rds"))
+test <- readRDS(here("data", "processed-data", "test-data.rds"))
+```
 
-# Define the dependent variable
+```{r}
+
+#defining the dependent variable
 dependent_var <- "Current_health_expenditure_per_capita_current_US"
 
-# Select 50 variables that are likely to predict the dependent variable
+#selecting 50 variables that are likely to predict the dependent variable
 selected_vars <- c(
@@
 # Subset the dataset with the selected variables
-mydata_selected <- mydata %>%
+train_selected <- train %>%
   select(all_of(selected_vars))
@@
 
+Now that we have selected our variables, we will try to look at the relation between our dependent variable "Health Expenditure Per Capita \$USD" with other variables from a specific category.
+
 ```{r}
-# Load necessary libraries
-library(dplyr)
-library(tidyr)
-library(broom)      # For tidy regression results
-library(purrr)      # For functional programming
-library(kableExtra) # For advanced table formatting
 
-# Ensure all selected variables are numeric
-mydata <- mydata %>%
-  mutate(across(everything(), as.numeric))
+# converting all columns to numeric except CountryName, CountryCode, and SeriesCode
+train <- train %>%
+  mutate(across(-c(CountryName, CountryCode, SeriesCode), as.numeric))
 
-# Function to run multiple regression for a specific category and country
+#saving the train dataset as we will be using it in future as well
+
+saveRDS(train, file = here("data", "processed-data", "train-data.rds"))
+
+# converting all columns to numeric except CountryName, CountryCode, and SeriesCode
+test <- test %>%
+  mutate(across(-c(CountryName, CountryCode, SeriesCode), as.numeric))
+
+#saving the train dataset as we will be using it in future as well
+
+saveRDS(test, file = here("data", "processed-data", "test-data.rds"))
+```
+
+```{r}
+
+# Function to run regression for a specific category of independent variables
 run_regression <- function(data, category_vars, category_name) {
-  # Drop variables with all NA values within the category
+  # dropping variables with all NA values within the category
   category_vars <- category_vars[sapply(data[category_vars], function(x) !all(is.na(x)))]
@@
 
-  # Create a subset of data with the dependent variable and independent variables
+  # creating a subset of data with the dependent variable and independent variables
   regression_data <- data %>%
@@
 
-  # Check if there are enough rows to run the regression
+  # checking if there are enough rows to run the regression
   if (nrow(regression_data) == 0) {
@@
 
-  # Create the regression formula
+  # creating the regression formula
   formula <- as.formula(paste("Current_health_expenditure_per_capita_current_US ~", 
@@
       Category = category_name,
-      Significant = ifelse(p.value <= 0.10, "YES", "NO")  # Add significance column
+      Significant = ifelse(p.value <= 0.05, "YES", "NO")  # Add significance column
     )
@@
 
-# Function to run regression for all countries and a specific category
-run_regression_all_countries <- function(category_vars, category_name) {
-  results <- mydata %>%
-    group_by(CountryName) %>%
-    group_modify(~ {
-      regression_results <- run_regression(.x, category_vars, category_name)
-      if (!is.null(regression_results)) {
-        return(regression_results)
-      } else {
-        return(tibble())  # Return empty tibble if no results
-      }
-    })
+# Run regression for each category and create tables
+economic_results <- run_regression(train, variable_categories$Economic, "Economic")
+health_results <- run_regression(train, variable_categories$Health, "Health")
+social_results <- run_regression(train, variable_categories$Social, "Social")
+government_results <- run_regression(train, variable_categories$Government, "Government")
+trade_results <- run_regression(train, variable_categories$Trade, "Trade")
 
-  return(results)
-}
-
-# Run regression for each category and create tables
-economic_results <- run_regression_all_countries(variable_categories$Economic, "Economic")
-health_results <- run_regression_all_countries(variable_categories$Health, "Health")
-social_results <- run_regression_all_countries(variable_categories$Social, "Social")
-government_results <- run_regression_all_countries(variable_categories$Government, "Government")
-trade_results <- run_regression_all_countries(variable_categories$Trade, "Trade")
-
-# Combine all results into one table
+# combine all results into one table
 all_results <- bind_rows(
@@
 
-# Ensure the "Significant" column has no NA values
+# ensuring the "Significant" column has no NA values
 all_results <- all_results %>%
   mutate(Significant = ifelse(is.na(Significant), "NO", Significant))  # Replace NA with "NO"
-
-# Remove the CountryName column
-all_results <- all_results %>%
-  select(-CountryName)
 
@@
 
-# Add category separators
+# Adding category separators
 all_results <- all_results %>%
@@
 
-# Create an advanced table using kableExtra
-all_results %>%
+# creating an advanced table using kableExtra
+table_html <- all_results %>%
   kable("html", caption = "Regression Results by Category") %>%
@@
   column_spec(1, bold = TRUE)  # Bold the "Predictor variables" column
+
+# Display the table
+table_html
 ```
+
+In the above table, we can observe that a lot of factors affect the dependent variable and a lot of them don't.
+We can also observe that a lot of our variables have missing values making it unavailable for the regression estimation.
+
+Now, we will save our regression result is a HTML table so that we can use to present it in our manuscript.
+
+```{r}
+# Save the table as a HTML file
+html_file <- here("results", "tables", "regression_results.html")
+save_kable(table_html, file = html_file)
+
+```
+
+Now that we have tested some variables in our model across different categories, we will pick the most useful variable and add them into our preliminary model.
+We will also add other variable that were not present in the above table.
+Variable selection is a continuous process and the variables selected in the preliminary model may also be dropped later to make space for new variables that may appear as important.
+
+```{r}
+# preliminary regression model
+
+model <- lm(Current_health_expenditure_per_capita_current_US ~ 
+              GDP_growth_annual_ + 
+              Income_share_held_by_lowest_20 + 
+              Control_of_Corruption_Estimate + 
+              Life_expectancy_at_birth_total_years + 
+Unemployment_youth_total__of_total_labor_force_ages_1524_modeled_ILO_estimate + 
+              Trade__of_GDP+ Foreign_direct_investment_net_inflows_BoP_current_US+ 
+  Access_to_electricity_rural__of_rural_population +
+  Age_dependency_ratio__of_workingage_population
+  , 
+            data = train)
+
+modelsummary(
+  model,
+  output = here("results", "tables", "preliminary_regression_results.png"),
+  title = "This table represents the preliminary regresssion model",
+  stars = c('*' = 0.1, '**' = 0.05, '***' = 0.01),
+  notes = list(
+    "Standard errors are shown in parentheses.",
+    "Significance levels: * p < 0.1, ** p < 0.05, *** p < 0.01."
+  )
+)
+
+```
+
+The variables have a significant impact on the dependent variable as expected.
+But, as this is a time series dataset, we must also test for autocorrelation.
+There are many other tests as well which can be performed on time-series models such ADF test for unit root, and complex models for variance modelling such as ARCH and GARCH can also be used.
+However, we will only perform simple procedures and try to build a simple model.
+
+# Autocorrelation test
+
+Below, we will perform the autocorrelation test.
+
+```{r}
+
+# Perform the Durbin-Watson test
+dw_test <- dwtest(model)
+
+# Extract the test statistic and p-value
+dw_statistic <- dw_test$statistic
+dw_p_value <- dw_test$p.value
+
+# Create a data frame with the results
+dw_results <- data.frame(
+  Statistic = dw_statistic,
+  P_Value = dw_p_value
+)
+
+# Generate a table using gt
+dw_table <- gt(dw_results) %>%
+  tab_header(
+    title = "Durbin-Watson Test Results"
+  ) %>%
+  fmt_number(
+    columns = vars(Statistic, P_Value),
+    decimals = 4
+  )
+
+# Define the file path for the PNG
+png_file <- here::here("results", "figures", "durbin_watson_test_results.png")
+
+# Save the table as a PNG file
+gtsave(dw_table, png_file)
+
+```
+
+Fortunately, the Durbin-Watson test is insignificant, and our model does not show presence of autocorrelation.
+
+We can therefore move forward with this model.
+
+Calculating the RMSE of a NULL Model.
+
+```{r}
+null_model <- lm(Current_health_expenditure_per_capita_current_US ~ 1, 
+                 data = train)
+
+# Calculate residuals
+residuals <- residuals(null_model)
+
+# Calculate RMSE manually
+null_rmse <- sqrt(mean(residuals^2))
+
+# Display results
+summary(null_model)
+cat("\nNull Model RMSE:", null_rmse, "\n")
+```
+
+ ADF test (combined)
+```{r}
+library(plm)
+
+# 1. Prepare panel data (handle NAs first)
+train_clean <- na.omit(train[, c("CountryName", "Year", "Current_health_expenditure_per_capita_current_US")])
+
+# 2. Check time dimensions
+time_counts <- table(train_clean$country)
+print(time_counts)
+
+# 3. Run Fisher-type test with reduced lags
+pdata <- pdata.frame(train_clean, index = c("CountryName", "Year"))
+fisher_test <- purtest(pdata$Current_health_expenditure_per_capita_current_US,
+                      test = "madwu",
+                      exo = "intercept",
+                      lags = 1) # Force shorter lags
+
+```
+
+```{r}
+library(kableExtra)
+library(here)
+
+# 1. Create properly formatted results table
+unit_root_table <- data.frame(
+  Test = "Fisher-Type Panel Unit Root Test (ADF)",
+  Statistic = fisher_test$statistic$statistic,
+  DF = fisher_test$statistic$parameter,
+  `p-value` = fisher_test$statistic$p.value,
+  check.names = FALSE
+)
+
+# 2. Create and save HTML table
+unit_root_table %>%
+  kbl(format = "html", digits = 4, align = "c",
+      caption = "Panel Unit Root Test Results") %>%
+  kable_styling(bootstrap_options = c("striped", "hover")) %>%
+  add_footnote("Null hypothesis: All series contain unit roots", 
+              notation = "none") %>%
+  save_kable(here("results", "tables", "unit_root_results.html"))
+```
+
+ADF test (Individual countries)
+
+```{r}
+library(tseries)
+
+# Option 1: Simple version with try()
+adf_results <- by(train, train$CountryName, function(x) {
+  try(adf.test(na.omit(x$Current_health_expenditure_per_capita_current_US)),
+  silent = TRUE)
+})
+
+# Option 2: More robust version with error handling
+adf_results <- lapply(split(train, train$CountryName), function(df) {
+  if (sum(!is.na(df$Current_health_expenditure_per_capita_current_US)) >= 3) {
+    tryCatch({
+      adf.test(na.omit(df$Current_health_expenditure_per_capita_current_US))
+    }, error = function(e) {
+      return(list(statistic = NA, p.value = NA, 
+                 note = paste("Error:", e$message)))
+    })
+  } else {
+    return(list(statistic = NA, p.value = NA, 
+               note = "Not enough observations"))
+  }
+})
+
+# Create a clean results table
+results_table <- do.call(rbind, lapply(names(adf_results), function(CountryName) {
+  res <- adf_results[[CountryName]]
+  data.frame(
+    Country = CountryName,
+    ADF_Statistic = ifelse(is.null(res$statistic), NA, res$statistic),
+    P_Value = ifelse(is.null(res$p.value), NA, res$p.value),
+    Note = ifelse(is.null(res$note), "", res$note)
+  )
+}))
+
+print(results_table)
+```

### 2025-04-03 code/eda-code/eda.qmd
part #4

@@
 
+```{r}
+# sorting the dataset by time
+mydata <- mydata %>% arrange(Year)  # sorting by year
+
+#splitting at 80% of the observations
+split_index <- floor(0.8 * nrow(mydata))
+train <- mydata[1:split_index, ]
+test <- mydata[(split_index + 1):nrow(mydata), ]
+```
+
 # Data exploration through tables
@@
 # Filtering data for the year 2010
-data_2010 <- mydata %>%
+data_2010 <- train %>%
   filter(Year == 2010)
@@
 # Converting column to numeric before plotting
-mydata <- mydata %>%
+train <- train %>%
   mutate(Current_health_expenditure__of_GDP = as.numeric(Current_health_expenditure__of_GDP))
@@
 # Filtering top 5 countries based on 2010 health expenditure
-top_5_countries <- mydata %>%
+top_5_countries <- train %>%
   filter(Year == 2010) %>%
@@
 # Filtering data for only 5 countries over time
-data_filtered <- mydata %>%
+data_filtered <- train %>%
   filter(CountryName %in% top_5_countries)
@@
 # Converting columns to numeric 
-mydata <- mydata %>%
+train <- train %>%
   mutate(
@@
 # Converting columns to numeric
-mydata <- mydata %>%
+train <- train %>%
   mutate(
@@
 # Filtering data for Australia
-australia_data <- mydata %>%
+australia_data <- train %>%
   filter(CountryName == "Australia")
@@
 # Filter data for France
-france_data <- mydata %>%
+france_data <- train %>%
   filter(CountryName == "France")
@@
 
+```{r}
+# Converting column to numeric before plotting
+train <- train %>%
+  mutate(Current_health_expenditure_per_capita_current_US = as.numeric(Current_health_expenditure_per_capita_current_US))
+
+# Filtering top 5 countries based on 2010 per capita health expenditure
+top_5_countries <- train %>%
+  filter(Year == 2010) %>%
+  arrange(desc(Current_health_expenditure_per_capita_current_US)) %>%
+  slice_head(n = 5) %>%
+  pull(CountryName)  # Extracting country names
+
+# Filtering data for only 5 countries over time
+data_filtered <- train %>%
+  filter(CountryName %in% top_5_countries)
+
+y_breaks <- seq(min(data_filtered$Current_health_expenditure_per_capita_current_US, na.rm = TRUE), 
+                max(data_filtered$Current_health_expenditure_per_capita_current_US, na.rm = TRUE), 
+                length.out = 4)
+
+# Creating the line plot
+ggplot(data_filtered, aes(x = Year, y = Current_health_expenditure_per_capita_current_US, color = CountryName, group = CountryName)) +
+  geom_line(size = 1) +
+  geom_point(size = 2) +  
+  labs(
+    title = "Current Per Capita Health Expenditure (2010–2020): Top 5 nations",
+    x = "Year",
+    y = "Health Expenditure (% of GDP)",
+    color = "Country"
+  ) +
+  scale_y_continuous(breaks = y_breaks) +  
+  theme_minimal()
+
+ggsave(
+  filename = here("results","figures", "per_capita_health_expenditure_plot.png"),  
+  plot = last_plot(),  # Saves the last plotted figure
+  width = 8, 
+  height = 6, 
+  dpi = 300
+)
+
+```
+
 # EDA (part2)
@@
 ```
-
-# Notes
-
-For your own explorations, tables and figures can be "quick and dirty".
-As long as you can see what's going on, there is no need to polish them.
-That's in contrast to figures you'll produce for your final products (paper, report, presentation, website, etc.).
-Those should look as nice, polished and easy to understand as possible.

### 2025-04-03 code/robustness-code/robustness-code.qmd
part #4

---
title: "robustness-check"
format: html
editor: visual
---

```{r}
# Load required packages
library(purrr)    # For map_df()
library(dplyr)    # For data manipulation
library(timetk)   # For time_series_cv()
library(rsample)  # For analysis() and assessment()
library(here)
library(modelsummary)
library(tidyr)
library(ranger)

```

Importing train/test dataset.

```{r}
train <- readRDS(here("data", "processed-data", "train-data.rds"))
test <- readRDS(here("data", "processed-data", "test-data.rds"))
```

testing the results

```{r}
# preliminary regression model

model <- lm(Current_health_expenditure_per_capita_current_US ~ 
              GDP_growth_annual_ + 
              Income_share_held_by_lowest_20 + 
              Control_of_Corruption_Estimate + 
              Life_expectancy_at_birth_total_years + 
Unemployment_youth_total__of_total_labor_force_ages_1524_modeled_ILO_estimate + 
              Trade__of_GDP+ Foreign_direct_investment_net_inflows_BoP_current_US+ 
  Access_to_electricity_rural__of_rural_population +
  Age_dependency_ratio__of_workingage_population
  , 
            data = test)

modelsummary(
  model,
  output = here("results", "tables", "testresult.png"),
  title = "This table represents regresssion results from the test data",
  stars = c('*' = 0.1, '**' = 0.05, '***' = 0.01),
  notes = list(
    "Standard errors are shown in parentheses.",
    "Significance levels: * p < 0.1, ** p < 0.05, *** p < 0.01."
  )
)

```

# cross validation

```{r}

# 1. Define your model formula (as you originally specified)
model_formula <- Current_health_expenditure_per_capita_current_US ~ 
  GDP_growth_annual_ + 
  Income_share_held_by_lowest_20 + 
  Control_of_Corruption_Estimate + 
  Life_expectancy_at_birth_total_years + 
  Unemployment_youth_total__of_total_labor_force_ages_1524_modeled_ILO_estimate + 
  Trade__of_GDP + 
  Foreign_direct_investment_net_inflows_BoP_current_US + 
  Access_to_electricity_rural__of_rural_population +
  Age_dependency_ratio__of_workingage_population

# 2. Create time-series CV splits
cv_splits <- time_series_cv(
  data = train,
  date_var = Year,
  initial = 5,  # 5-year initial window
  assess = 1,   # 1-year test window
  skip = 1,     # Move forward 1 year each split
  slice_limit = 3
)
```

```{r}

set.seed(123)

cv <- function(splits, model_formula) {
  map_df(splits, function(split) {
    train_cv <- analysis(split)
    test_cv <- assessment(split)

    # Train model
    model <- lm(model_formula, data = train_cv)

    # Get predictions and actuals
    preds <- predict(model, newdata = test_cv)
    actuals <- test_cv$Current_health_expenditure_per_capita_current_US

    # Convert to numeric safely
    preds_num <- suppressWarnings(as.numeric(preds))
    actuals_num <- suppressWarnings(as.numeric(actuals))

    # Calculate RMSE with error handling
    if(length(preds_num) == 0 || length(actuals_num) == 0) {
      rmse_val <- NA_real_
      status <- "Empty predictions/actuals"
    } else if(any(is.na(actuals_num))) {
      valid <- !is.na(actuals_num)
      rmse_val <- sqrt(mean((preds_num[valid] - actuals_num[valid])^2, na.rm = TRUE))
      status <- "NA in actuals"
    } else {
      rmse_val <- sqrt(mean((preds_num - actuals_num)^2, na.rm = TRUE))
      status <- "Success"
    }

    data.frame(
      Train_Years = paste(min(train_cv$Year), max(train_cv$Year), sep = "-"),
      Test_Year = max(test_cv$Year),
      RMSE = rmse_val,
      stringsAsFactors = FALSE
    )
  })
}

# Run the analysis
results <- cv(cv_splits$splits, model_formula)
print(results)
```

```{r}
# View summary of CV results
cv_summary <- results %>%
  summarise(
    Mean_RMSE = mean(RMSE, na.rm = TRUE),
    SD_RMSE = sd(RMSE, na.rm = TRUE),
  )

print(cv_summary)
```

# Random Forest

```{r}
rf_cv <- function(splits, model_formula) {
  map_df(splits, function(split) {
    train_cv <- analysis(split)
    test_cv <- assessment(split)

    # 1. Get target variable from formula
    target_var <- all.vars(model_formula)[1]

    # 2. Check if Year column exists (case-insensitive)
    year_col <- grep("^year$", names(train_cv), ignore.case = TRUE, value = TRUE)

    # 3. Select required variables (including Year if it exists)
    required_vars <- unique(c(all.vars(model_formula), year_col))
    train_cv <- train_cv %>% 
      select(any_of(required_vars)) %>% 
      drop_na(all_of(target_var))

    test_cv <- test_cv %>% 
      select(any_of(required_vars)) %>% 
      drop_na(all_of(target_var))

    # 4. Get year information
    get_years <- function(df) {
      if (length(year_col) > 0) {
        list(
          min_year = min(df[[year_col]], na.rm = TRUE),
          max_year = max(df[[year_col]], na.rm = TRUE)
        )
      } else {
        list(min_year = NA_real_, max_year = NA_real_)
      }
    }

    train_years <- get_years(train_cv)
    test_year <- get_years(test_cv)$max_year

    # 5. Calculate mtry
    n_predictors <- length(setdiff(required_vars, c(target_var, year_col)))
    mtry_value <- max(1, min(floor(n_predictors / 3), n_predictors))

    # 6. Train model
    set.seed(123)
    rf_model <- ranger(
      formula = model_formula,
      data = train_cv,
      num.trees = 500,
      mtry = mtry_value,
      importance = 'permutation',
      seed = 123
    )

    # 7. Calculate RMSE
    preds <- predict(rf_model, data = test_cv)$predictions
    actuals <- pull(test_cv, target_var)
    rmse_val <- sqrt(mean((preds - actuals)^2, na.rm = TRUE))

    # 8. Return results
    tibble(
      Train_Years = if (!is.na(train_years$min_year)) {
        paste(train_years$min_year, train_years$max_year, sep = "-")
      } else {NA_character_},
      Test_Year = test_year,
      RMSE = rmse_val,
      Model = "Random Forest",
      Predictors_Used = n_predictors,
      mtry_Used = mtry_value,
      Year_Column_Found = ifelse(length(year_col) > 0, year_col, "Not found")
    )
  })
}
```

```{r}
rf_results <- rf_cv(cv_splits$splits, model_formula)
print(rf_results)
```

# Random Forest on test dataset

```{r}
library(tidyverse)
library(ranger)

# Prepare training data (remove NA in target variable)
target_var <- all.vars(model_formula)[1]
train_processed <- train %>%
  drop_na(all_of(target_var))

# Calculate appropriate mtry
n_predictors <- length(all.vars(model_formula)) - 1
mtry_value <- max(1, min(floor(n_predictors / 3), n_predictors))

# Train Random Forest model
set.seed(123)
rf_model <- ranger(
  formula = model_formula,
  data = train_processed,
  num.trees = 500,
  mtry = mtry_value,
  importance = 'permutation',
  seed = 123
)
```

```{r}
# Ensure test data has same structure as training data
test_processed <- test %>%
  select(any_of(names(train_processed))) %>%  # Keep only columns present in training
  drop_na(all_of(target_var))  # Remove rows with NA in target
```

```{r}
# Generate predictions
test_predictions <- predict(rf_model, data = test_processed)$predictions

# Create results dataframe
test_results <- test_processed %>%
  mutate(Predicted = test_predictions,
         Residual = .data[[target_var]] - Predicted)
```

```{r}
# Calculate metrics
performance_metrics <- test_results %>%
  summarise(
    RMSE = sqrt(mean(Residual^2, na.rm = TRUE)),
    MAE = mean(abs(Residual), na.rm = TRUE),
    R2 = 1 - (sum(Residual^2) / sum((.data[[target_var]] - mean(.data[[target_var]]))^2))
  )

# Print metrics
cat("\nTest Set Performance:\n")
print(performance_metrics)

# View predictions vs actual
head(test_results %>% select(all_of(target_var), Predicted, Residual))
```

### 2025-04-03 products/manuscript/Manuscript.qmd
part #4

--- 2025-03-14 products/manuscript/Manuscript.qmd
@@
 bibliography: ../../assets/dataanalysis-references.bib
-csl: ../../assets/american-journal-of-epidemiology.csl
+csl: ../../assets/vancouver-author-date.csl
 ---
 
-This uses HTML as output format.
+This project uses HTML as output format.
 
@@
 
-# Summary/Abstract
+# Abstract
 
@@
 We are using data from OECD members for analysis, and there have already been some research which have focused on GDP as predictor for health sector expenditure.
-Some research have also focused on the relation between life expectancy and health expenditure.
-However, these research were based on fixed effect and random effect models.
+Some research have such as [@Nkemgha2020] studied the link among public health expenditure, private health expenditure and life expectancy in Cameroon.
+[@CHRISTOPOULOS2020195] using panel data analyzed the fiscal effects of health expenditures on health outcomes and found a significant impact of healthcare expenditures on increasing revenue.
 In our analysis we will try to follow OLS multiple regression model.
@@
 This organization is a unique forum with governments from 37 democracies collaborate to develop policy standards to promote sustainable economic growth.
-However, the data obtained from this dataset is crude and needs to be treated before attempting analysis based on it.
-For example, in its present form, the variable name is a long sentense describing the data it contains.The years are listed as column names and variables are listed as Series.
+However, the data obtained from this dataset is crude and needed to be treated before attempting analysis based on it.
+The details about the data cleaning process can be assessed from this github repository.
 
@@
 
-## Data aquisition
-
-The data for this research is available at the official World Bank- Data Bank website.It is a free database managed by world bank group and can be accessed anytime.
-We can apply various filters as per our need and export the data as a .CSV or an .xlsx file.
-
-Link to the source: <https://databank.worldbank.org/source/world-development-indicators/Type/TABLE/preview/on#>
-
 # Methods
 
-We wll be following a panel data time series regression model. The details will be discussed as we progress.
+We will be following a panel data time series regression model. A general form of the OLS model for time series can be represented as:
 
-# Importing the original dataset
+$y_{it}$ = $b_0$ + $b_1$$X_{it1}$ + $b_2$$X_{it2}$ + $b_3$$X_{it3}$ + $b_4$$X_{it4}$ + $b_5$$X_{it5}$ + $e_{it}$
 
-```{r}
-file_path <- here("data","raw-data", "wdi-oecd.xlsx")
-data <- read_excel(file_path, sheet="Data")
-```
+where $e_{it}$ is the error term.
+$X_{itk}$ is the $k^{th}$ independent variable ( k= 1 to 5, in the above equation but our research will likely have k>5) $y_{it}$ is the dependent variable, $b_0$ is the intercept, and $b_k$ is the estimated parameter for the $k^{th}$ independent variable.
 
-## A look into the original format of the dataset
+# Dataset
 
-This was the state of the data when first downloaded
+The original data in it's raw form was obtained from World Bank.
+However, the dataset needed processing, and some modification.
+The final dataset that is eligible for our use contains 462 observations and 1496 variables for the OECD countries.
+However, we will be using 80% of the dataset for our whole analysis and we will test the robustness of the results on the un-used 20% at the end of the analysis.
 
-```{r}
-head(data)
-```
-
-# Importing the cleaned dataset
-
-```{r}
-#Path to data. 
-data_location <- here::here("data","processed-data","processeddata.rds")
-#load data
-mydata <- readRDS(data_location)
-```
-
-## A look into the cleaned dataset
-
-This is the present state of the dataset
-
-```{r}
-head(mydata)
-```
-
-{{< pagebreak >}}
-
-# Results
-
-In exploratory data analysis, we will just look at data from Australia as a reference for relation between variables in most of the cases.
-We are currently thinking of using "Current Health Expenditure as a % of GDP" as our dependent variable.
-But, we will see the it's distribution and connections with other variables before making decision.
-
-## Exploratory/Descriptive analysis
-
-*Here, we will now use a combination of tables and figues to have a look into the data.*
-
-As this is a time series data, it is quite difficult to visualize all the variables across the years for all the countries.
-So, we will try to keep it simple and see simple relations between variables within a year or trend of a particular variable for some countries across the years.
+## A look into the train dataset
 
@@
 
-### Looking into the trend for our possible dependent variable
+{{< pagebreak >}}
+
+# Results
+
+In exploratory data analysis, we will just look at data from Australia as a reference for relation between variables in most of the cases.
+We are currently thinking of using "Current Health Expenditure as a % of GDP" as our dependent variable.
+But, we will see it's distribution and other alternatives before making decision.
+
+## Exploratory/Descriptive analysis
+
+*Here, we will now use a combination of tables and figues to have a look into the data.*
+
+As this is a time series data, it is quite difficult to visualize all the variables across the years for all the countries.
+So, we will try to keep it simple and see simple relations between variables within a year or trend of a particular variable for some countries across the years.
+
+### Looking into the trend for our possible dependent variables
 
@@
 
-@fig-health-exp-aus shows the distribution of "Health Expenditure as a Percentage of GDP changes over time for Australia.
-Here we can see that there is not much variation in the data.
-Therefore we will also look at other similar variables.
-
 ```{r}
-#| label: fig-health-exp-aus
-#| fig-cap: "Health Expenditure (% of GDP), Country: Australia"
+#| label: fig-per-capita-health-expenditure-plot
+#| fig-cap: "Per Capita Current Health Expenditure (2010–2020): Top 5 nations"
 #| echo: FALSE
-#| out.width: "60%"
-knitr::include_graphics(here("results", "figures", "health_exp_aus.png"))
+knitr::include_graphics(here("results","figures","per_capita_health_expenditure_plot.png"))
 ```
 
-@fig-he_per_capita shows the distribution of Health Expenditure Per Capita (USD) over the Years.
-It can be observed that this variable indeed has variation and could be useful as our dependent variable.
-
-```{r}
-#| label: fig-he_per_capita
-#| fig-cap: "Health Expenditure Per Capita (USD, Country: Australia"
-#| echo: FALSE
-#| out.width: "60%"
-knitr::include_graphics(here("results","figures","health_exp_per_capita_aus.png"))
-```
-
-@fig-pop-growth-aus shows the distribution of population growth of australia.This figure shows that it has very less in common with @fig-health-exp-aus and more in common with @fig-he_per_capita
-
-```{r}
-#| label: fig-pop-growth-aus
-#| fig-cap: "Population Growth, Country: Australia"
-#| echo: FALSE
-#| out.width: "60%"
-knitr::include_graphics(here("results","figures","pop_growth_aus.png"))
-```
-
-I think it will be better for us to use the variable "Health Expenditure Per Capita \$USD" as a dependent variable in our analysis as it show more variation and connections with other variables.
+@fig-per-capita-health-expenditure-plot shows the distribution of "Per Capita Health Expenditure of top 5 nations between 2010-2020.
+This variable shows more variability across the countries, therefore we will go forward with "Per Capita Health Expenditure" as the factor that will represent the spending of different countries in health sectors.
 
@@
 
-@fig-regression-results shows the different regresion results based on different categories.
+@tbl-regression-results shows the different regresion results based on different categories.
 The categories are: Economic, Health, Social, Government, Trade.
@@
 ```{r}
-#| label: fig-regression-results
-#| fig-cap: "Regression Results by Category"
+#| label: tbl-regression-results
+#| tbl-cap: "Regression Results by Category"
 #| echo: FALSE
@@
 
+We can observe that health related variables are either missing observations or not statistically significant.
+But, it should not cause us much trouble because even if we had good observation within health related variables, there would be an equal chance of endogeniety and simultaneity bias with the dependent variable, leading to their exclusion from the model.
+
 ## Observing the relationship with dependent variable
 
-@fig-correlation-p1 show the correlation of different variables that had a statistical significant effect in @fig-regression-results (and other variables of interest).
+@fig-correlation-p1 show the correlation of different variables that had a statistical significant effect in @tbl-regression-results (and other variables of interest).
 We are just trying to see the pattern present in the dataset for the year 2010, with all the countries treated as cross-sectional units.
@@
 
-## Proposed Model for the research
+## Proposed Model for the research {#model}
 
-Based on the informations obtained in @fig-regression-results, we have selected variable that would best explain our dependent variable. The model is as follow:
+Based on the informations obtained in @tbl-regression-results, we have selected variable that would best explain our dependent variable. The model is as follow:
 
-Current Health Expenditure Per Capita (\$USD) = $b_0$ + $b_1$ Annual GDP growth + $b_2$ Income share held by lowest 20% + $b_3$ Control of corruption + $b_4$ Life expectancy at birth + $b_5$ Unemployed youth population as a percentage of total labor force + $b_6$ Adult literacy rate + $b_7$ Trade as a percentage of GDP + $b_8$ FDI net inflow + $b_9$ Access to electricity (rural) + $b_{10}$ Age dependency ratio
+Current Health Expenditure Per Capita (\$USD) = $b_0$ + $b_1$ Annual GDP growth + $b_2$ Income share held by lowest 20% + $b_3$ Control of corruption + $b_4$ Life expectancy at birth + $b_5$ Unemployed youth population as a percentage of total labor force + $b_6$ Trade as a percentage of GDP + $b_7$ FDI net inflow + $b_8$ Access to electricity (rural) + $b_{9}$ Age dependency ratio
+
+## Full analysis
+
+We ran the regression based on our primary model described in the [proposed model for the research](#model) section of this article.
+The results obtained from the regression analysis are presented by @tbl-preliminary-regression-results.
 
 ```{r}
-#| label: fig-preliminary-regression-results
-#| fig-cap: "Preliminary Regression Results"
+#| label: tbl-preliminary-regression-results
+#| tbl-cap: "Preliminary Regression Results"
 #| echo: FALSE
@@
 knitr::include_graphics(here("results", "tables", "preliminary_regression_results.png"))
-
 ```
 
-But, it turns out that the model shows significant amount of autocorrelation
+The result shows that annual GDP growth of a country has statistically significant negative impact on the per capita health expenditure of the country at 1% level of significance.
+Specifically, one percentage oint growth in annual GDP decreases the per capita health expenditure by 117.468 US dollars.
+Similarly, the estimate for the control of corruption in a country has a statistically significant negative effect at 1% level of significance.
+Specifically, one unit increase in control increases the per capita health expenditure by 1822 US dollars.
+The level of unemployment among youth population out of the total labor force population had a statistically significant negative effect at 1% level of significance.
+To be precise, one percent increase in unemployment rate among youth decreases the per capita health expenditure by 25 US dollars.
+Similarly, foreign direct investment (FDI) had a positive but very low magnitude effect on the health expenditure at 1% level of significance.
+Life expectancy at birth had a statistically significant positive effect on the health expenditure at 1% significance level, with increase in health expectancy by one year resulting in an increase in per capita health expenditure by 208 US dollars.
+Factors such as income share held by the lowest 20-percent population, trade as a percentage of GDP, access to electricity among rural popuation, and age dependency ration had no statistical significant impact on the health expenditure.
+
+The model showed a "good" descriptive performance based on it's adjusted R-squared value of 73.4%.
+The RMSE value for our model (RMSE=1213) is also quite lower compared to the NULL Model which showed a RMSE value of 2364.
+
+We performed the autocorrelation test. The results from the DW test presented by @tbl-DW-test-results confirms the absence of autocorrelation, therefore we can move forward with our model.
 
 ```{r}
-#| label: fig-DW-test-results
-#| fig-cap: "Durbin Watson Test Results"
+#| label: tbl-DW-test-results
+#| tbl-cap: "Durbin Watson Test Results"
 #| echo: FALSE
@@
 knitr::include_graphics(here("results", "figures", "durbin_watson_test_results.png"))
-
 ```
 
-**Therefore, we will use a regression model with country fixed effect which is given below:**
-
-Current Health Expenditure Per Capita (\$USD) = $b_0$ + $b_1$ Annual GDP growth + $b_2$ Income share held by lowest 20% + $b_3$ Control of corruption + $b_4$ Life expectancy at birth + $b_5$ Unemployed youth population as a percentage of total labor force + $b_6$ Adult literacy rate + $b_7$ Trade as a percentage of GDP + $b_8$ FDI net inflow + $b_9$ Access to electricity (rural) + $b_{10}$ Age dependency ratio + country fixed effect.
+We also performed the Fisher-type Augmented Dickey Fuller test for presence of unit root. The results confirmed that the not all series in our panel data are non-stationary.
 
 ```{r}
-#| label: fig-DW-test-results-fe
-#| fig-cap: "Durbin Watson Test Results FE-Model"
+#| label: tbl-unit-root-results
+#| tbl-cap: "Panel Unit Root Test Results"
 #| echo: FALSE
-#| out.width: "30%"
-knitr::include_graphics(here("results", "figures", "durbin_watson_test_results_fe.png"))
 
+htmltools::tags$iframe(
+  src = here("results", "tables", "unit_root_results.html"),
+  width = "100%",
+  height = "150px",
+  scrolling = "auto",
+  frameborder = "0"
+)
 ```
 
-## Full analysis
+We also performed the analysis on test dataset. The regression result obtained for the testing dataset is as follows.
 
-*To be added.*
+```{r}
+#| label: tbl-test-regression-results
+#| tbl-cap: "Regression Results for the test dataset"
+#| echo: FALSE
+#| out.width: "70%"
+knitr::include_graphics(here("results", "tables", "testresult.png"))
+```
+
+We can observe that the variables in general show some common pattern between the two datasets.
+The difference is acceptable as they are based on entirely different time frames.
+If we observe the model performance, it is evident that the model performance is weaker than obtined in the train dataset.
+The RMSE has increased by 383 from 1213 to 1596, however this is still 768 lower than the NULL model.
+The adjusted R-squared in our test dataset has also dropped by 12.3% to 62%
 
@@
 
-## Summary and Interpretation
+Our results shows that annual GDP growth, control of corruption, life expectancy at birth, unemployment rate, and FDI all have a statistically significant effect on the per capita health expenditure of a country.
+[@Nghiem2017] also found evidence that annual GDP and unemployment rate have a significant effect on the health expenditures, however the variables used in the studies are different.
+[@Factor2015] has found that country with high level of corruption have a lower spending on health care as a percentage of GDP.
 
-*To be added.*
+Factors such as income share held by the lowest 20-percent population, trade as a percentage of GDP, access to electricity among rural popuation, and age dependency ration had no statistical significant impact on the health expenditure.
+Among these variables, age dependency ratio was expected to have a significant impact as higher pressure on the working population to support the non-active population reduces the chances of surplus revenues to invest in different sectors of the economy.
 
-## Strengths and Limitations
+## Conclusions and Policy implications
 
-*To be added.*
+The results from our study concludes that corruption GDP growth, control of corruption, and  unemployment rate have a significant and an important role in determining the amount of money spent on health sector by the country.
+Therefore, if we are to improve the investment on health sectors and therefore improve the citizen's access to the heath services we also need to consider the unemployment rate, corruption and the GDP growth as well.
+This information could be handy for donor agencies and INGO's who work to improve access of health cervice in different nations.
 
-## Conclusions
+A special concideration should be given to nations that have an increasing GDP, high corruption, and high unemployment rate, as these nations have lower investments in health sectors.
+These type of countries should therefore be priotized by world bank, World Health Organization, and other institutions where allocating their resources to help countries.
 
-*To be added.*
+## Data aquisition
+
+The data for this research is available at the official World Bank- Data Bank website.It is a free database managed by world bank group and can be accessed anytime.
+We can apply various filters as per our need and export the data as a .CSV or an .xlsx file.
+
+Link to the source: <https://databank.worldbank.org/source/world-development-indicators/Type/TABLE/preview/on#>
 

### 2025-04-04 code/analysis-code/statistical-analysis.qmd
changes to regression tables

@@
 library(Metrics)
+library(kableExtra)
+library(here)
 
@@
 
+model1 <- lm(Current_health_expenditure_per_capita_current_US ~ 
+              GDP_growth_annual_ + 
+              Income_share_held_by_lowest_20 + 
+              Control_of_Corruption_Estimate + 
+              Life_expectancy_at_birth_total_years + 
+Unemployment_youth_total__of_total_labor_force_ages_1524_modeled_ILO_estimate + 
+              Trade__of_GDP+ Foreign_direct_investment_net_inflows_BoP_current_US+
+  Access_to_electricity_rural__of_rural_population +
+  Age_dependency_ratio__of_workingage_population +
+  Military_expenditure__of_GDP +
+  Literacy_rate_adult_total__of_people_ages_15_and_above
+
+  , 
+            data = train)
+
+modelsummary(
+  model1,
+  output = here("results", "tables", "preliminary_regression_results.png"),
+  title = "This table represents the preliminary regresssion model",
+  stars = c('*' = 0.1, '**' = 0.05, '***' = 0.01),
+  notes = list(
+    "Standard errors are shown in parentheses.",
+    "Significance levels: * p < 0.1, ** p < 0.05, *** p < 0.01."
+  )
+)
+
+```
+
+However, due to the missing values in the independent variables the total number of observation used in the regression is too low.
+So, we will use such variables which are available as much as possible for all countries.
+
+```{r}
+# primary regression model
+
 model <- lm(Current_health_expenditure_per_capita_current_US ~ 
@@
   model,
-  output = here("results", "tables", "preliminary_regression_results.png"),
-  title = "This table represents the preliminary regresssion model",
+  output = here("results", "tables", "primary_regression_results.png"),
+  title = "This table represents the primary regresssion model",
   stars = c('*' = 0.1, '**' = 0.05, '***' = 0.01),
@@
 ADF test (combined)
+
 ```{r}
@@
 
-# 1. Prepare panel data (handle NAs first)
+#handling NAs first
 train_clean <- na.omit(train[, c("CountryName", "Year", "Current_health_expenditure_per_capita_current_US")])
 
-# 2. Check time dimensions
+# time dimensions
 time_counts <- table(train_clean$country)
@@
 
-# 3. Run Fisher-type test with reduced lags
+# Fisher-type test with reduced lags
 pdata <- pdata.frame(train_clean, index = c("CountryName", "Year"))
@@
 ```
+ADF test (combined) continued....
 
 ```{r}
-library(kableExtra)
-library(here)
 
-# 1. Create properly formatted results table
+# 1. Creating properly formatted results table
 unit_root_table <- data.frame(
@@
 
-# 2. Create and save HTML table
+# creating and save HTML table
 unit_root_table %>%
   kbl(format = "html", digits = 4, align = "c",
-      caption = "Panel Unit Root Test Results") %>%
+      caption = "Primary model Unit Root Test Results") %>%
   kable_styling(bootstrap_options = c("striped", "hover")) %>%
@@
 
-# Option 1: Simple version with try()
-adf_results <- by(train, train$CountryName, function(x) {
-  try(adf.test(na.omit(x$Current_health_expenditure_per_capita_current_US)),
-  silent = TRUE)
-})
-
-# Option 2: More robust version with error handling
+# adf result
 adf_results <- lapply(split(train, train$CountryName), function(df) {
@@
 ```
+
+We will also test another version of our primary model excluding life expectancy to test if the results differ.
+
+```{r}
+# primary regression model
+
+model2 <- lm(Current_health_expenditure_per_capita_current_US ~ 
+              GDP_growth_annual_ + 
+              Income_share_held_by_lowest_20 + 
+              Control_of_Corruption_Estimate + 
+Unemployment_youth_total__of_total_labor_force_ages_1524_modeled_ILO_estimate + 
+              Trade__of_GDP+ Foreign_direct_investment_net_inflows_BoP_current_US+
+  Access_to_electricity_rural__of_rural_population +
+  Age_dependency_ratio__of_workingage_population
+  , 
+            data = train)
+
+modelsummary(
+  model2,
+  output = here("results", "tables", "primary_regression_excl_LE.png"),
+  title = "This table represents the primary regresssion model without life expectancy",
+  stars = c('*' = 0.1, '**' = 0.05, '***' = 0.01),
+  notes = list(
+    "Standard errors are shown in parentheses.",
+    "Significance levels: * p < 0.1, ** p < 0.05, *** p < 0.01."
+  )
+)
+
+```
+
+We can observe that the estimates and standard error don't change much even after the exclusion of life expectancy at birth, so our suspicion of endogeniety might not be that serious.

### 2025-04-04 code/eda-code/eda.qmd
changes to regression tables

@@
 ```
+
 We can not see a visible relation between these two factors.  Let's try it for France now
@@
 ```
+
 We can not observe a distinct relation between the two variable, we can observe that the Health Expenditure (% of GDP) variable also show less fluctuation across the years for multiple countries.
@@
 ```
+
 The Health Expenditure Per Capita (USD) varriable in fact shows more variation and also show some connections with annual population growth.

### 2025-04-04 products/manuscript/Manuscript.qmd
changes to regression tables

@@
 We ran the regression based on our primary model described in the [proposed model for the research](#model) section of this article.
-The results obtained from the regression analysis are presented by @tbl-preliminary-regression-results.
+The results obtained from the regression analysis are presented by @tbl-primary-regression-results.
 
 ```{r}
-#| label: tbl-preliminary-regression-results
-#| tbl-cap: "Preliminary Regression Results"
+#| label: tbl-primary-regression-results
+#| tbl-cap: "Primary Regression Results"
 #| echo: FALSE
 #| out.width: "70%"
-knitr::include_graphics(here("results", "tables", "preliminary_regression_results.png"))
+knitr::include_graphics(here("results", "tables", "primary_regression_results.png"))
 ```

### 2025-04-05 code/analysis-code/statistical-analysis.qmd
added RF results

@@
 ```
+
+Grouping the variables.
 
@@
 Now that we have selected our variables, we will try to look at the relation between our dependent variable "Health Expenditure Per Capita \$USD" with other variables from a specific category.
-
-```{r}
-
-# converting all columns to numeric except CountryName, CountryCode, and SeriesCode
-train <- train %>%
-  mutate(across(-c(CountryName, CountryCode, SeriesCode), as.numeric))
-
-#saving the train dataset as we will be using it in future as well
-
-saveRDS(train, file = here("data", "processed-data", "train-data.rds"))
-
-# converting all columns to numeric except CountryName, CountryCode, and SeriesCode
-test <- test %>%
-  mutate(across(-c(CountryName, CountryCode, SeriesCode), as.numeric))
-
-#saving the train dataset as we will be using it in future as well
-
-saveRDS(test, file = here("data", "processed-data", "test-data.rds"))
-```
 
@@
 ```
+
 ADF test (combined) continued....
@@
 ```{r}
-# primary regression model
+# primary regression model without LE
 

### 2025-04-05 code/eda-code/eda.qmd
added RF results

@@
 test <- mydata[(split_index + 1):nrow(mydata), ]
+```
+
+But first, let's save the dataset
+
+```{r}
+# converting all columns to numeric except CountryName, CountryCode, and SeriesCode
+train <- train %>%
+  mutate(across(-c(CountryName, CountryCode, SeriesCode), as.numeric))
+
+#saving the train dataset as we will be using it in future as well
+
+saveRDS(train, file = here("data", "processed-data", "train-data.rds"))
+
+# converting all columns to numeric except CountryName, CountryCode, and SeriesCode
+test <- test %>%
+  mutate(across(-c(CountryName, CountryCode, SeriesCode), as.numeric))
+
+#saving the train dataset as we will be using it in future as well
+
+saveRDS(test, file = here("data", "processed-data", "test-data.rds"))
 ```

### 2025-04-05 code/robustness-code/robustness-code.qmd
added RF results

@@
 ```{r}
-# Load required packages
+# Loading required packages
 library(purrr)    # For map_df()
@@
 
-testing the results
+Here, we will be testing the regression model on out test data
 
 ```{r}
-# preliminary regression model
+# primary regression model on test data
 
@@
 
+The results from our test dataset is show moderate performance of our model. Check the manuscript for more details on the comparision between train and test results.
+
 # cross validation
+
+Here, we will use cross validation to test the robustness of our model. We will have 5 year initial window and 1 year test window.
 
@@
 
-# 1. Define your model formula (as you originally specified)
+# defining the model  
 model_formula <- Current_health_expenditure_per_capita_current_US ~ 
@@
 
-# 2. Create time-series CV splits
+#creating time-series CV splits
 cv_splits <- time_series_cv(
@@
 
-    # Train model
+    #training model
     model <- lm(model_formula, data = train_cv)
 
-    # Get predictions and actuals
+    # Getting predictions and actuals
     preds <- predict(model, newdata = test_cv)
@@
 
-    # Convert to numeric safely
+    #converting to numeric 
     preds_num <- suppressWarnings(as.numeric(preds))
@@
 
-    # Calculate RMSE with error handling
+    #calculating RMSE
     if(length(preds_num) == 0 || length(actuals_num) == 0) {
@@
 
-# Run the analysis
+#saving the results
 results <- cv(cv_splits$splits, model_formula)
@@
 ```{r}
-# View summary of CV results
+#view summary of CV results
 cv_summary <- results %>%
@@
 
+The mean RMSE is 1151 which is simillar to the RMSE found in the original result (RMSE=1213). Therefore CV estimates show that our result is quite robust with very low Standard Deviation of 74.8
+
+We will also test if there is any other model that performs better than OLS regression. We will use Random Forest model.
+
 # Random Forest
@@
 
-    # 1. Get target variable from formula
+    # setting the target variables
     target_var <- all.vars(model_formula)[1]
 
-    # 2. Check if Year column exists (case-insensitive)
+    # checking if Year column exists
     year_col <- grep("^year$", names(train_cv), ignore.case = TRUE, value = TRUE)
 
-    # 3. Select required variables (including Year if it exists)
+    #selecting required variables
     required_vars <- unique(c(all.vars(model_formula), year_col))
@@
 
-    # 4. Get year information
+    # getting years
     get_years <- function(df) {
@@
 
-    # 5. Calculate mtry
+    #calculate mtry
     n_predictors <- length(setdiff(required_vars, c(target_var, year_col)))
@@
 
-    # 6. Train model
+    #train model
     set.seed(123)
@@
 
-    # 7. Calculate RMSE
+    #calculate RMSE
     preds <- predict(rf_model, data = test_cv)$predictions
@@
 
-    # 8. Return results
+    #return results
     tibble(
@@
 
+The results of Random Forest are also quite good. They show very low RMSE compared to the OLS Moldel.
+
 # Random Forest on test dataset
@@
 
-# Prepare training data (remove NA in target variable)
+#preparing training data
 target_var <- all.vars(model_formula)[1]
@@
 
-# Calculate appropriate mtry
+#calculate appropriate mtry
 n_predictors <- length(all.vars(model_formula)) - 1
@@
 
-# Train Random Forest model
+#training Random Forest model
 set.seed(123)
@@
 ```{r}
-# Ensure test data has same structure as training data
+#processing test data
 test_processed <- test %>%
-  select(any_of(names(train_processed))) %>%  # Keep only columns present in training
-  drop_na(all_of(target_var))  # Remove rows with NA in target
+  select(any_of(names(train_processed))) %>%  # keeping only columns present in training
+  drop_na(all_of(target_var))  # Removing rows with NA in target
 ```
@@
 ```{r}
-# Generate predictions
+# Generating predictions
 test_predictions <- predict(rf_model, data = test_processed)$predictions
 
-# Create results dataframe
+#creating results dataframe
 test_results <- test_processed %>%
@@
 ```{r}
-# Calculate metrics
+#calculating metrics
 performance_metrics <- test_results %>%
@@
 
-# Print metrics
+# printing metrics
 cat("\nTest Set Performance:\n")
@@
 
-# View predictions vs actual
+```
+
+However, the results from test dataset are not promising for our Random Forest model. So, we will stick with our OLS model.
+
+```{r}
+#viewing the predictions vs actuals
 head(test_results %>% select(all_of(target_var), Predicted, Residual))

### 2025-04-05 products/manuscript/Manuscript.qmd
added RF results

@@
 The RMSE has increased by 383 from 1213 to 1596, however this is still 768 lower than the NULL model.
-The adjusted R-squared in our test dataset has also dropped by 12.3% to 62%
+The adjusted R-squared in our test dataset has also dropped by 12.3% to 62%.
+We also tested Random Forest model as an alternative of the OLS model, and althouth Random Forest showed promising results in our train dataset, the results were worse than OLS model for the test dataset.
 

### 2025-04-05 products/manuscript/Manuscript.qmd
abstract

@@
 
-$\land$ Corresponding author: some\@email.com
+$\land$ Corresponding author: prasanga.paudel@email.com
 
@@
 
-The data is based on World Development Indicator from World Bank Group.
-This is a comprehensive database that includes information including employment, public and private expenditure in health sector, fertility, immunization, literacy, mortality, and many more.The dataset has a total of 1496 variables and across 11 years from 2010 to 2020.
-The actual source of data has information from 1960 to 2023 but to deal with the problem of missing values in recent years and overcomplication in analysis due to too many year data , we will focus on 10 recent years between 2010 and 2020.
+_We used Ordinary Least Squares method to analyze the factors affecting per capita health expenditure of a country among the OECD nations during 2010 to 2020.
+We found that corruption GDP growth, control of corruption, and  unemployment rate have a significant and an important role in determining the amount of money spent on health sector by the country.
+Foreign Direct Investment had a significant but low magnitude impact.
+Factors such as income share held by the lowest 20-percent population, trade as a percentage of GDP, access to electricity among rural popuation, and age dependency ration had no statistical significant impact on the health expenditure._
 
@@
 We can relate the expenditure with different national development indicators and hope to find important connections between them.
-This will help government to make specific interventions through their policy to enhance their health sector.
+This will help government and international institutions to make specific interventions through their policy to enhance their health sector.
 We can also find out how countries are currently operating and which factor has helped or negatively affected their health sector.
@@
 
-The original data in it's raw form was obtained from World Bank.
+The original data (2010-2020) in it's raw form was obtained from World Bank.
 However, the dataset needed processing, and some modification.

### 2025-04-19 code/analysis-code/statistical-analysis.qmd
project #5

@@
 library(broom)      
-library(kableExtra) 
 library(webshot)
@@
 ```
+
+# Importing datasets
 
@@
 
-Grouping the variables.
+# Grouping the variables.
 
@@
 Now that we have selected our variables, we will try to look at the relation between our dependent variable "Health Expenditure Per Capita \$USD" with other variables from a specific category.
+
+# Preliminary Regression
 
@@
 # Save the table as a HTML file
-html_file <- here("results", "tables", "regression_results.html")
+html_file <- here("products", "supplementary-files", "regression_results.html")
 save_kable(table_html, file = html_file)
@@
 Variable selection is a continuous process and the variables selected in the preliminary model may also be dropped later to make space for new variables that may appear as important.
+
+# Suggested Model
 
@@
 So, we will use such variables which are available as much as possible for all countries.
+
+# Primary regression model
 
@@
 
+# Null Model
+
 Calculating the RMSE of a NULL Model.
@@
 ```
+
+# ADF test
 
@@
   kbl(format = "html", digits = 4, align = "c",
-      caption = "Primary model Unit Root Test Results") %>%
+      caption = "") %>%
   kable_styling(bootstrap_options = c("striped", "hover")) %>%
-  add_footnote("Null hypothesis: All series contain unit roots", 
+  add_footnote("", 
               notation = "none") %>%
@@
 
+# Regression model without Life Expectancy
+
 ```{r}

### 2025-04-19 code/eda-code/eda.qmd
project #5

@@
 library(webshot2)
+library(patchwork)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
 library(patchwork)
@@
 
-# Create a summary table, excluding specific rows
+# creating a summary table, excluding specific rows
 country_summary_table <- data_2010 %>%
@@
 
-# Generate the HTML table
+# generate the HTML table
 table_html <- country_summary_table %>%
@@
 
-# Save the table object as an RDS file
+# saving the table object as an RDS file
 rds_file <- here::here("results", "tables", "health_exp_2010.rds")
@@
 # Creating the line plot
-ggplot(data_filtered, aes(x = Year, y = Current_health_expenditure__of_GDP, color = CountryName, group = CountryName)) +
+p1<- ggplot(data_filtered, aes(x = Year, y = Current_health_expenditure__of_GDP, color = CountryName, group = CountryName)) +
   geom_line(size = 1) +
@@
 The Health Expenditure Per Capita (USD) varriable in fact shows more variation and also show some connections with annual population growth.
+
 ```{r}
@@
 # Creating the line plot
-ggplot(data_filtered, aes(x = Year, y = Current_health_expenditure_per_capita_current_US, color = CountryName, group = CountryName)) +
+p2<- ggplot(data_filtered, aes(x = Year, y = Current_health_expenditure_per_capita_current_US, color = CountryName, group = CountryName)) +
   geom_line(size = 1) +
@@
     x = "Year",
-    y = "Health Expenditure (% of GDP)",
+    y = "Per capita Health Expenditure (USD)",
     color = "Country"
@@
 
+saving the health expenditure plot
+
+```{r}
+combined_plot <- p1 / p2 + 
+plot_layout(widths = c(1, 1)) +  # Equal width for both plots
+plot_annotation(tag_levels = 'A')
+
+ggsave(
+  filename = here("results", "figures", "combined_health_expenditure_plots.png"),
+  plot = combined_plot,
+  width = 10,  #  width as needed
+  height = 8,  #  height as needed
+  dpi = 300
+)
+```
+
 # EDA (part2)
@@
 ```{r}
-# Define dataset
+# defining dataset
 
@@
 
-# Arrange plots in a 2x2 grid
+# arranging plots in a 2x2 grid
 final_plot1 <- (plot1 + plot2) / (plot3 + plot4)
@@
 
-# Generate individual plots with different colors
+# generating individual plots with different colors
 plot5 <- create_corr_plot("Literacy_rate_adult_total__of_people_ages_15_and_above", df, "green")
@@
 
-# Arrange plots in a 2x2 grid
+# arranging plots in a 2x2 grid
 final_plot2 <- (plot5 + plot6) / (plot7 + plot8)
@@
 ```
+
+# EDA (part 3)
+
+## Figures for manuscript
+
+```{r}
+
+#cleaning and preparing the data
+trend_data <- train %>%
+  # Remove non-country rows
+  filter(!CountryName %in% c(
+    "Data from database: World Development Indicators",
+    "OECD members",
+    "Last Updated: 01/28/2025")) %>%
+
+  # renaming columns
+  select(
+    year = Year,
+    country = CountryName,
+    gdp = GDP_current_US,
+    population = Population_total,
+    life_exp = Life_expectancy_at_birth_total_years,
+    physicians = Physicians_per_1000_people,
+    hospital_beds = Hospital_beds_per_1000_people,
+    gov_effectiveness = Government_Effectiveness_Estimate,
+    corruption = Control_of_Corruption_Estimate,
+    rule_of_law = Rule_of_Law_Estimate,
+    electricity = Access_to_electricity__of_population,
+    unemployment = Unemployment_total__of_total_labor_force_modeled_ILO_estimate,
+    poverty = Poverty_headcount_ratio_at_national_poverty_lines__of_population
+  ) %>%
+
+  #converting to long format
+  pivot_longer(
+    cols = -c(year, country),
+    names_to = "indicator",
+    values_to = "value"
+  ) %>%
+
+  #calculating yearly averages
+  group_by(year, indicator) %>%
+  summarise(avg_value = mean(value, na.rm = TRUE)) %>%
+  ungroup()
+
+#making a reusable plotting function
+create_trend_plot <- function(data, indicator_name, plot_title, y_label, color) {
+  data %>%
+    filter(indicator == indicator_name) %>%
+    ggplot(aes(x = year, y = avg_value)) +
+    geom_line(color = color, linewidth = 1) +
+    geom_point(color = color, size = 2) +
+    labs(title = plot_title, y = y_label, x = NULL) +
+    theme_minimal(base_size = 12) +
+    theme(plot.title = element_text(size = 11))
+}
+
+#now defining colors for consistency
+plot_colors <- c(
+  "#1f77b4", "#ff7f0e", "#2ca02c", "#d62728",
+  "#9467bd", "#8c564b", "#e377c2", "#7f7f7f",
+  "#bcbd22", "#17becf", "#393b79"
+)
+
+#generating the plots
+plots <- list(
+  gdp = create_trend_plot(trend_data, "gdp", "GDP (current US$)", "USD", plot_colors[1]),
+  population = create_trend_plot(trend_data, "population", "Total Population", "Count", plot_colors[2]),
+  life_exp = create_trend_plot(trend_data, "life_exp", "Life Expectancy", "Years", plot_colors[3]),
+  physicians = create_trend_plot(trend_data, "physicians", "Physicians per 1000", "Count", plot_colors[4]),
+  hospital_beds = create_trend_plot(trend_data, "hospital_beds", "Hospital Beds per 1000", "Count", plot_colors[5]),
+  gov_effectiveness = create_trend_plot(trend_data, "gov_effectiveness", "Government Effectiveness", "Score", plot_colors[6]),
+  corruption = create_trend_plot(trend_data, "corruption", "Control of Corruption", "Score", plot_colors[7]),
+  rule_of_law = create_trend_plot(trend_data, "rule_of_law", "Rule of Law", "Score", plot_colors[8]),
+  electricity = create_trend_plot(trend_data, "electricity", "Rural Electricity Access", " Percentage", plot_colors[9]),
+  unemployment = create_trend_plot(trend_data, "unemployment", "Unemployment Rate", "% Labor Force", plot_colors[10]),
+  poverty = create_trend_plot(trend_data, "poverty", "Poverty Headcount", "% of Population", plot_colors[11])
+)
+
+#combining plots ----
+#create a 3-column layout 
+combined_plot <- wrap_plots(plots, ncol = 3) +
+  plot_annotation(
+    title = "Trends in Key Development Indicators",
+    subtitle = "Country averages across time",
+    caption = "Source: World Development Indicators",
+    theme = theme(
+      plot.title = element_text(size = 16, face = "bold"),
+      plot.subtitle = element_text(size = 12)
+    )
+  )
+
+#saving the plot
+ggsave(
+  filename = here::here("results", "figures", "development_indicators_trends.png"),
+  plot = combined_plot,
+  width = 16,
+  height = 14,
+  dpi = 300
+)
+
+```

### 2025-04-19 code/processing-code/processingfile-v1.qmd
part #4
project #5

@@
 ---
+
+Please see the readme file before running the codes.
 
@@
 Here, we will transform the variables as column names with appropriates values assigned to them correctly.
+
+CAUTION: This chunck might take a long time to process. Please wait pariently for few minutes (2-5 minutes if your PC has 16gb RAM) until the code processes itself.
+
 ```{r}
@@
 
+CAUTION: This chunck might also consume a good amoung of your PC's memory, expect upto 1-2 minutes for the code to process.
+
 ```{r}
@@
 
+CAUTION: This chunk could also take upto 1-2 minutes to process.
+
 ```{r}
@@
 xl_path <- here("data", "processed-data", "processeddata.xlsx")
-writexl::write_xlsx(processeddata, xl_path)
+writexl::write_xlsx(data_collapsed, xl_path)
 ```

### 2025-04-19 code/robustness-code/robustness-code.qmd
project #5

@@
 ---
+
+NOTE: Not all codes may produce an output, you can find those outputs saved in the respective folder location.
 
@@
 library(ranger) 
+library(knitr)
+library(kableExtra)
 
 ```
+# Importing dataset
 
@@
 ```
+
+# OLS regression on dataset
 
@@
 
-# cross validation
+# cross validation of OLS model
 
@@
 print(results)
+
+#saving the results
+#saving the table
+cv_results <- kable(results, format = "html", digits = 3) %>%
+kable_styling()
+save_kable(cv_results, file = here("results", "tables","OLS-CV-results-full.html"))
 ```
+
+Now we will see the summary of the CV results
 
@@
 print(cv_summary)
+
+#saving the table
+cv_summary_kable <- kable(cv_summary, format = "html", digits = 3) %>%
+kable_styling()
+save_kable(cv_summary_kable, file = here("results", "tables","OLS-CV-results.html"))
 ```
@@
 
+## Random Forest using train dataset
+
 ```{r}
-rf_cv <- function(splits, model_formula) {
+
+target_var <- all.vars(model_formula)[1]
+train_processed <- train %>%
+  drop_na(all_of(target_var))
+
+#calculate appropriate mtry
+n_predictors <- length(all.vars(model_formula)) - 1
+mtry_value <- max(1, min(floor(n_predictors / 3), n_predictors))
+
+# Train Random Forest on full training dataset
+set.seed(123)
+rf_full_model <- ranger(
+  formula = model_formula,
+  data = train_processed,
+  num.trees = 500,
+  mtry = mtry_value,
+  importance = 'permutation',
+  seed = 123
+)
+
+# Extract variable importance
+importance_df <- data.frame(
+  Variable = names(rf_full_model$variable.importance),
+  Importance = rf_full_model$variable.importance
+) %>% 
+  arrange(desc(Importance))
+
+# Create importance plot
+library(ggplot2)
+importance_plot <- ggplot(importance_df, aes(x = reorder(Variable, Importance), 
+                         y = Importance)) +  # Fixed this line - removed extra parenthesis
+  geom_col(fill = "steelblue") +  # Simplified this line
+  coord_flip() +
+  labs(title = "Random Forest Variable Importance",
+       x = "Predictor Variables",
+       y = "Importance (Permutation)") +
+  theme_minimal()
+
+# Save the plot
+ggsave(here("results", "figures", "rf_importance_plot.png"), 
+       plot = importance_plot,
+       width = 8, height = 6, dpi = 300)
+```
+
+Here we will calcualte the performance metrics of our RF model.
+
+```{r}
+train_predictions <- predict(rf_full_model, data = train_processed)$predictions
+
+# Calculate performance metrics
+train_performance <- train_processed %>%
+  mutate(Predicted = train_predictions) %>%
+  summarise(
+    RMSE = sqrt(mean((.data[[target_var]] - Predicted)^2)),
+    MAE = mean(abs(.data[[target_var]] - Predicted)),
+    R2 = cor(.data[[target_var]], Predicted)^2
+  ) %>%
+  mutate(across(everything(), ~round(., 3)))  # Round all metrics to 3 decimal places
+
+# Display the performance metrics
+print(train_performance)
+
+```
+
+Saving the table
+
+```{r}
+library(kableExtra)
+ train_table <- train_performance %>%
+  pivot_longer(everything(), names_to = "Metric", values_to = "Value") %>%
+  mutate(Metric = case_when(
+    Metric == "RMSE" ~ "Root Mean Squared Error",
+    Metric == "MAE" ~ "Mean Absolute Error",
+    Metric == "R2" ~ "R-squared",
+    TRUE ~ Metric
+  )) %>%
+  kable(format = "html", align = c("l", "r")) %>%
+  kable_styling(bootstrap_options = c("striped", "hover"))
+
+kableExtra::save_kable(train_table, 
+                       file = here("results", "tables", "rf_train_performance.html"))
+```
+
+## Random Forest with CV on train dataset
+
+```{r}
+rf_cv <- function(splits, model_formula, include_importance = TRUE) {
   map_df(splits, function(split) {
@@
 
+Saving the RF results
+
+```{r}
+# Select only the desired columns
+rf_results_table <- rf_results %>%
+  select(Train_Years, Test_Year, RMSE, Predictors_Used, mtry_Used)
+
+# Create and save the HTML table
+rf_results_html <- kable(rf_results_table, 
+                         format = "html",
+                         digits = 2,
+                         caption = "") %>%
+  kable_styling(bootstrap_options = c("striped", "hover"),
+                full_width = FALSE,
+                position = "left") %>%
+  add_header_above(c("Training Period" = 1, "Test Year" = 1, 
+                     "Performance" = 1, "Model Parameters" = 2))
+
+# Save the table
+save_kable(rf_results_html, 
+           file = here("results", "tables", "rf_cv_results.html"))
+```
+
+Now we will calculate the summary of the above results.
+
+```{r}
+
+# Create summary table first
+rf_summary_table <- rf_results %>%
+  summarise(
+    `Mean RMSE` = mean(RMSE, na.rm = TRUE),
+    `SD RMSE` = sd(RMSE, na.rm = TRUE),
+    `Min RMSE` = min(RMSE, na.rm = TRUE),
+    `Max RMSE` = max(RMSE, na.rm = TRUE)
+  ) %>%
+  mutate(across(everything(), ~round(., 3)))  # Round all values
+
+# Then create and save HTML table
+rf_summary_html <- kable(rf_summary_table, 
+                         format = "html",
+                         caption = "") %>%
+  kable_styling(bootstrap_options = c("striped", "hover"),
+                full_width = FALSE,
+                position = "center") %>%
+  row_spec(0, bold = TRUE)
+
+save_kable(rf_summary_html, file = here("results", "tables", "rf_cv_summary.html"))
+```
+
 # Random Forest on test dataset
@@
 ```{r}
-#calculating metrics
-performance_metrics <- test_results %>%
-  summarise(
-    RMSE = sqrt(mean(Residual^2, na.rm = TRUE)),
-    MAE = mean(abs(Residual), na.rm = TRUE),
-    R2 = 1 - (sum(Residual^2) / sum((.data[[target_var]] - mean(.data[[target_var]]))^2))
+# Calculate metrics manually
+performance_metrics <- data.frame(
+  RMSE = sqrt(mean(test_results$Residual^2)),
+  MAE = mean(abs(test_results$Residual)),
+  R2 = cor(test_results[[target_var]], test_results$Predicted)^2
 )
 
-# printing metrics
-cat("\nTest Set Performance:\n")
 print(performance_metrics)
+```
+
+```{r}
+
+# Format the performance metrics into a clean table
+performance_table <- performance_metrics %>%
+  pivot_longer(cols = everything(), 
+               names_to = "Metric", 
+               values_to = "Value") %>%
+  mutate(Metric = case_when(
+    Metric == "RMSE" ~ "Root Mean Squared Error",
+    Metric == "MAE" ~ "Mean Absolute Error",
+    Metric == "R2" ~ "R-squared",
+    TRUE ~ Metric
+  ),
+  Value = round(Value, 3))  # Round to 3 decimal places
+
+# Create and save the HTML table
+performance_html <- kable(performance_table, 
+                         format = "html",
+                         col.names = c("Performance Metric", "Value"),
+                         align = c("l", "r"),
+                         caption = "") %>%
+  kable_styling(bootstrap_options = c("striped", "hover"),
+                full_width = FALSE,
+                position = "center") %>%
+  row_spec(0, bold = TRUE, background = "#f8f9fa") %>%
+  column_spec(2, width = "5cm")
+
+# Save the table
+save_kable(performance_html, 
+           file = here("results", "tables", "rf_test_performance.html"))
 
@@
 
+# Visualizing the models train dataset
+
 ```{r}
-#viewing the predictions vs actuals
-head(test_results %>% select(all_of(target_var), Predicted, Residual))
+# For Random Forest model 
+rf_train_predictions <- predict(rf_full_model, data = train_processed)$predictions
+
+# For OLS model 
+ols_model <- lm(model_formula, data = train_processed)
+ols_train_predictions <- predict(ols_model, newdata = train_processed)
+
+#creating combined data frame for plotting
+plot_data <- train_processed %>%
+  mutate(
+    RF_Predicted = rf_train_predictions,
+    OLS_Predicted = ols_train_predictions,
+    Observed = .data[[target_var]]
+  )
+
+#creating the comparison plot
+library(ggplot2)
+pred_vs_obs_plot <- ggplot(plot_data, aes(x = Observed)) +
+  geom_point(aes(y = RF_Predicted, color = "Random Forest"), alpha = 0.6) +
+  geom_point(aes(y = OLS_Predicted, color = "OLS Regression"), alpha = 0.6) +
+  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black") +
+  labs(
+    title = "Predicted vs Observed Values (Training Data)",
+    x = "Observed Health Expenditure per Capita (US$)",
+    y = "Predicted Health Expenditure per Capita (US$)",
+    color = "Model Type"
+  ) +
+  scale_color_manual(values = c("Random Forest" = "#1f77b4", 
+                               "OLS Regression" = "#ff7f0e")) +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+#saing the plot
+ggsave(here("results", "figures", "pred_vs_obs_train.png"),
+       plot = pred_vs_obs_plot,
+       width = 8, height = 6, dpi = 300)
+
+# Display  plot
+print(pred_vs_obs_plot)
 ```
+
+# Visualizing the models test dataset
+
+```{r}
+#generating predictions for test data as we did for train data
+
+# Random Forest predictions 
+rf_test_predictions <- predict(rf_model, data = test_processed)$predictions
+
+# OLS predictions on test data
+ols_test_predictions <- predict(model, newdata = test_processed)  #
+
+# combined data frame for plotting
+test_plot_data <- test_processed %>%
+  mutate(
+    RF_Predicted = rf_test_predictions,
+    OLS_Predicted = ols_test_predictions,
+    Observed = .data[[target_var]]
+  )
+
+# comparison plot
+test_pred_vs_obs_plot <- ggplot(test_plot_data, aes(x = Observed)) +
+  geom_point(aes(y = RF_Predicted, color = "Random Forest"), alpha = 0.6, size = 2) +
+  geom_point(aes(y = OLS_Predicted, color = "OLS Regression"), alpha = 0.6, size = 2) +
+  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black") +
+  labs(
+    title = "Predicted vs Observed Values (Test Data)",
+    x = "Observed Health Expenditure per Capita (US$)",
+    y = "Predicted Health Expenditure per Capita (US$)",
+    color = "Model Type"
+  ) +
+  scale_color_manual(values = c("Random Forest" = "#E69F00",  
+                               "OLS Regression" = "#56B4E9")) +  
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        plot.title = element_text(hjust = 0.5)) +  # Center title
+  coord_equal()  # Equal axis scaling for proper 45° line
+
+#adding performance metrics to plot
+test_performance <- test_plot_data %>%
+  summarise(
+    RF_RMSE = sqrt(mean((Observed - RF_Predicted)^2)),
+    OLS_RMSE = sqrt(mean((Observed - OLS_Predicted)^2))
+  )
+
+test_pred_vs_obs_plot <- test_pred_vs_obs_plot +
+  annotate("text", x = min(test_plot_data$Observed), 
+           y = max(c(test_plot_data$RF_Predicted, test_plot_data$OLS_Predicted)),
+           hjust = 0, vjust = 1,
+           label = sprintf("RMSE:\nRandom Forest: %.1f\nOLS: %.1f",
+                          test_performance$RF_RMSE, 
+                          test_performance$OLS_RMSE))
+
+#saving the plot
+ggsave(here("results", "figures", "pred_vs_obs_test.png"),
+       plot = test_pred_vs_obs_plot,
+       width = 8, height = 6, dpi = 300)
+
+#show the plot
+print(test_pred_vs_obs_plot)
+```

### 2025-04-19 products/manuscript/Manuscript.qmd
project #5

@@
 ---
-
-This project uses HTML as output format.
 
@@
 
--   Prasanga Paudel$^{1,2,*}$ (ORCID: 0000-0000-1234-5678)
+-   Prasanga Paudel$^{1}$ (ORCID: 0000-0000-1234-5678)
 
@@
 1.  College of Agricultural and Environmental Science, University of Georgia, Athens, GA, USA.
-
-$*$ These authors contributed equally to this work.
-
-$\land$ Corresponding author: prasanga.paudel@email.com
 
@@
 _We used Ordinary Least Squares method to analyze the factors affecting per capita health expenditure of a country among the OECD nations during 2010 to 2020.
-We found that corruption GDP growth, control of corruption, and  unemployment rate have a significant and an important role in determining the amount of money spent on health sector by the country.
+We found that GDP growth, control of corruption, and  unemployment rate among the youth in the labor force have a significant and an important role in determining the amount of money spent on health sector by the country.
 Foreign Direct Investment had a significant but low magnitude impact.
-Factors such as income share held by the lowest 20-percent population, trade as a percentage of GDP, access to electricity among rural popuation, and age dependency ration had no statistical significant impact on the health expenditure._
+Factors such as income share held by the lowest 20-percent population, trade as a percentage of GDP, access to electricity among rural popuation, and age dependency ration had no statistical significant impact on the health expenditure.
+We also applied Random Forest model to explain the factors affecting per capita health expenditure of a country.
+The results of both the OLS model and Random forest was tested using cross-validation and train/test technique, and we found that our OLS model shows a consistent result across all test.
+Although Random Model was found to have a better explanation capability, the inconsistensies raised a question on it's validity.
+_
 
@@
 
-## Description of data and data source
+## Data source
 
@@
 However, the data obtained from this dataset is crude and needed to be treated before attempting analysis based on it.
-The details about the data cleaning process can be assessed from this github repository.
+The details about the data cleaning process can be assessed from this [ GitHub](https://github.com/xprasanga/PAUDEL-MADA-project) repository.
 
@@
 
-*The main objective of this research is to:*
+The main objective of this research is to:
 
-*1. To analyze the factors that affect the expenditure in health sector among OECD countries.*
+1. To analyze the factors that affect the expenditure in health sector among OECD countries.
 
-*2. To analyze the magnitude of the influence of these factors on health expenditure*
+2. To analyze the magnitude of the influence of these factors on health expenditure.
 
@@
 
-## A look into the train dataset
-
-@tbl-summary shows the values under interesting variables for all OECD countries for the year 2010.
-
-```{r}
-#| label: tbl-summary
-#| tbl-cap: "Summary of variables"
-#| echo: FALSE
-table_html <- readRDS(here::here("results", "tables", "health_exp_2010.rds"))
-table_html
-```
-
 {{< pagebreak >}}
@@
 
-In exploratory data analysis, we will just look at data from Australia as a reference for relation between variables in most of the cases.
-We are currently thinking of using "Current Health Expenditure as a % of GDP" as our dependent variable.
-But, we will see it's distribution and other alternatives before making decision.
+In this section, we will we will explore the data based.
 
-## Exploratory/Descriptive analysis
+## Descriptive analysis
 
-*Here, we will now use a combination of tables and figues to have a look into the data.*
+Here, we will now use a combination of tables and figues to have a look into the data.
 
@@
 
-### Looking into the trend for our possible dependent variables
+### Trends of Development Indicators
 
-@fig-health-expenditure-plot shows how "Health Expenditure as a Percentage of GDP Over Time (2010–2020)" is distributed over the years for top 5 OECD countries.
+In @fig-development-indicators-trends, we will explore differenct variables of interest by observing the average trends across the countries.
+We can observe that average life expectancy has been increasing consistently through 2010 to 2020.
+The average physicians per 1000 people shows an upward trend.
+However, hospital beds show a decreasing trend.
+The average government effectiveness seems to be dereasing across the countries.
+The unemployment rate in the OECD countries seems to be consistently decreasing.
+Finally, if we observe the average Gross Domestic Product(GDP) , we can see that the GDP of OECD countries are somewhat consistent except for the last year 2018.
 
 ```{r}
-#| label: fig-health-expenditure-plot
-#| fig-cap: "Current Health Expenditure as a % of GDP (2010–2020): Top 5 nations"
+#| label: fig-development-indicators-trends
+#| fig-cap: "Trends of Development Indicators (Average across countries)"
 #| echo: FALSE
-knitr::include_graphics(here("results","figures","health_expenditure_plot.png"))
+knitr::include_graphics(here("results","figures","development_indicators_trends.png"))
 ```
 
-But, before moving on to be fixed at a single variable as our dependent variable, we will also look if there are any alternative dependent varibles that could prove to be more useful.
+### Trends of health expenditure
+
+@fig-combined-health-expenditure-plot shows how "Health Expenditure as a Percentage of GDP Over Time (2010–2020)" and "Current Per Capita Health Expenditure (2010-2020)" are distributed over the years for top 5 OECD countries.
+We can observe that among the two factors, Health Expenditure as a percentage of GDP is relatively more consistent across the years.
+Per capita health expenditure across the countries, on the other hand, shows variation across times.
+It is evident that per capaita health expenditure is more sensitive to external factors and shows immediate response through each years.
+Our objective is to explain this variation in expenditure.
 
 ```{r}
-#| label: fig-per-capita-health-expenditure-plot
-#| fig-cap: "Per Capita Current Health Expenditure (2010–2020): Top 5 nations"
+#| label: fig-combined-health-expenditure-plot
+#| fig-cap: "Health expenditure across top 5 countries"
 #| echo: FALSE
-knitr::include_graphics(here("results","figures","per_capita_health_expenditure_plot.png"))
+knitr::include_graphics(here("results","figures","combined_health_expenditure_plots.png"))
 ```
 
-@fig-per-capita-health-expenditure-plot shows the distribution of "Per Capita Health Expenditure of top 5 nations between 2010-2020.
-This variable shows more variability across the countries, therefore we will go forward with "Per Capita Health Expenditure" as the factor that will represent the spending of different countries in health sectors.
+## Basic statistical analysis
 
-# Basic statistical analysis
+### Preparing a model
 
-## Preparing a model
-
-To help finalize our model, we will first try to look at the relation between our dependent variable "Health Expenditure Per Capita \$USD" with other variables from a specific category.
-Specifically, we will pick the most useful variables from the large pool of 1400+ variables in our dataset, and we will then categorize these variables based on the aspects they cover.
-For example, if there is avariable named "Annual GDP growth", we will categorize it as an "Economic" variable.
+To help finalize our model, we first tried to look at the relation between our dependent variable "Health Expenditure Per Capita $USD" with other variables from a specific category.
+Specifically, we picked the most useful variables from the large pool of 1400+ variables in our dataset, and we then categorized these variables based on the aspects they cover.
+For example, if there is a variable named "Annual GDP growth", we will categorize it as an "Economic" variable.
 Similarly, if there is a variable named "Unemployment rate" or "Total population", we will categorize them as "Social" variable
 
-We will then run our initial regression with "Health Expenditure Per Capita \$USD" as dependent variable, and variables from each of the category as independent variables in separate regressions
+We then ran our initial regression with "Health Expenditure Per Capita $USD" as dependent variable, and variables from each of the category as independent variables in separate regressions.
+The regression results for this preliminary test can be observed under this [ GitHub](https://github.com/xprasanga/PAUDEL-MADA-project) repository under supplementary-materials.
 
-@tbl-regression-results shows the different regresion results based on different categories.
-The categories are: Economic, Health, Social, Government, Trade.
-We will pick the most useful variables among the variables that came up significant and create our preliminary model for our research.
-However, this would not be the end of our variable selection process.
+After observing the relation of almost 60 important variables with our variable of interest "Health Expenditure Per Capita $USD", we filtered few variables worth pursuing, and we will discuss them in the sections that will come.
 
-```{r}
-#| label: tbl-regression-results
-#| tbl-cap: "Regression Results by Category"
-#| echo: FALSE
-htmltools::tags$iframe(
-  src = here("results", "tables", "regression_results.html"),
-  width = "100%",
-  height = "600px"
-)
-```
+### Observing the relationship with dependent variable
 
-We can observe that health related variables are either missing observations or not statistically significant.
-But, it should not cause us much trouble because even if we had good observation within health related variables, there would be an equal chance of endogeniety and simultaneity bias with the dependent variable, leading to their exclusion from the model.
-
-## Observing the relationship with dependent variable
-
-@fig-correlation-p1 show the correlation of different variables that had a statistical significant effect in @tbl-regression-results (and other variables of interest).
-We are just trying to see the pattern present in the dataset for the year 2010, with all the countries treated as cross-sectional units.
+@fig-correlation-p1 shows the correlation of per capita health expenditure with different variables for the year 2010.
+This fugure shows the pattern present in the dataset for the year 2010, with all the countries treated as cross-sectional units.
+We are only using the information from one year, year 2010, because visualization across years can be quite complex as we would also have to consider for the trends of each country across years.
 
@@
 
-We do the same with few other variables.
+@fig-correlation-p1 shows that income share held by lowest 20% of the population is positively correlated with the per capita health expenditure.
+Similary, estimate of control of corruption and life expectancy at birth also show positive correlation.
+Annual GDP growth on the other hand shows a low but negative correlation.
+
+We will check few other relation in @fig-correlation-p2.
 
@@
 
-It can be seen that most of the variables seem to have a high correlation with our variable of interest.
+It can be seen that most of the variables seem to have a moderate and positive correlation with our variable of interest.
 Although some variables lack observations, we must recall that this is just for the year 2010, and the whole dataset might have a different distribution pattern.
 
-## Proposed Model for the research {#model}
+### Proposed Model for the research {#model}
 
-Based on the informations obtained in @tbl-regression-results, we have selected variable that would best explain our dependent variable. The model is as follow:
+Based on the informations obtained from our preliminary results, we have selected variable that would best explain our dependent variable. The model is as follow:
 
@@
 
-We ran the regression based on our primary model described in the [proposed model for the research](#model) section of this article.
+### Regression Results
+
+We ran the regression based on our primary model described in the section [ 5.2.3](#model) of this article.
 The results obtained from the regression analysis are presented by @tbl-primary-regression-results.
@@
 
-The model showed a "good" descriptive performance based on it's adjusted R-squared value of 73.4%.
+The model showed a "good" descriptive performance based on it's adjusted R-squared value of 74.3%.
 The RMSE value for our model (RMSE=1213) is also quite lower compared to the NULL Model which showed a RMSE value of 2364.
 
-We performed the autocorrelation test. The results from the DW test presented by @tbl-DW-test-results confirms the absence of autocorrelation, therefore we can move forward with our model.
+#### Autocorrelation Test
+
+We performed the Durbin Watson autocorrelation test for our model.
+The results from the DW test, presented by @tbl-DW-test-results, confirms the absence of autocorrelation, therefore we can move forward with our model.
 
@@
 #| echo: FALSE
-#| out.width: "25%"
+#| out.width: "40%"
 knitr::include_graphics(here("results", "figures", "durbin_watson_test_results.png"))
@@
 
-We also performed the Fisher-type Augmented Dickey Fuller test for presence of unit root. The results confirmed that the not all series in our panel data are non-stationary.
+#### ADF unit root test
+
+We also performed the Fisher-type Augmented Dickey Fuller (ADF) test for presence of unit root. The results confirmed that the not all series in our panel data are non-stationary.
 
@@
 #| label: tbl-unit-root-results
-#| tbl-cap: "Panel Unit Root Test Results"
+#| tbl-cap: "ADF Unit Root Test Results"
 #| echo: FALSE
@@
 ```
+
+Since there is no significant presence of correlation and unit root, we will claim our model to be a valid model.
+Next, we will test the robustness of our OLS results throuh various machine learning techniques.
+
+#### Cross-Validation of OLS results
+
+In this section, we will observe how our model will perform across different subsets of datasets within our train dataset.As this is time-series dataset, we will consider the fact that the values across different years are correlated with each other.
+We, therefore, use three different subset of years with five years within each subset as training data within the train dataset.
+We will use the year after the training dataset as a testing dataset.
+Therefore, we ended up with three subsets of training data and three testing data.
+
+```{r}
+#| label: tbl-OLS-CV-results-full
+#| tbl-cap: "OLS cross-validation results "
+#| echo: FALSE
+
+htmltools::tags$iframe(
+  src = here("results", "tables", "OLS-CV-results-full.html"),
+  width = "100%",
+  height = "200px",
+  scrolling = "auto",
+  frameborder = "0"
+)
+```
+
+It can be observed that the RMSE values across the different combinations of train/test dataset show a somewhat consistent result.
+The RMSE value for the primary dataset was 1213.
+The results from CV, as observed in @tbl-ols-cv-results, show even less RMSE in some cases.
+This is a good thing as our model performs consistently better in cross-validation.
+
+```{r}
+#| label: tbl-ols-cv-results
+#| tbl-cap: "OLS cross-validation results summary"
+#| echo: FALSE
+
+htmltools::tags$div(
+  style = "display: flex; justify-content: center;",
+  htmltools::tags$iframe(
+    src = here("results", "tables", "OLS-CV-results.html"),
+    width = "40%",
+    height = "110px",
+    scrolling = "auto",
+    frameborder = "20"
+  )
+)
+```
+
+If we observe the mean RMSE across the three cross-validation subsets, the mean is lower than our primary dataset with a small standard deviation.
+
+### Random Forest model as an Alternative
+
+Next, we will perform our analyses with a different technique other than ordinary least squares method.
+We will use a Random Forest (RF) model to explain per capita health expenditure of a country.
+we will use a tree size of 300, and mtry value equal to 3.
+
+#### Random Forest Results
+
+@tbl-rf-importance-results represents the importance factor obtained from the random forest model.
+It can be observed that the variables that were found to have a significant and high impact on the per capita health expenditure in the OLS model also show a high importance in the random forest model as well, although results are different from OLS as well.
+
+```{r}
+#| label: tbl-rf-importance-results
+#| tbl-cap: "RF importance plot"
+#| echo: FALSE
+#| out.width: "100%"
+knitr::include_graphics(here("results", "figures", "rf_importance_plot.png"))
+```
+
+#### Performance of RF model
+
+The performance of our random forest model is quite exciting.
+The R-squared value obtained is quite high (0.988.
+The RMSE value is also 288, which is almost 1000 less than the RMSE value (1213) obtained from our OLS model.
+This shows that RF can explain our data way more appropriately than an OLS model.
+
+However, we need to test if this result is consistent across various dataset.
+
+```{r}
+#| label: tbl-rf-train-perform
+#| tbl-cap: "Random Forest TRAIN dataset performance"
+#| echo: FALSE
+
+htmltools::tags$iframe(
+  src = here("results", "tables", "rf_train_performance.html"),
+  width = "100%",
+  height = "200px",
+  scrolling = "auto",
+  frameborder = "0"
+)
+```
+
+#### Cross-validation of RF model
+
+In this section, we will try to see how our RF performs across various cross-validation datasets.
+We will perform cv based on the same combination of datasets we used for our OLS model.
+There are three different subsets of training dataset with five years in each dataset.
+Also, there will be a one year test dataset for the year immediately coming after the last year of testing dataset.
+
+```{r}
+#| label: tbl-rf-cv-results
+#| tbl-cap: "Random Forest CV Results"
+#| echo: FALSE
+
+htmltools::tags$iframe(
+  src = here("results", "tables", "rf_cv_results.html"),
+  width = "100%",
+  height = "300px",
+  scrolling = "auto",
+  frameborder = "0"
+)
+```
+
+The results obtained from the cross-validation as observed in @tbl-rf-cv-results show that the RMSE values have more than doubled in the cross-validation sets.Infact, the highest RMSE value observed in cv is as high as 2.6 times more than the RMSE obtained in the primary train dataset.
+
+```{r}
+#| label: tbl-rf-cv-summary
+#| tbl-cap: "Random Forest CV Summary"
+#| echo: FALSE
+
+htmltools::div(
+  style = "text-align: center;",
+  htmltools::tags$iframe(
+    src = here("results", "tables", "rf_cv_summary.html"),
+    width = "100%",
+    height = "120px",
+    scrolling = "auto",
+    frameborder = "0"
+  )
+)
+```
+
+The mean RMSE, 644, despite showing a standard deviation comparable to the OLS model, has a significantly higher value compared to RMSE (288) obtained with the train dataset.
+This means that our RF model is performing poorly in the case of different subsections of datasets.
+
+### Visual inspection of the two models in train dataset
+
+Finally, in this section, we will try to visually observe the performance of the two models (OLS and Random Forest) in our train dataset that we discussed and worked on till now.
+
+```{r}
+#| label: fig-pred-vs-obs-train
+#| fig-cap: "Predicted vs Observed values based on TRAIN dataset"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","pred_vs_obs_train.png"))
+```
+
+It can be observed in @fig-pred-vs-obs-train that the Random Forest clearly out-performs the OLS model in correctly predicting the outcomes.
+The datapoints for OLS regression are comparitively very hapazardly distributed compared to the datapoints of RF model.
+
+However, we should remember that although the RF model had an outstanding performance in our primary train dataset with RMSE almost 1000 less than RMSE observed in the OLS model, the cross-validation results showed significant fluctuations, due to which robustness of our RF model can be questioned.
+
+### Test dataset
+
+#### Perfrormance of OLS on test dataset
 
@@
 The adjusted R-squared in our test dataset has also dropped by 12.3% to 62%.
-We also tested Random Forest model as an alternative of the OLS model, and althouth Random Forest showed promising results in our train dataset, the results were worse than OLS model for the test dataset.
+
+#### Performance of Random Forest on test dataset
+
+We also ran our Random Forest model with the train test dataset. The performance metrics obtained from the model are presented in @tbl-rf-test-perform.
+
+```{r}
+#| label: tbl-rf-test-perform
+#| tbl-cap: "Random Forest TEST dataset performance"
+#| echo: FALSE
+
+htmltools::tags$iframe(
+  src = here("results", "tables", "rf_test_performance.html"),
+  width = "80%",
+  height = "250px",
+  scrolling = "auto",
+  frameborder = "0"
+)
+```
+
+We can observe that the RMSE value has again largely increased.
+The RMSE from test dataset is now as large as 4.6 times more than the RMSE obtained in the primary train dataset.
+This value is also double than the mean RMSE obtained from cross-validations.
+Therefore, there is a high suspicion that our our primary RF model sufferes from overfitting, although it performs relatively better than our OLS model.
+
+But, as of now, as our OLS model showed a consistent and robust result across all our datasets and robustness tests, we can conclude that OLS model should be preferred over the Random Forest model despite RF performs better than OLS.
+Our conclusion is that the RF model can not be trusted because of it's inconsistency and suspicious results.
+
+#### Visual inspection in test dataset
+
+Finally, we will visually observe the performance of both the models in the test dataset.
+@fig-pred-vs-obs-test shows the distribution of observed values and predicted values using test dataset for both the models.
+
+```{r}
+#| label: fig-pred-vs-obs-test
+#| fig-cap: "Predicted vs Observed values based on TEST dataset"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","pred_vs_obs_test.png"))
+```
+
+As expected based on the RMSE values, the RF model does perform better than our OLS model, However, due to inconsistency, we will prefer our OLS model over the RF model.
+
+One thing that should be noted is that the RF model should be optimized and formulated in an efficient manner because it clearly outperformed the OLS model in every comparision across different datasets despite the inconsistency.
 
@@
 [@Factor2015] has found that country with high level of corruption have a lower spending on health care as a percentage of GDP.
+[@Giammanco2019] have found that FDI is positively related to factors related to public health.
+[@Wang2015] argues that the impact of health expenditures on economic growth is somewhat ambiguous.
 
@@
 
-## Conclusions and Policy implications
+# Conclusions and Policy implications
 
@@
 
-## Data aquisition
+# Data aquisition
 
@@
 
-Link to the source: <https://databank.worldbank.org/source/world-development-indicators/Type/TABLE/preview/on#>
+The dataset can be assessed through this [link.](https://databank.worldbank.org/source/world-development-indicators/Type/TABLE/preview/on#)
 

### 2025-04-24 review-zuber/project-review-template.qmd
Part 5 review_ Mohammed Zuber

---
title: Project Review Template 
author: Mohammed Zuber
date: 4/23/2025
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: Factors affecting health sector expenditure in OECD countries

Name of project author(s): Prasanga Paudel

Name of project reviewer: Mohammed Zuber

# Score 
Overall Score: 3 (Sufficient). 
The project is complete, detailed, and well-documented. It includes a comprehensive analysis using both statistical and machine learning approaches, with thorough explanation and reproducibility. Minor improvements are suggested, but they do not detract from the overall high quality of the work.
# Instructions
## Background, Context and Motivation
How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

The background provides a clear rationale for studying health expenditure with references to prior studies. The objective to understand determinants such as GDP growth, corruption, and demographic variables is clearly articulated. However, a more detailed summary of the gaps in previous research and how this study addresses them would strengthen the motivation.

### Summary assessment 
* strong contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments
The research questions are explicitly stated and directly linked to the outcome variable (health expenditure per capita). The two-part objective — assessing key factors and their magnitude of influence — is well aligned with the analytical plan.

### Summary assessment
* question/hypotheses fully clear

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments
The data source (World Bank via OECD) is clearly identified, and the use of 462 observations from 38 countries is explained well. The GitHub repository is mentioned as the location for preprocessing steps, which is helpful. However, a codebook or variable dictionary is missing from the project folder, which makes it harder to understand the meaning and units of the 60+ variables used. Including even a brief data dictionary or a summary table with variable names and definitions would enhance transparency and usability.

### Summary assessment
* source and overall structure of data somewhat explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments
Exploratory analysis is well-executed with appropriate figures (e.g., trends, correlation plots) and justified variable selection. The well documented comments on how much time the code takes to run is commendable.

### Summary assessment
* essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

The OLS regression is appropriate and well-justified for panel data. The inclusion of cross-validation and RF as robustness checks adds value.

### Summary assessment
* strong and reasonable analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments
Figures and tables are informative and generally clear. However, a few visuals could benefit from better axis labeling and reduced clutter. The narrative does a good job walking the reader through the findings.

### Summary assessment
* results are very well presented

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

The discussion effectively synthesizes the key findings and connects them to existing literature, reinforcing the credibility of the results. The implications for public policy and international resource allocation are thoughtfully addressed. However, the discussion lacks a clear summary of the study’s strengths and limitations. For instance, a major strength is the comprehensive modeling approach, including OLS and Random Forest with validation across datasets. A notable limitation is the potential overfitting in the RF model and the absence of detail on how missing data and variable collinearity were handled. Including these would provide a more balanced and critical reflection on the results.

### Summary assessment
* minor parts wrong, missing or unclear

## Further comments
none

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

The project folder is clearly organized, with appropriately named files and a logical directory structure. There are no unnecessary or junk files, and the layout makes it easy to follow the workflow from raw data to final analysis and figures. The repository reflects good practices in reproducible research and is easy to navigate.

### Summary assessment
* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

The project is well documented, with clear and consistent comments throughout the code. Each line of code is annotated to explain its purpose, making the analysis easy to follow. The Rmd files guide the reader through the workflow step by step, and the folder structure further supports understanding. Overall, the documentation reflects thoughtful effort to ensure transparency and reproducibility.

### Summary assessment
* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

The project is fully reproducible, with all code executing smoothly and generating the expected outputs, including figures and tables. The analysis runs without any need for manual intervention, and all necessary data files and scripts are included in the repository. This reflects strong adherence to reproducible research standards.

### Summary assessment
* fully reproducible without issues

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

The project demonstrates a strong level of thoroughness. Multiple modeling approaches were considered — including OLS and Random Forest — and each was tested using cross-validation and out-of-sample performance.

### Summary assessment
* strong level of thorougness

## Further comments

Overall, Great Job!

### 2025-04-25 code/analysis-code/statistical-analysis.qmd

@@
 library(plm)
+library(zoo)
 library(lmtest)

### 2025-04-27 products/manuscript/Manuscript.qmd
intro #1

@@
 ## General Background Information
+
+Health expenditure is an important indicator of a country's committment to citizen's well-being and efficiency of it's healthcare systems.
+Among the OECD countries, the per capita health expenditure vary significantly ranging from 1,000 USD to over 10,000 [@2021].
+Understanding the factors that drive these differences is important for policymakers to optimize healthcare delivery and improving healhcare access.
+While GDP had been one of the primary determinant of health expending [@Newhouse1977], recent studies suggest that institutional, demographic, and socio-economic factors may also play a crucial role in determination of health expenditure among different countries [@Ziegler2017].
+
+The relationship between health expenditure and their determinants because particularly relevant in the context of recent global challenges.
+The recent COVID-19 pandemic also exposed the vulnerabilities in global healthcare systems with countries showing high pre-pandemic health expenditure generally demonstrating greater resilience [@WHO2022]
 

### 2025-04-27 code/robustness-code/robustness-code.qmd
Part 5 review_ Mohammed Zuber
intro #2

@@
 ```
+
 # Importing dataset
@@
                 full_width = FALSE,
-                position = "left") %>%
+                position = "center") %>%
   add_header_above(c("Training Period" = 1, "Test Year" = 1, 

### 2025-04-27 products/manuscript/Manuscript.qmd
intro #2

@@
 While GDP had been one of the primary determinant of health expending [@Newhouse1977], recent studies suggest that institutional, demographic, and socio-economic factors may also play a crucial role in determination of health expenditure among different countries [@Ziegler2017].
-
 The relationship between health expenditure and their determinants because particularly relevant in the context of recent global challenges.
@@
 In our analysis we will try to follow OLS multiple regression model.
+We will also use methodological inventions in comparing different models primarily OLS regression and Random Forest to address the recent suggestions in analytical apporaches in health economics [@ParkerLue2020]
 
@@
 #| echo: FALSE
-#| out.width: "40%"
+#| out.width: "30%"
 knitr::include_graphics(here("results", "figures", "durbin_watson_test_results.png"))
@@
 
+# Limitations
+
+Our analysis covers 2010-2020, excluding the full impact of the COVID-19 pandemic, which dramatically altered health expenditure patterns worldwide.
+Our OLS approach assumes linear relationships that may not capture complex interactions present in health systems, and although we employ Random Forest to address this, the "black box" nature of machine learning models limits policy interpretability.
+
 # Data aquisition
@@
 
-{{< pagebreak >}}
-
-# References
+# Reference

### 2025-05-01 products/manuscript/Manuscript.qmd
#codebook

@@
 Health expenditure is an important indicator of a country's committment to citizen's well-being and efficiency of it's healthcare systems.
-Among the OECD countries, the per capita health expenditure vary significantly ranging from 1,000 USD to over 10,000 [@2021].
+Among the OECD (Organisation for Economic Co-operation and Development) countries, the per capita health expenditure vary significantly ranging from 1,000 USD to over 10,000 [@2021].
 Understanding the factors that drive these differences is important for policymakers to optimize healthcare delivery and improving healhcare access.
@@
 However, the data obtained from this dataset is crude and needed to be treated before attempting analysis based on it.
-The details about the data cleaning process can be assessed from this [ GitHub](https://github.com/xprasanga/PAUDEL-MADA-project) repository.
+The details about the data cleaning process is available at [ GitHub](https://github.com/xprasanga/PAUDEL-MADA-project).
+
+The original data (2010-2020) in it's raw form was obtained from World Bank.
+However, the dataset needed processing, and some modification.
+The final dataset that is eligible for our use contains 462 observations and 1496 variables for the OECD countries.
+The variables include information on various aspect of a country's development indicators, such as information on education, health, population distribution, trade, energy access, GDP, infrastructure, market, corruption, investments, and so on.
+However, we will be using 80% of the dataset for our whole analysis and we will test the robustness of the results on the un-used 20% at the end of the analysis.
 
@@
 $X_{itk}$ is the $k^{th}$ independent variable ( k= 1 to 5, in the above equation but our research will likely have k>5) $y_{it}$ is the dependent variable, $b_0$ is the intercept, and $b_k$ is the estimated parameter for the $k^{th}$ independent variable.
-
-# Dataset
-
-The original data (2010-2020) in it's raw form was obtained from World Bank.
-However, the dataset needed processing, and some modification.
-The final dataset that is eligible for our use contains 462 observations and 1496 variables for the OECD countries.
-However, we will be using 80% of the dataset for our whole analysis and we will test the robustness of the results on the un-used 20% at the end of the analysis.
 

### 2025-05-01 code/eda-code/eda.qmd
codebook #2
#limitations

@@
 
-#combining plots ----
-#create a 3-column layout 
 combined_plot <- wrap_plots(plots, ncol = 3) +
@@
     theme = theme(
-      plot.title = element_text(size = 16, face = "bold"),
-      plot.subtitle = element_text(size = 12)
+      plot.title = element_text(size = 16, face = "bold", hjust = 0.5),  # Center title
+      plot.subtitle = element_text(size = 12, hjust = 0.5),              # Center subtitle
+      plot.caption = element_text(hjust = 0.5)                           # Center caption
     )

### 2025-05-01 products/manuscript/Manuscript.qmd
#limitations

@@
 
-In this section, we will we will explore the data based.
+In this section, we will we will explore the data and perform analysis based on ordinary least squares and random forest model.
 
@@
 As this is a time series data, it is quite difficult to visualize all the variables across the years for all the countries.
-So, we will try to keep it simple and see simple relations between variables within a year or trend of a particular variable for some countries across the years.
+So, we will try to keep it simple and see simple relations between variables within a year or trend of a particular variable for some countries across the years or the average trend.
 
@@
 #| label: fig-development-indicators-trends
-#| fig-cap: "Trends of Development Indicators (Average across countries)"
+#| fig-cap: "Trends of Development Indicators (average across countries)"
 #| echo: FALSE
@@
 we will use a tree size of 300, and mtry value equal to 3.
+The motivation behind using a random forest model is to check if we are being able to capture the relation between per capita health expenditure and our predictors variable well enough using the OLS model.
+The non-linear effects, if present, would be captured better by the RF model.
 
@@
 
-The results obtained from the cross-validation as observed in @tbl-rf-cv-results show that the RMSE values have more than doubled in the cross-validation sets.Infact, the highest RMSE value observed in cv is as high as 2.6 times more than the RMSE obtained in the primary train dataset.
+The results obtained from the cross-validation as observed in @tbl-rf-cv-results show that the RMSE values have more than doubled in the cross-validation sets.
+Infact, the highest RMSE value observed in cv is as high as 2.6 times more than the RMSE obtained in the primary train dataset.
 
@@
 
-Our results shows that annual GDP growth, control of corruption, life expectancy at birth, unemployment rate, and FDI all have a statistically significant effect on the per capita health expenditure of a country.
+Our results show that annual GDP growth, control of corruption, life expectancy at birth, unemployment rate, and FDI all have a statistically significant effect on the per capita health expenditure of a country.
 [@Nghiem2017] also found evidence that annual GDP and unemployment rate have a significant effect on the health expenditures, however the variables used in the studies are different.
-[@Factor2015] has found that country with high level of corruption have a lower spending on health care as a percentage of GDP.
-[@Giammanco2019] have found that FDI is positively related to factors related to public health.
-[@Wang2015] argues that the impact of health expenditures on economic growth is somewhat ambiguous.
+Similarly, [@Factor2015] has found that country with high level of corruption have a lower spending on health care as a percentage of GDP.
+In a similar study, [@Giammanco2019] have found that FDI is positively related to factors related to public health.
+Therefore, our result somewhat follow the general trend of findings in the field of health economics.
+In another study, [@Wang2015] argues that the impact of health expenditures on economic growth is somewhat ambiguous.
 
-Factors such as income share held by the lowest 20-percent population, trade as a percentage of GDP, access to electricity among rural popuation, and age dependency ration had no statistical significant impact on the health expenditure.
+Our study showed that factors such as income share held by the lowest 20-percent population, trade as a percentage of GDP, access to electricity among rural popuation, and age dependency ration had no statistical significant impact on the health expenditure.
 Among these variables, age dependency ratio was expected to have a significant impact as higher pressure on the working population to support the non-active population reduces the chances of surplus revenues to invest in different sectors of the economy.
@@
 A special concideration should be given to nations that have an increasing GDP, high corruption, and high unemployment rate, as these nations have lower investments in health sectors.
-These type of countries should therefore be priotized by world bank, World Health Organization, and other institutions where allocating their resources to help countries.
+These type of countries should therefore be priotized by world bank, World Health Organization, and other institutions who allocate their resources to help countries while also considering the fact that corruption and high unemployment should also be addressed as they show causal effect.
 
@@
 Our OLS approach assumes linear relationships that may not capture complex interactions present in health systems, and although we employ Random Forest to address this, the "black box" nature of machine learning models limits policy interpretability.
+As with time-series datasets for countries, our datasets had a lot of missing values for a lot of variables across different countries.
+This is quite common as countries differ in their time and duration of census and surveys and not all countries folllow the sam procedure to collect informations.
+Also, not all countries would collect the same set of information.
+We had to drop the missing values due to which we lost a lot of information which coould have been a valuable source of information.
 
@@
 
-The data for this research is available at the official World Bank- Data Bank website.It is a free database managed by world bank group and can be accessed anytime.
+The data for this research is available at the official World Bank- Data Bank website.
+It is a free database managed by world bank group and can be accessed anytime.
 We can apply various filters as per our need and export the data as a .CSV or an .xlsx file.

### 2025-05-02 code/analysis-code/statistical-analysis.qmd
intro #1
finalizing

@@
 
-Then we will import the dataset
+Then we will import the dataset.
 
@@
 Similarly, if there is a variable named "Unemployment rate" or "Total population", we will categorize them as "Social" variable.
+This will help us look into the data more easily
 
@@
 
-Now, we will save our regression result is a HTML table so that we can use to present it in our manuscript.
+Now, we will save our regression result is a HTML table so that we can use to present it in our manuscript or our supplementary material.
 

### 2025-05-02 products/manuscript/Manuscript.qmd
finalizing

@@
 where $e_{it}$ is the error term.
-$X_{itk}$ is the $k^{th}$ independent variable ( k= 1 to 5, in the above equation but our research will likely have k>5) $y_{it}$ is the dependent variable, $b_0$ is the intercept, and $b_k$ is the estimated parameter for the $k^{th}$ independent variable.
+$X_{itk}$ is the $k^{th}$ independent variable ( k= 1 to 5 in the above equation but our research will likely have k>5), $y_{it}$ is the dependent variable, $b_0$ is the intercept, and $b_k$ is the estimated parameter for the $k^{th}$ independent variable.
 

### 2025-05-03 products/manuscript/Manuscript.qmd
#final edits

@@
 
-# Abstract
+<strong>Abstract</strong>
 
-_We used Ordinary Least Squares method to analyze the factors affecting per capita health expenditure of a country among the OECD nations during 2010 to 2020.
-We found that GDP growth, control of corruption, and  unemployment rate among the youth in the labor force have a significant and an important role in determining the amount of money spent on health sector by the country.
-Foreign Direct Investment had a significant but low magnitude impact.
-Factors such as income share held by the lowest 20-percent population, trade as a percentage of GDP, access to electricity among rural popuation, and age dependency ration had no statistical significant impact on the health expenditure.
+_Understanding the determinants of health expenditures is a critical job in order to optimize the healthcare systems and policymaking.
+This study utilizes the Ordinary Least Squares method to analyze the factors affecting per capita health expenditure of a country among the OECD nations during 2010 to 2020.
+Our study identifies that GDP growth, control of corruption, and  unemployment rate among the youth in the labor force have a significant and an important role in determining the amount of money spent on health sector by the country, while Foreign Direct Investment exhibits a significant but low magnitude impact.
+Factors such as income share held by the lowest 20-percent population, trade as a percentage of GDP, access to electricity among rural popuation, and age dependency ration showed no statistical significant impact on the health expenditure.
 We also applied Random Forest model to explain the factors affecting per capita health expenditure of a country.
-The results of both the OLS model and Random forest was tested using cross-validation and train/test technique, and we found that our OLS model shows a consistent result across all test.
+The results of both the OLS model and Random forest was tested using cross-validation and train/test technique, and we found that our OLS model shows a consistent result across all tests.
 Although Random Model was found to have a better explanation capability, the inconsistensies raised a question on it's validity.
+Our findings highlight the importance of economic and institutional factors in shaping the health expenditure and offer insights for policimakers aiming to enhance healthcare accessibility and efficiency.
 _
@@
 
-## General Background Information
+## Health expenditure as an indicator
 
 Health expenditure is an important indicator of a country's committment to citizen's well-being and efficiency of it's healthcare systems.
-Among the OECD (Organisation for Economic Co-operation and Development) countries, the per capita health expenditure vary significantly ranging from 1,000 USD to over 10,000 [@2021].
-Understanding the factors that drive these differences is important for policymakers to optimize healthcare delivery and improving healhcare access.
+Among the OECD (Organisation for Economic Co-operation and Development) countries, the per capita health expenditure vary significantly ranging from 1,000 USD to over 10,000 USD [@2021].
+Understanding the factors that drive these differences is important for policymakers to optimize healthcare delivery and improving healthcare access.
 While GDP had been one of the primary determinant of health expending [@Newhouse1977], recent studies suggest that institutional, demographic, and socio-economic factors may also play a crucial role in determination of health expenditure among different countries [@Ziegler2017].
@@
 
-We then ran our initial regression with "Health Expenditure Per Capita $USD" as dependent variable, and variables from each of the category as independent variables in separate regressions.
-The regression results for this preliminary test can be observed under this [ GitHub](https://github.com/xprasanga/PAUDEL-MADA-project) repository under supplementary-materials.
-
+We then ran our initial regression with "Health Expenditure Per Capita (USD)" as dependent variable, and variables from each of the category as independent variables in separate regressions.
+The motivation here was to find the most influential factor from each category.
+The regression results for this preliminary test can be found in  [ GitHub](https://github.com/xprasanga/PAUDEL-MADA-project).
 After observing the relation of almost 60 important variables with our variable of interest "Health Expenditure Per Capita $USD", we filtered few variables worth pursuing, and we will discuss them in the sections that will come.
@@
 #| label: fig-correlation-p2
-#| fig-cap: "Correlation based on 2010 data"
+#| fig-cap: "Additional correlation based on 2010 data"
 #| echo: FALSE
@@
 
-A special concideration should be given to nations that have an increasing GDP, high corruption, and high unemployment rate, as these nations have lower investments in health sectors.
+A special consideration should be given to nations that have an increasing GDP, high corruption, and high unemployment rate, as these nations have lower investments in health sectors.
 These type of countries should therefore be priotized by world bank, World Health Organization, and other institutions who allocate their resources to help countries while also considering the fact that corruption and high unemployment should also be addressed as they show causal effect.
@@
 As with time-series datasets for countries, our datasets had a lot of missing values for a lot of variables across different countries.
-This is quite common as countries differ in their time and duration of census and surveys and not all countries folllow the sam procedure to collect informations.
+This is quite common as countries differ in their time and duration of census and surveys and not all countries folllow the same procedure to collect informations.
 Also, not all countries would collect the same set of information.
@@
 
-The data for this research is available at the official World Bank- Data Bank website.
-It is a free database managed by world bank group and can be accessed anytime.
+The data for this research is available at the official World Bank- Data Bank website accessible [here.
+](https://databank.worldbank.org/source/world-development-indicators/Type/TABLE/preview/on#) It is a free database managed by world bank group and can be accessed anytime.
 We can apply various filters as per our need and export the data as a .CSV or an .xlsx file.
-
-The dataset can be assessed through this [link.](https://databank.worldbank.org/source/world-development-indicators/Type/TABLE/preview/on#)
-
-# Reference

### 2025-05-03 review-Crammer/Cramer-project-review.qmd
# appropriately placing crammer's review

--- 2025-04-24 review-zuber/project-review-template.qmd
@@
 ---
-title: Project Review Template 
-author: Mohammed Zuber
-date: 4/23/2025
+title: Project Review
+author: Annalise Cramer
+date: 4/25/25
 format: 
@@
 
-Name of project reviewer: Mohammed Zuber
+Name of project reviewer: Annalise Cramer
 
-# Score 
-Overall Score: 3 (Sufficient). 
-The project is complete, detailed, and well-documented.
-It includes a comprehensive analysis using both statistical and machine learning approaches, with thorough explanation and reproducibility.
-Minor improvements are suggested, but they do not detract from the overall high quality of the work.
-# Instructions
 ## Background, Context and Motivation
@@
 
-The background provides a clear rationale for studying health expenditure with references to prior studies.
-The objective to understand determinants such as GDP growth, corruption, and demographic variables is clearly articulated.
-However, a more detailed summary of the gaps in previous research and how this study addresses them would strengthen the motivation.
+I think defined OECD starting in the abstract would be helpful.
+More background information on the problem would be helpful-- why do we care about health expenditure?
+Why look at different countries?
+As someone who doesn't know much about this field, I would guess this has something to do with different healthcare systems in various countires and is this some sort of measure of efficency of the different systems?
+Idk, but add some more to fill it out.
 
-### Summary assessment 
-* strong contextualization and motivation
+### Summary assessment (PICK ONE, DELETE THE OTHERS)
+
+Some contextualization and motivation
 
@@
 ### Feedback and Comments
-The research questions are explicitly stated and directly linked to the outcome variable (health expenditure per capita).
-The two-part objective — assessing key factors and their magnitude of influence — is well aligned with the analytical plan.
+
+I liked the part about how this research could inform policy.
 
 ### Summary assessment
-* question/hypotheses fully clear
+
+Question/hypotheses fully clear
 
@@
 ### Feedback and Comments
-The data source (World Bank via OECD) is clearly identified, and the use of 462 observations from 38 countries is explained well.
-The GitHub repository is mentioned as the location for preprocessing steps, which is helpful.
-However, a codebook or variable dictionary is missing from the project folder, which makes it harder to understand the meaning and units of the 60+ variables used.
-Including even a brief data dictionary or a summary table with variable names and definitions would enhance transparency and usability.
+
+Data source is well explained and 38 countries is described. What years does this data cover, and what types of variables are in this dataset?
+
+Edit: I see there's more information on the data that answers some of my questions. I think combining this information into one place would be helpful.
 
 ### Summary assessment
-* source and overall structure of data somewhat explained
+
+Source and overall structure of data somewhat explained
 
@@
 ### Feedback and Comments
-Exploratory analysis is well-executed with appropriate figures (e.g., trends, correlation plots) and justified variable selection.
-The well documented comments on how much time the code takes to run is commendable.
+
+This is some of the best documented code I've seen all semester.
+I was easily able to follow along with data manipulation and what the figures were for, and the figures looked excellent.
+I got an error at line 435 for france_data, maybe you had a line in there ealier that got removed?
+As I continue, correlation plots look great.
+You provided plenty of description in your manuscript for EDA.
 
 ### Summary assessment
-* essentially no weaknesses in wrangling and exploratory component
+
+Essentially no weaknesses in wrangling and exploratory component
 
@@
 
-The OLS regression is appropriate and well-justified for panel data. The inclusion of cross-validation and RF as robustness checks adds value.
+I'm able to follow the analysis successfully.
+First, you group the variables into theme categories.
+I'm used to seeing forwards or backawards selection for model building, so I'm a lttle unfamiliar with this method-- could you explain more?
+The regression ran smoothly.
+
+For 5.2.3 in your manuscript where you type the regression equation, at line 163 in your manuscript, the equation needs some reformatting. Maybe make the vairable names abgreviated subscripts?
+
+Adding in a little bit more motivation for exploring the random forest model would help
+
+You can save yourself some time but just loading library(tidyverse), which includes: ggplot2, dplyr, tidyr, readr, purrr, tibble, stringr, and forcats, and some others too.
+It's a little easier than typing a bunch of those out :)
 
 ### Summary assessment
-* strong and reasonable analysis
+
+Strong and reasonable analysis
 
@@
 ### Feedback and Comments
-Figures and tables are informative and generally clear.
-However, a few visuals could benefit from better axis labeling and reduced clutter.
-The narrative does a good job walking the reader through the findings.
+
+10/10 figures and tables
 
 ### Summary assessment
-* results are very well presented
+
+Results are very well presented
 
@@
 
-The discussion effectively synthesizes the key findings and connects them to existing literature, reinforcing the credibility of the results.
-The implications for public policy and international resource allocation are thoughtfully addressed.
-However, the discussion lacks a clear summary of the study’s strengths and limitations.
-For instance, a major strength is the comprehensive modeling approach, including OLS and Random Forest with validation across datasets.
-A notable limitation is the potential overfitting in the RF model and the absence of detail on how missing data and variable collinearity were handled.
-Including these would provide a more balanced and critical reflection on the results.
+Good explanation and interpretation of results and their implications. Could use discussion of limitations.
 
 ### Summary assessment
-* minor parts wrong, missing or unclear
+
+Minor parts wrong, missing or unclear
 
 ## Further comments
-none
 
@@
 
-The project folder is clearly organized, with appropriately named files and a logical directory structure.
-There are no unnecessary or junk files, and the layout makes it easy to follow the workflow from raw data to final analysis and figures.
-The repository reflects good practices in reproducible research and is easy to navigate.
+I was able to find everything with no issues. Copy/pasting the codebook into some of the EDA code could be helpful.
 
 ### Summary assessment
-* well structured
+
+Well structured
 
@@
 
-The project is well documented, with clear and consistent comments throughout the code.
-Each line of code is annotated to explain its purpose, making the analysis easy to follow.
-The Rmd files guide the reader through the workflow step by step, and the folder structure further supports understanding.
-Overall, the documentation reflects thoughtful effort to ensure transparency and reproducibility.
+Excellent no notes
 
 ### Summary assessment
-* fully and well documented
+
+Fully and well documented
 
@@
 
-The project is fully reproducible, with all code executing smoothly and generating the expected outputs, including figures and tables.
-The analysis runs without any need for manual intervention, and all necessary data files and scripts are included in the repository.
-This reflects strong adherence to reproducible research standards.
+I was able to run everything with no issues
 
 ### Summary assessment
-* fully reproducible without issues
+
+Fully reproducible without issues
 
@@
 
-The project demonstrates a strong level of thoroughness.
-Multiple modeling approaches were considered — including OLS and Random Forest — and each was tested using cross-validation and out-of-sample performance.
+There was lots of exploration on which vairables to include in the model, and ML altetnatives were epxlored.
 
 ### Summary assessment
-* strong level of thorougness
+
+Strong level of thorougness
 
@@
 
-Overall, Great Job!
+Overall you're in really good shape. Filling it out a little more in some parts will make it even better.

### 2025-05-03 code/eda-code/eda.qmd
Update eda.qmd

--- 2025-04-19 code/eda-code/eda.qmd
@@
     labs(title = plot_title, y = y_label, x = NULL) +
-    theme_minimal(base_size = 12) +
-    theme(plot.title = element_text(size = 11))
+    theme_minimal(base_size = 18) +
+    theme(plot.title = element_text(size = 18))
 }
@@
     title = "Trends in Key Development Indicators",
-    subtitle = "Country averages across time",
+    subtitle = "(The x-axis is labeled at the top of each plot, y-axis is labeled at the left side )",
     caption = "Source: World Development Indicators",
     theme = theme(
-      plot.title = element_text(size = 16, face = "bold"),
-      plot.subtitle = element_text(size = 12)
+      plot.title = element_text(size = 19, face = "bold", hjust = 0.5),  # Center title
+      plot.subtitle = element_text(size = 17, hjust = 0.5),              # Center subtitle
+      plot.caption = element_text(size = 14)                         
     )
@@
   width = 16,
-  height = 14,
+  height = 18,
   dpi = 300

### 2025-05-03 code/analysis-code/statistical-analysis.qmd
removing underscore

@@
 # Save the table as a HTML file
-html_file <- here("products", "supplementary-files", "regression_results.html")
+html_file <- here("results", "supplementary-tables", "regression_results.html")
 save_kable(table_html, file = html_file)
@@
 
+# Define a named vector for pretty variable names
+pretty_names <- c(
+  "(Intercept)" = "Intercept",
+  "GDP_growth_annual_" = "GDP Growth (annual %)",
+  "Income_share_held_by_lowest_20" = "Income Share (Lowest 20%)",
+  "Control_of_Corruption_Estimate" = "Control of Corruption",
+  "Life_expectancy_at_birth_total_years" = "Life Expectancy (years)",
+  "Unemployment_youth_total__of_total_labor_force_ages_1524_modeled_ILO_estimate" = "Youth Unemployment (% of labor force, ages 15-24)",
+  "Trade__of_GDP" = "Trade (% of GDP)",
+  "Foreign_direct_investment_net_inflows_BoP_current_US" = "FDI Net Inflows (current US$)",
+  "Access_to_electricity_rural__of_rural_population" = "Rural Electricity Access (% of rural population)",
+  "Age_dependency_ratio__of_workingage_population" = "Age Dependency Ratio"
+)
+
+# Generate the table with formatted names
 modelsummary(
@@
   output = here("results", "tables", "primary_regression_results.png"),
-  title = "This table represents the primary regresssion model",
+  title = "",
   stars = c('*' = 0.1, '**' = 0.05, '***' = 0.01),
+  coef_rename = pretty_names,
   notes = list(

### 2025-05-03 code/eda-code/eda.qmd
removing underscore

@@
 
-# Function to create a scatter plot with correlation
+# Function to create a scatter plot with correlation and formatted x-axis labels
 create_corr_plot <- function(x_var, df, point_color) {
+  # Create a named vector for neat variable names
+  pretty_names <- c(
+    "GDP_growth_annual_" = "GDP Growth (annual %)",
+    "Income_share_held_by_lowest_20" = "Income Share held by lowest 20%",
+    "Control_of_Corruption_Estimate" = "Control of Corruption Estimate",
+    "Life_expectancy_at_birth_total_years" = "Life expectancy at birth (years)",
+    "Literacy_rate_adult_total__of_people_ages_15_and_above" = "Adult literacy rate (% ages 15+)",
+    "Trade__of_GDP" = "Trade (% of GDP)",
+    "Foreign_direct_investment_net_inflows_BoP_current_US" = "FDI Net inflows (current US$)",
+    "Age_dependency_ratio__of_workingage_population" = "Age Dependency Ratio (% of working-age population)"
+  )
+
   cor_value <- round(cor(df[[x_var]], df$Current_health_expenditure_per_capita_current_US, use = "complete.obs"), 2)
@@
   ggplot(df, aes_string(x = x_var, y = "Current_health_expenditure_per_capita_current_US")) +
-    geom_point(color = point_color, alpha = 0.6) +  # Set color based on input
+    geom_point(color = point_color, alpha = 0.6) +
     geom_smooth(method = "lm", color = "red", se = FALSE) +
-    labs(title = paste("Correlation:", cor_value), x = x_var, y = "Health Expenditure per Capita") +
+    labs(
+      title = paste("Correlation:", cor_value), 
+      x = pretty_names[x_var], 
+      y = "Health Expenditure per Capita (current US$)"
+    ) +
     theme_minimal()

### 2025-05-03 products/manuscript/Manuscript.qmd
removing underscore

@@
 #| echo: FALSE
-#| out.width: "70%"
+#| out.width: "55%"
 knitr::include_graphics(here("results", "tables", "primary_regression_results.png"))
@@
 #| echo: FALSE
-#| out.width: "30%"
+#| out.width: "27%"
 knitr::include_graphics(here("results", "figures", "durbin_watson_test_results.png"))

### 2025-05-03 code/analysis-code/statistical-analysis.qmd
made tables formal

--- 2025-05-02 code/analysis-code/statistical-analysis.qmd
@@
 
+# Define a named vector for clean variable names
+clean_names <- c(
+  "(Intercept)" = "Intercept",
+  "GDP_growth_annual_" = "GDP Growth (annual %)",
+  "Income_share_held_by_lowest_20" = "Income Share (Lowest 20%)",
+  "Control_of_Corruption_Estimate" = "Control of Corruption",
+  "Life_expectancy_at_birth_total_years" = "Life Expectancy (years)",
+  "Unemployment_youth_total__of_total_labor_force_ages_1524_modeled_ILO_estimate" = "Youth Unemployment (% of labor force, ages 15-24)",
+  "Trade__of_GDP" = "Trade (% of GDP)",
+  "Foreign_direct_investment_net_inflows_BoP_current_US" = "FDI Net Inflows (current US$)",
+  "Access_to_electricity_rural__of_rural_population" = "Rural Electricity Access (% of rural population)",
+  "Age_dependency_ratio__of_workingage_population" = "Age Dependency Ratio"
+)
+
+# Generate the table with formatted names
 modelsummary(
@@
   output = here("results", "tables", "primary_regression_results.png"),
-  title = "This table represents the primary regresssion model",
+  title = "",
   stars = c('*' = 0.1, '**' = 0.05, '***' = 0.01),
+  coef_rename = clean_names,
   notes = list(
@@
   Statistic = dw_statistic,
-  P_Value = dw_p_value
+  `p-value` = dw_p_value,  
+  check.names = FALSE      # Preserve the column name formatting
 )
@@
   fmt_number(
-    columns = vars(Statistic, P_Value),
+    columns = everything(),  # Applies to all columns
     decimals = 4

### 2025-05-03 code/robustness-code/robustness-code.qmd
made tables formal

@@
 library(timetk)   # For time_series_cv()
-library(rsample)  # For analysis() and assessment()
+library(rsample)  # For analysis() and assessmen()
 library(here)
+library(ggplot2)
 library(modelsummary)
@@
   output = here("results", "tables", "testresult.png"),
-  title = "This table represents regresssion results from the test data",
+  title = "",
   stars = c('*' = 0.1, '**' = 0.05, '***' = 0.01),
+  coef_rename = clean_names,
   notes = list(
@@
 
-    data.frame(
-      Train_Years = paste(min(train_cv$Year), max(train_cv$Year), sep = "-"),
-      Test_Year = max(test_cv$Year),
+results <- data.frame(
+  `Train Years` = paste(min(train_cv$Year), max(train_cv$Year), sep = "-"),
+  `Test Year` = max(test_cv$Year),
   RMSE = rmse_val,
-      stringsAsFactors = FALSE
+  stringsAsFactors = FALSE,
+  check.names = FALSE  # Preserves the spaces in column names
 )
@@
   summarise(
-    Mean_RMSE = mean(RMSE, na.rm = TRUE),
-    SD_RMSE = sd(RMSE, na.rm = TRUE),
+    `Mean RMSE` = mean(RMSE, na.rm = TRUE),
+    `Std.D (RMSE)` = sd(RMSE, na.rm = TRUE)
   )
@@
 
-# Extract variable importance
+# First create a named vector for pretty variable names
+neat_names <- c(
+  "GDP_growth_annual_" = "GDP Growth",
+  "Income_share_held_by_lowest_20" = "Income Share (Lowest 20%)",
+  "Control_of_Corruption_Estimate" = "Control of Corruption",
+  "Life_expectancy_at_birth_total_years" = "Life Expectancy",
+  "Unemployment_youth_total__of_total_labor_force_ages_1524_modeled_ILO_estimate" = "Youth Unemployment",
+  "Trade__of_GDP" = "Trade (% of GDP)",
+  "Foreign_direct_investment_net_inflows_BoP_current_US" = "FDI Net Inflows",
+  "Access_to_electricity_rural__of_rural_population" = "Rural Electricity Access",
+  "Age_dependency_ratio__of_workingage_population" = "Age Dependency Ratio"
+)
+
+# Extract and prepare importance data with neat names
 importance_df <- data.frame(
   Variable = names(rf_full_model$variable.importance),
-  Importance = rf_full_model$variable.importance
+  Importance = rf_full_model$variable.importance,
+  stringsAsFactors = FALSE
 ) %>% 
+  mutate(Variable = neat_names[Variable]) %>%  # Replace with neat names
   arrange(desc(Importance))
 
-# Create importance plot
-library(ggplot2)
+# Create importance plot with better formatting
 importance_plot <- ggplot(importance_df, aes(x = reorder(Variable, Importance), 
-                         y = Importance)) +  # Fixed this line - removed extra parenthesis
-  geom_col(fill = "steelblue") +  # Simplified this line
+                                           y = Importance)) +
+  geom_col(fill = "steelblue", width = 0.7) +
   coord_flip() +
   labs(title = "Random Forest Variable Importance",
-       x = "Predictor Variables",
+       x = NULL,  #  redundant axis title
        y = "Importance (Permutation)") +
-  theme_minimal()
+  theme_minimal(base_size = 12) +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.grid.minor = element_blank(),
+    plot.title = element_text(hjust = 0.5, face = "bold")
+  )
 
-# Save the plot
+# Saving the plot
 ggsave(here("results", "figures", "rf_importance_plot.png"), 
        plot = importance_plot,
-       width = 8, height = 6, dpi = 300)
+       width = 10,  #  wider for better text fit
+       height = 6, 
+       dpi = 300)
 ```
@@
 ```{r}
-# Select only the desired columns
+# Select and rename columns with clean names
 rf_results_table <- rf_results %>%
-  select(Train_Years, Test_Year, RMSE, Predictors_Used, mtry_Used)
+  select(
+    `Train Years` = Train_Years,
+    `Test Year` = Test_Year,
+    RMSE,
+    `Predictors Used` = Predictors_Used,
+    `mtry Used` = mtry_Used
+  )
 
-# Create and save the HTML table
+# Create and save the HTML table with formatted headers
 rf_results_html <- kable(rf_results_table, 
@@
                         digits = 2,
+                        col.names = c("Train Years", "Test Year", "RMSE", 
+                                     "No. of predictors", "mtry"),
                         caption = "") %>%
@@
                 position = "center") %>%
-  add_header_above(c("Training Period" = 1, "Test Year" = 1, 
-                     "Performance" = 1, "Model Parameters" = 2))
+  add_header_above(c(" " = 1, " " = 1,  # Empty spaces for first two columns
+                    "Performance" = 1, 
+                    "Model Parameters" = 2)) %>%
+  row_spec(0, bold = TRUE)  # Make header row bold
 

### 2025-05-03 products/manuscript/Manuscript.qmd
made tables formal

@@
 #| label: tbl-rf-importance-results
-#| tbl-cap: "RF importance plot"
+#| tbl-cap: "Random Forest importance plot"
 #| echo: FALSE
@@
 #| echo: FALSE
-#| out.width: "70%"
+#| out.width: "55%"
 knitr::include_graphics(here("results", "tables", "testresult.png"))
